### 2025-05-03 doc/paper.qmd
Initial commit of huwiwiQuartoThesis based on CQA template

---
title: |
  | How Much New Information Is There in Earnings?
  | Corporate Decision-Making and Quantitative Analysis
  | Winter 2024/25 - Individual Report\vspace{1cm}
author:
  - name: Mel Mzv
    email: m@student.hu-berlin.de
    affiliations:
      - Humboldt-Universität zu Berlin  
date: today
date-format: MMM D, YYYY [\vspace{0cm}] # Adjusted the vertical space after the date from 1cm
abstract: |
  | This project uses the TRR 266 Template for Reproducible Empirical Accounting Research (TREAT) to provide infrastructure for open science-oriented empirical projects. Leveraging data from the CRSP and Compustat databases, alongside Worldscope and Datastream via WRDS, this repository showcases a reproducible workflow integrating Python scripts for data preparation, analysis, and visualization. Integrating multiple databases adds complexity, requiring a detailed understanding of their structures and careful scripting to extract, align, and analyze data effectively. The project replicates and extends Ball and Shivakumar (2008) to analyze the informativeness of quarterly earnings announcements and their contribution to share price movements, highlighting their critical role in investment decisions and impact on investors and analysts. Key tasks include replicating and comparing original results, extending the analysis to 2007–2023, and applying the methodology to a non-U.S. country. The project also documents research design choices, discusses variations between original and reproduced results, and provides insights into earnings informativeness across different timeframes and jurisdictions. Additionally, it sketches a research design for a non-archival study to evaluate the paper’s findings. This code base, adapted from TREAT, demonstrates how the template applies to the project and serves as a structured guide for reproducible empirical research in accounting.
  <!--
  | \vspace{6cm} commenting away to decrease distance
  -->

bibliography: references.bib
biblio-style: apsr
format:
  pdf:
    documentclass: article
    number-sections: true
    toc: true
    lof: true  # Enables List of Figures
    lot: false  # Disables List of Tables
fig_caption: yes
fontsize: 11pt
ident: yes
always_allow_html: yes
number-sections: true 
header-includes:
  - \usepackage[nolists]{endfloat}    
  - \usepackage{setspace}\doublespacing
  - \setlength{\parindent}{4em}
  - \setlength{\parskip}{0em}
  - \usepackage[hang,flushmargin]{footmisc}
  - \usepackage{caption} 
  - \captionsetup[table]{skip=24pt,font=bf}
  - \usepackage{array}
  - \usepackage{threeparttable}
  - \usepackage{adjustbox}
  - \usepackage{graphicx}
  - \usepackage{csquotes}
  - \usepackage{indentfirst}  # Added this line to ensure the first paragraph is indented for better readability
  - \usepackage[margin=1in]{geometry}
---

\pagebreak

# List of Abbreviations
\setlength{\parindent}{0em}
**BHR**: Buy-and-Hold Returns   
**CRSP**: Center for Research in Security Prices  
**CQA**: Corporate Decision-Making and Quantitative Analysis  
**DS**: Datastream  
**DV**: Dependent Variable  
**HU**: Humboldt-Universität zu Berlin  
**IDE**: Integrated Development Environment  
**IV**: Independent Variable  
**TREAT**: TRR 266 Template for Reproducible Empirical Accounting Research  
**US**: United States  
**WRDS**: Wharton Research Data Services  
**WSCP**: Worldscope

\setlength{\parindent}{4em}

\pagebreak

# Introduction {#sec-introduction}
The aim of this paper is to illustrate the use of open science tools in empirical accounting research. This project builds on the methodological framework established in the Corporate Decision-Making and Quantitative Analysis (CQA) course, which explored a range of empirical research methods—including archival analysis, field experiments, and survey-based approaches—to develop a comprehensive understanding of corporate decision-making and quantitative analysis. Expanding on prior empirical research conducted in Assignments I and II, this paper extends the empirical focus of the course. Assignment I examined audit market concentration in the EU using Transparency Reports data, offering insights into the dominance of major audit firms and the structure of the European audit market. Assignment II explored the economic implications of corporate financial reporting, focusing on @graham_economic_2005 to analyze managerial decision-making, earnings management strategies, and voluntary disclosure practices.

Building on these foundations, Assignment III investigates the informativeness of quarterly earnings announcements and their contribution to annual share price movements. The project replicates and extends @ball_how_2008, assessing the extent to which earnings releases serve as a source of new information for financial markets. By integrating multiple datasets from WRDS and applying an event-study methodology, this study aims to evaluate the role of earnings disclosures in market efficiency, showcasing the valuation purpose of accounting. At the same time, earnings reports may play a crucial role in contracting by influencing managerial compensation, debt agreements, and other contractual mechanisms tied to financial performance [@ball_how_2008]. The integration of multiple databases adds complexity, requiring a detailed understanding of their structures and careful scripting to extract, align, and analyze data effectively.

Beyond empirical replication, the project also documents research design choices, replication steps, and explicit assumptions made whenever the original paper was unclear on how to proceed. It discusses variations between original and reproduced results and provides insights into the timeliness and market impact of earnings announcements across different timeframes and jurisdictions. Additionally, it sketches a research design for a non-archival study to evaluate the paper’s findings through an alternative methodological lens. This project leverages the TRR 266 Template for Reproducible Empirical Accounting Research (TREAT) to establish an open science-oriented infrastructure, ensuring transparency and replicability for empirical accounting research.

Earnings announcements provide investors with valuable information about a firm’s market value, with stock prices reacting significantly when earnings news deviate from expectations [@fink_review_2021, p.2]. The seminal study by @ball_empirical_1968 was the first to document this relationship, showing that stock prices anticipate earnings surprises, with most of the market reaction occurring before the official announcement, suggesting that earnings reports primarily confirm rather than introduce new information. Over the past decades, more than a thousand studies have examined the interplay between capital markets and financial statements, a research stream that originated with Ball and Brown (1968)  [@kothari_capital_2001].

Building on this foundation, the original study by @ball_how_2008 investigates the extent to which quarterly earnings announcements contribute new information to the market, assessing their role in shaping annual share price movements. By estimating the R² from regressions of annual stock returns on earnings announcement window returns, the paper quantifies the informativeness of earnings releases, finding that they account for only 1% to 2% of total annual volatility. This challenges the assumption that earnings provide substantial new information and instead suggests a confirmatory role in financial reporting. These findings align with the semi-strong form of market efficiency, which suggests that stock prices already reflect all publicly available information, including prior signals about earnings. Unlike traditional event studies that assume market efficiency, @ball_how_2008 allow for potential market mispricing by not imposing a unit slope restriction in their regressions, meaning that their approach can capture under- or overreaction to earnings announcements. They also emphasize that extending the event window beyond the earnings announcement period would shift the research focus from earnings “surprise” to broader financial reporting effects, such as its role in debt contracting and executive compensation.

While @ball_how_2008 provide key insights into the informativeness of earnings announcements, their study was conducted within a specific timeframe and market context. Since then, financial markets have undergone significant changes due to regulatory reforms, economic crises, and technological advancements in financial information processing. Moreover, the extent to which earnings informativeness varies across international markets remains underexplored. By replicating their study with updated data from 2007–2023 and applying the methodology to Canada as a non-U.S. country, this project reassesses the robustness of their findings. Through this approach, it critically evaluates the generalizability of the original results.

The paper is structured into sections corresponding to Tasks 1–3 (@sec-task1, @sec-task2, @sec-task3), each detailing the research design choices and assumptions, documenting the replication steps and results. @sec-task1 follows an exact replication approach by replicating key tables and figures from the original study using US data (1972–2006) to establish baseline results. @sec-task2 extends the analysis through empirical generalization, applying the same methodology to 2007–2023 US data to compare how results evolve over time. @sec-task3 employs a generalization and extension approach by applying the methodology to a non-US country, requiring adjustments for different databases and market structures. This framework aligns with the replication taxonomy outlined by @salterio_replication_2022. Finally, @sec-task4 sketches the survey design as a non-archival study that allows evaluating the key findings of the seminal paper, following Bloomfield’s (2016) triangulation principle. The concluding remarks are provided in @sec-conclusion.

# Task 1 - Replication of Key Tables and Figures {#sec-task1}
Task 1 is the most comprehensive, replicating the methodology of key tables and figures from @ball_how_2008 using daily US stock return data (1972–2006) to establish baseline results, ensuring alignment with the original findings before extending the analysis in @sec-task2. By reconstructing Table 1 (Panel A), Table 2 (all panels), and Figure 1 (all panels), the task evaluates the robustness of the original results and identifies potential discrepancies, offering insights into data consistency, methodology, and market dynamics over the examined period. The use of daily CRSP data is essential for capturing short-term stock price movements around earnings announcements, ensuring precise event window calculations. The discussion of replication findings is presented in @sec-results1.

## Research Design Choices and Assumptions {#sec-research_design_assumptions1}
Following @ball_how_2008, I focus on earnings announcements and stock returns from 1972 to 2006, ensuring alignment with the original study’s methodology. This replication mirrors the research design, maintaining consistency in sample selection, event window definitions, and return calculations.

Based on the original paper, I do not gather analyst expectations or estimate earnings deviations, requiring only stock return data and earnings announcement dates, simplifying data collection.

To ensure clarity where @ball_how_2008 lacks explicit guidance, I impose the following assumptions:

1.	Earnings announcement windows are defined as the three-day period surrounding the release (-1 to +1). According to @center_for_research_in_security_prices_crsp_crsp_2024 guide,since CRSP excludes weekends and holidays, if Day 0 falls on a non-trading day, it shifts to the next trading day. Similarly, Day -1 or +1 adjusts to the nearest available trading day.

2.	Firm-years are selected based on available CRSP stock return data. While @ball_how_2008 restricts the sample to firms with at least 240 trading days per year, I follow course guidance and do not impose this restriction, allowing for a slightly larger sample.

3.	Compustat includes active and inactive firms; since the original study does not filter based on firm activity, all firms meeting sample selection criteria (four quarterly earnings announcements per year, available stock returns) are retained.

4.	Permanent identifiers (`PERMNO`, `PERMCO`, `GVKEY`) ensure accurate historical tracking [@gow_empirical_2025]. `PERMNO` is used instead of `PERMCO` as it uniquely identifies individual securities rather than companies, preventing aggregation issues from multiple share classes.

5.	The replication relies on Compustat and CRSP, linked via the CRSP/Compustat Merged (CCM) database. As Compustat updates daily, while CRSP and CCM update annually, timing mismatches may arise. Unlike the original study, this replication uses the latest WRDS version as of January 2025, meaning database updates may introduce firm coverage or value differences.

6. Since @ball_how_2008 analyze returns, I use `RET` variable in CRSP too to simplify the process. Bid and ask prices are ignored.

7. Same as @ball_how_2008, I group earnings announcements based on their actual announcement dates (`RDQ` from Compustat) rather than the fiscal periods they cover. Since Compustat provides the precise earnings announcement date and CRSP aligns stock returns with trading days (`date` in CRSP), this approach ensures consistency. So, no adjustments are made to realign quarters.

By following the steps outlined in @sec-replication_steps1 and adhering to the assumptions made, I aim to replicate the results from @ball_how_2008. A thorough step-by-step approach, with each step clearly documented, helps to understand and verify the outputs.

## Replication Steps {#sec-replication_steps1}
The analysis involves gathering and filtering stock return and earnings announcement data, linking CRSP stock data with Compustat firm-level accounting data fundamentals, and estimating the relationship between annual returns and earnings-announcement window returns. The replication process follows a structured workflow of data pull, preparation, and analysis, ensuring consistency with the original study.

**Step 1: Pulling the Data and Managing the Databases**  
Data is pulled from WRDS using Python to integrate daily stock returns from CRSP (`crsp_a_stock.dsf`) with quarterly earnings announcement data from Compustat  
(`comp_na_daily_all.fundq`), linked via the Compustat-CRSP link table  
(`crsp_a_ccm.ccmxpf_lnkhist`). `permno` unique identifier matches stock returns to firm fundamentals via `gvkey` for exact firm/year match. Key variables include `ret` (daily stock return) and `rdq` (earnings announcement date), enabling event study analysis from 1972 to 2023. While WRDS provides extensive financial datasets, documentation on precise variable selection and dataset integration is limited, requiring additional validation and cross-referencing with prior research. The data is pulled for the entire range 1972-2023 required for Task 1 and 2 to prevent double preparation step.

**Step 2: Data Preparation**  
Although not finished, the plan would be as follows: The merged dataset would integrate stock return data from CRSP (`crsp_daily_stock_returns`), firm fundamentals from Compustat (`compustat_fundq_1972_2023`), and the CRSP-Compustat linking table (`linkdata_compustat_crsp.parquet`). The `PERMNO` identifier from CRSP and `GVKEY` from Compustat are connected via `LPERMNO` in the linking table, ensuring correct firm-year alignment. Unnecessary variables from linked table, such as `linkprim`, `liid`, `linktype`, and `usedflag`, are dropped. The resulting dataset would retain only daily stock returns (`ret`), earnings announcement dates (`rdq`), fiscal year-end indicators (`fyr`), and firm identifiers (`permno, gvkey`), used further for Buy-and-Hold Returns (BHR) calculation and regression analysis.

After defining the event windows for each earnings announcement day in Q1-Q4, retain only the firms that have four earnings announcements per year, such that each of them falls within the same calendar year. Additionally, for the subgroup replication of Table 2 Panels C-D classify the firms based on their fiscal year-end (`fyr`) to define whether firms have December fiscal year-end firm or in any another month. Control for the duplicate and missing firm/years.

**Step 3: Analysis Implementation and Reproduction of Tables and Figure**  
The analysis step computes the BHR returns and performs a cross-sectional regression for each distinct firm/year.

Since @ball_how_2008 do not explicitly describe their return computation, I compute Buy-and-Hold Returns (BHR) following the cumulative return formula as discussed in @gundersen_returns_2022. The BHR formula captures compounded returns over multiple days, reflecting the stock’s actual performance, which is used to replicate Table 1.

The following formula is used to calculate calendar-year buy-and-hold returns, that measures the stock’s total performance over the year for each firm using daily CRSP returns:
$$
BHR_{\text{annual}} = \prod_{t=1}^{T} (1 + R_t) - 1, \quad \text{where } R_t \text{ is the daily return and } T \text{ is the total trading days in a year.}
$$

Specifically, buy-and-hold return over the three-day event window, that measures the stock’s reaction only to earnings news, is computed as given below:
$$
\begin{aligned}
BHR_{\text{event}} &= (1 + R_{t_1}) (1 + R_{t_2}) \dots (1 + R_{t_T}) - 1
\\
\text{where } R_t & \text{ is the daily return and } T \text{ is the total trading days in a year.}
\end{aligned}
$$

Next, perform cross-sectional regression for each firm/year - the main objective of @ball_how_2008's research:
$$
R_i(\text{annual}) = a_0 + a_1 R_i(\text{window1}) + a_2 R_i(\text{window2}) + a_3 R_i(\text{window3}) + a_4 R_i(\text{window4}) + \varepsilon_i
$$

The DV (BHR_annual) is the firm’s calendar-year buy-and-hold return from CRSP, aggregated over all trading days, while the IVs (BHR_event) are the four quarterly earnings-announcement returns, computed as three-day buy-and-hold returns (-1, 0, +1) around each earnings announcement date (RDQ) from Compustat. Additionally, I compute abnormal R^2 by benchmarking the adjusted  R^2  from annual regressions against the 4.8% baseline. This lays the ground for Table 2 replication.

## Results {#sec-results1}
Since Tables and Figure replication not finished, this section presents only the original @fig-fig1_original.

```{python}
#| label: fig-fig1_original
#| fig-cap: "Original Figure 1 from Ball and Shivakumar (2008)"
#| echo: false

import matplotlib.pyplot as plt
import matplotlib.image as mpimg

img = mpimg.imread("fig1_original.png")
plt.figure(figsize=(12, 20))
plt.imshow(img)
plt.axis("off")
plt.show()

# used in AssI Solution, I use knitr for availability of cross-referencing.
#| label: fig1_original_png
#| fig-cap: "Original Figure 1 from Ball and Shivakumar (2008)"
#| echo: false

#from IPython.display import Image

#Image(filename="fig1_original.png")
#
```

# Task 2 - Extending the Analysis to 2007–2023 {#sec-task2}
This Section extends the analysis to 2007–2023, a period defined by major economic events such as the 2008 financial crisis and the COVID-19 pandemic, which may have influenced the informativeness of earnings announcements. While the task is relatively straightforward in terms of programming—requiring adjustments to the sample period—it demands careful economic interpretation of observed trends, market shifts, and potential institutional details that may influence the relationship between earnings announcements and stock returns.

## Research Design Choices and Assumptions {#sec-research_design_assumptions2}

Assumptions from @sec-research_design_assumptions1 still hold.

## Replication Steps {#sec-replication_steps2}
The pull, preparation, and analysis step would remain the same as designed in @sec-replication_steps1 with the only change in data range.

## Results {#sec-results2}
The sample shift to 2007–2023 is expected to bring potential shifts in earnings announcement informativeness due to significant economic and regulatory events. The 2008 Global Financial Crisis likely disrupted earnings predictability, with heightened volatility and uncertainty increasing the market’s sensitivity to earnings reports. The COVID-19 pandemic 2020 introduced extreme market uncertainty, potentially amplifying the informativeness of earnings announcements as investors relied on them to gauge firm resilience and recovery. Additionally, the introduction of Regulation (EU) No 537/2014 may have influenced earnings informativeness by improving transparency, audit quality, and financial reporting compliance across EU countries. Differences in compliance and audit coverage could have led to variations in how earnings announcements reflect new information in stock prices, potentially contributing to cross-country disparities in earnings informativeness.

# Task 3 - Cross-Country Replication {#sec-task3}
Task 3 extends the analysis beyond the U.S. by replicating Figure 1 using 1972–2023 data from a non-U.S. market, examining how earnings informativeness varies across regulatory environments and market structures. This requires mapping CRSP/Compustat variables to their Worldscope/Datastream equivalents while adjusting for accounting standards, market liquidity, and institutional factors. Worldscope provides earnings announcement data, and Datastream covers stock returns. While the core methodology remains, careful data handling and interpretation are essential for cross-country analysis, particularly with Canada.

## Research Design Choices and Assumptions {#sec-research_design_assumptions3}
Since this task employs a generalization and extension approach by applying the same methodology to a country different from original report, it requires adjustments for different databases and market structures. Based on the original paper, I do not gather analyst expectations or estimate earnings deviations for my replication. Instead, I only require stock return data and earnings announcement dates.

Canada was chosen for its quarterly reporting frequency, ensuring comparability with U.S. data, as most firms report four times per year, with some semi-annual reports [@short_short_2025]. Despite its smaller market size, Canada’s regulatory environment, liquidity, and institutional settings closely align with the U.S., making it an ideal candidate. Additionally, Canadian firms registered with the SEC can report under U.S. GAAP, making their financial disclosures fully compatible with U.S. firms.

In addition, I impose the following assumptions apart from those listed in @sec-research_design_assumptions1 to ensure generalization and extension approach due to usage of other databases or where @ball_how_2008 do not provide explicit guidance:

1. To ensure a smooth and conflict-free process, I handle the WSCP/Datastream workflow separately, preventing potential interference with the CRSP/Compustat pull process and improving debugging efficiency. Although a unified pull/prepare/do_analysis file for all databases would be ideal and streamline management, the large-scale nature of the extracted datasets demands modular approach.

2. Since I do not perform subgroup analysis for December and non-December fiscal year end firms as in Table 2 Panels C-D, `item5350` for FISCAL PERIOD END DATE is not relevant for this task, as Figure 1 applies to the analysis of the entire dataset. So, the sample is all firm-years with available data on the quarterly Worlscope and daily Datastream files.

3. I verified whether Worldscope and Datastream follow the same earnings announcement date convention as Compustat or if they report earnings based on fiscal periods. I decided to use Worldscope items 5901–5904 from historical stock-related data in `wrds_ws_stock`, which represent the actual date when a company publicly reports its earnings for each quarter, where, however, data for these fields is generally not available prior to 1992 [@thomson_financial_worldscope_2007, p.281]. This may cause limited data sample size.

4. This task is a generalization to a non-U.S. setting, meaning methodological differences will certainly arise due to variations in data structure and reporting standards. Worldscope - Stock Data `wrds_ws_stock` table is updated quarterly, while Datastream Daily Stock File `wrds_ds2dsf` is updated weekly, which may introduce timing mismatches when aligning financial fundamentals with stock returns. Given that, future applications should consider database updates and coverage differences between CRSP and Datastream.

5. Similar to @sec-task1, I use returns `ret` variable in Datastream for returns, presented as percentage change. Bid and ask prices are ignored.

6. I ignore the currency effect (CAD or USD) in DS table, as returns are measured in percentages, making them relative rather than absolute values.

## Replication Steps {#sec-replication_steps3}
The process is same as in @sec-task1.

**Step 1: Pulling the Data and Managing the Databases**  
I follow the linking approach between Worldscope and Datastream based on @dai_research_2021. The original study uses data from 1972 onward, while Worldscope's `year_` variable only provides data starting from 1980 [@wharton_research_data_services_worldscope_2025]. This constraint may lead to differences in sample coverage. Hence, filter Datastream `marketdate` to 1980 to prevent mismatched observations where stock data exists but earnings do not.

Next, retrieve quarterly earnings announcement dates from Worldscope and daily stock returns from Datastream, filtering for Canadian firms from 1980 onward. The datasets are saved separately and later merged using firm-level identifiers. Only common equity (`typecode` = EQ) is included to ensure consistency with CRSP.

The pull step results into three comprehensive datasets: DS table with 27,950,012, Link table with 77,944, WSCP table with 1,942,202 observations.

**Step 2: Data Preparation**  
First, I merged Worldscope stock data with the linking table using `code` as unique identifier, resulting in 747,760 observations. This merged dataset was then joined with Datastream daily stock data using unique `infocode`, producing a final dataset with 10,249,737 observations.

If a firm had missing values for any of the Earnings Per Share Report Date fields (item5901–item5904), it did not meet the criteria set by @ball_how_2008 and was excluded from the sample. This filtering step removed 3,609,492 missing firm-year observations. Additionally, 5,996,793 firm-year entries were dropped because their four earnings announcements did not fall within the same calendar year. Furthermore, I extend dynamically shift the missing values for trading days -1,0,1 within 3 day window, due to small sample size to keep more earning announcemnet windows.

From this cleaned dataset, BHR event returns are computed. Then, annual stock returns for the firms of the filtered sample are extracted from Datastream pulled data . Finally, the Annual BHR Returns are calculated. Notably, after pulling the annual stock data for further annual return calculation, all infocodes present an average of 250 trading days, highlighting full year stock coverage.

**Step 3: Analysis Implementation and Reproduction of Tables and Figure**  
I conduct the earnings informativeness analysis by computing summary statistics (Table 1), running annual regressions (Table 2), and generating a replication of Figure 1. First, I loaded the computed event-window and annual buy-and-hold returns (BHR_3day, BHR_Annual) and produced summary statistics similar to @ball_how_2008. Next, I performed annual cross-sectional regressions of stock returns on earnings-announcement returns, benchmarking against an Abnormal R² (adjusted R² minus 4.8%), while handling missing data and ensuring a valid number of observations per year. Finally, the script plots Figure 1, visualizing Abnormal R² trends (Panel A) and slope coefficients (Panel B) over time, saving it as both a PNG and a pickle file for further use.

## Results {#sec-results3}
Comparing the replication results with those from @ball_how_2008 (see @fig-fig1_original) reveals both similarities and discrepancies. The original figure shows a general upward trend in abnormal  R² , peaking in the early 2000s, whereas the replicated figure captures greater volatility in the later years (2010–2023). Additionally, Ball’s figure suggests a smoother long-term dynamic, while the replication exhibits more pronounced fluctuations, particularly post-2010. Panel B in replicated figure exhibits higher volatility and larger coefficient magnitudes, suggesting increased earnings informativeness post-2006, while Ball’s original figure has more stable and lower-magnitude coefficients.

The post-2000 period in the replication shows more inconsistent abnormal  R^2  values, possibly reflecting evolving market structures, shifts in disclosure practices, and increased earnings informativeness following regulatory changes (e.g., IFRS adoption, post-SOX adjustments, different financial market structure). The pronounced fluctuations could also be driven by global financial crisis, Canada's higher market concentration, fewer publicly traded firms, and potential regulatory differences, which may affect earnings informativeness and stock return patterns. The variability in slope coefficients further suggests that earnings announcements’ impact on stock returns is less stable than in prior decades, potentially due to differences in investor response.

```{python}
#| label: fig-abnormal-r2
#| fig-cap: "Replicated Figure 1: Abnormal Adjusted R² Values and Slope Coefficients from Annual Regressions"
#| fig-align: center
#| echo: false

import pickle
import matplotlib.pyplot as plt

# Load the figure from the pickle file
with open('../output/figure1_replication.pickle', 'rb') as f:
    fig = pickle.load(f)

# Render the figure explicitly
plt.figure(fig.number)
plt.show()
```

# Task 4 - Research Design for a Non-Archival Study {#sec-task4}

The research design follows the generalization goal in empirical literature, emphasizing the need to validate findings across different methods [@bloomfield_gathering_2016]. @ball_how_2008 assess earnings informativeness through stock price reactions, but triangulation—beyond replication—requires diverse methods to examine the same question [@bloomfield_gathering_2016, p.353]. While price reactions indicate market response, they do not capture investor perceptions. A survey directly evaluates whether market participants view earnings announcements as confirmatory or informative. If investors confirm a secondary role, it reinforces Ball and Shivakumar’s (2008) findings, ensuring informativeness reflects cognitive and behavioral mechanisms rather than price fluctuations. These insights could inform corporate reporting strategies by clarifying how investors prioritize disclosures.

A survey is the most suitable non-archival method, capturing investor sentiment beyond what archival data provides. Unlike lab experiments, which lack real-world applicability, surveys measure the dependent variable (DV) by eliciting participant perceptions [@bloomfield_gathering_2016, p.358]. Field experiments, though capturing real-time reactions, require direct intervention, posing ethical and logistical challenges. A long-term research program combining surveys and field studies would enhance validity, leveraging unlimited funds to assess investor decision-making and market responses.

To ensure survey reliability, design strategies follow @bruggemann_worku_2024. A sample of 300 investors, analysts, and financial professionals will provide institutional and retail perspectives while balancing statistical power and feasibility. The survey will be distributed via email through financial networks, investor associations, and professional platforms (e.g., LinkedIn, CFA societies) and in paper format at finance conferences. The Humboldt-Universität zu Berlin (HU) brand may be used, subject to approval. Respondents will be assured confidentiality and offered an €20 honorarium. A pilot study with 30 participants will refine question clarity, survey length, and response distribution, helping identify unclear or missing questions and assess completion time.

To minimize response bias, questions will be neutrally worded to capture pre- and post-announcement perceptions. The fully anonymous survey ensures candid responses without requiring personal financial history.

The **DV** is the perceived informativeness of earnings announcements, measured through survey responses on whether they are confirmatory or informative. Independent variables (**IVs**) include investor preparation, market awareness, reaction timing, perceived novelty of earnings, key financial indicators, and external conditions. These are measured via Likert-scale, ranking, and multiple-choice questions, supplemented by qualitative insights from follow-up interviews.

To deepen findings, selected participants will be invited for confidential follow-up interviews, ensuring diversity in investment strategies, industries, and expertise. These interviews will be conducted via video conferencing or phone calls. The following sections provide detailed sketches of the survey and interview designs.

## Survey Questions

1. **How do you typically prepare for earnings announcements in your investment decisions?**
    * I conduct in-depth research and adjust my positions in advance 
    * I monitor but rarely adjust positions pre-announcement
    * I rely on market consensus, analyst forecasts, and AI-driven insights
    * I do not make investment decisions based on earnings announcements

2. **How do you typically react to earnings announcements?**
    * I adjust my investment strategy immediately based on the announcement.
    * I wait for further analysis before making changes.
    * I rarely make investment decisions based on earnings announcements.

3. **To what extent do you believe that earnings announcements provide new information beyond what is already reflected in market prices?**
    * Always
    * Most of the time
    * Sometimes
    * Rarely
    * Never

4. **Rank the following factors in order of importance when evaluating earnings announcements (1 = most important, 5 = least important):**
    * Earnings per share (EPS) compared to analyst forecasts
    * Revenue growth
    * Management guidance and commentary
    * Market reaction on the day of the announcement
    * Industry trends and macroeconomic conditions

5. **Do you consider earnings announcements to be more confirmatory or informative?**
    * Primarily confirmatory (reinforce existing expectations)
    * Primarily informative (provide new insights)
    * A mix of both

## Follow-Up Interview Questions

1. Can you describe a recent instance where an earnings announcement significantly influenced your investment decision?
2. In your experience, are there specific industries where earnings announcements are more informative than confirmatory?
3. Do you use earnings announcements differently depending on market conditions (e.g., economic downturn vs. growth periods)?

# Conclusion {#sec-conclusion}
This project effectively demonstrates the use of a systematic and collaborative workflow for empirical accounting research, leveraging the TRR 266 Template for Reproducible Empirical Accounting Research. By following an open science approach, I aimed to replicate key tables and figures from @ball_how_2008, providing insights into how earnings announcements contribute to price formation and market efficiency. By extending the analysis beyond the original study’s U.S. market focus, the project highlights how market structures and investor behaviors influence the informativeness of earnings disclosures in different jurisdictions.

The study contributes to the ongoing discussion on market efficiency and the economic role of corporate disclosures by providing empirical evidence on how earnings informativeness has evolved over time and across jurisdictions. The cross-country analysis with Canada highlights variations in market responses to earnings announcements, suggesting that institutional factors, market structures, and investor behavior play a role in shaping how financial information is processed. These findings have implications for investors, analysts, and policymakers, as they underscore the importance of regulatory consistency and market transparency in shaping the effectiveness of financial reporting.

This assignment required a comprehensive application of programming skills and institutional knowledge gained throughout the course, integrating data analysis, replication, and visualization techniques in line with open science principles. In the future, this repository can be cloned or forked (if made public) to facilitate further research on earnings informativeness, enabling additional extensions or robustness tests. Additionally, the survey developed as part of the non-archival study proposal can be further refined and expanded, providing a structured framework for gathering primary data on market participants’ interpretations of earnings announcements. Through this approach, the study not only revisits a fundamental question in financial research but also provides a foundation for future empirical investigations into the evolving role of earnings announcements in global capital markets, both through archival and non-archival methodologies. Thanks for reading!

\pagebreak

\setcounter{table}{0}
\renewcommand{\thetable}{\arabic{table}}

# References {-}
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\noindent

### 2025-05-03 doc/presentation.qmd
Initial commit of huwiwiQuartoThesis based on CQA template

--- ## Not rendered in this project; to be used as draft
title: |
  | Earnings Management and Investor Protection:
  | Accounting Reading Group - Assignment III
author: "Mel Mzv"
institute: "Humboldt-Universität zu Berlin – Summer 2024"
date: today
fontsize: "9pt"
pdf-engine: xelatex # pdflatex creates rastered fonts
format: 
  beamer:
    slide-level: 3
    number-sections: true
    toc: false
header-includes:
- \usepackage{booktabs} 
- \usepackage{threeparttable}
- \usepackage{graphicx}
- \usepackage{etoolbox}
- \AtBeginEnvironment{verbatim}{\fontsize{7}{7}\selectfont}  # This sets the font size for verbatim environments
- \input{beamer_theme_trr266.sty}
bibliography: references.bib
biblio-style: apsr
---

```{python}
#| echo: false
#| output: false

import pickle
import pandas as pd

# Load the results from the existing pickle file for table_1
with open('../data/generated/table_1.pickle', 'rb') as f:
    results = pickle.load(f)

# Load the results from the new pickle file for final_table and summary_stats
with open('../output/em_results.pickle', 'rb') as f:
    em_results = pickle.load(f)

def escape_special_characters(df):
    df.columns = [col.replace('_', '\\_').replace('#', '\\#') for col in df.columns]
    return df

# Custom formatter function
def custom_float_format(x):
    return f"{x:.3f}".lstrip('0') if pd.notnull(x) else ""

def prep_latex_table(df, caption=None, label=None, fontsize='\\fontsize{11}{13}\\selectfont', description=None):
    df = escape_special_characters(df.reset_index(drop=True))  # Drop the index column from output tables for better readability
    num_columns = len(df.columns)
    column_format = 'l' + 'r' * (num_columns - 1)

    latex_table = df.to_latex(
        column_format=column_format,
        index=False, 
        float_format=custom_float_format,
        escape=False  # Important to disable escaping so LaTeX commands in column names are respected
    )

    # Construct description to the table
    full_caption = f"{caption}\\\\\n\\textnormal{{{description}}}" if description else caption

    latex_table_lines = [
        "\\begin{table}[htbp]",
        "\\centering",
        fontsize,  # Set font size for the table
        latex_table,
        f"\\caption{{{full_caption}}}" if full_caption else "",
        f"\\label{{{label}}}" if label else "",
        "\\end{table}"
    ]
    return "\n".join(line for line in latex_table_lines if line)
```

# Motivation

## Understanding Systematic Differences in Earnings Management across Countries

- The paper of @Leuz_2003 focuses on the relationship between investor protection and earnings management.
- Investor protection plays a crucial role in limiting the extent of earnings management by reducing the opportunities for managers to manipulate earnings.
- Analyzing the relationship between EM and investor protection across different countries provides insights into how legal and institutional frameworks impact financial reporting practices.
- Earnings management (EM) measures reflect the extent to which managers manipulate financial reports to meet certain objectives, such as meeting earnings targets or influencing stock prices.

## Project Objective

- The primary objective is to demonstrate a reproducible and collaborative research workflow by calculating and analyzing EM measures across different countries.
- The study aims to replicate and extend the findings of @Leuz_2003, exploring the relationship between EM and investor protection using Worldscope datasets for the time range 1990 to 1999.
- The analysis provides insights into the levels of earnings management across countries, comparing them with the original findings of Leuz, Nanda, and Wysocki (2003).

## Project Relevance

- Understanding EM trends across countries aids in assessing global financial transparency.
- Promotes transparency and reproducibility in empirical accounting research.

# Research Design

- Data sourced from the Worldscope Database (1990-1999) July 2024 version.
- Replication includes calculation of **four earnings management measures**:
  1. Earnings smoothing using accruals (EM1)
  2. Correlation between changes in accruals and cash flows (EM2)
  3. Magnitude of accruals (EM3)
  4. Ratio of small profits to small losses (EM4)

# Assumptions

- **Database Version**: Analysis uses the July 2024 Worldscope data, which may differ from the November 2000 version due to updates and restatements.
- **Key Terms**: Earnings management (EM) and investor protection defined as per @Leuz_2003 for consistency.
- **Handling Negative Values**: Negative values in key metrics (e.g., operating income) were included to capture the full scope of EM activities.
- **Variable Selection**: Chose specific Worldscope variables (e.g., item1151 over item4051) based on best practices; differences may affect results.
- **Currency Consistency**: EM measures are based on scaled variables; currency consistency ensures comparability across countries.

# Data - Replication Steps

Step 1: Pulling the Data and Managing the Databases

Step 2: Data Preparation

+  To verify the robustness of descriptive statistics before the main replication, Table 1 from Leuz et al. was replicated, yielding results that closely align with the original study:
---
```{python}
#| label: table_1 
#| echo: false
#| output: asis

print(prep_latex_table(
    results['table_1'],
    caption='Replicated table - Descriptive statistics: Number of firm-year observations per country',
    label='tab:table_1',
    fontsize='\\fontsize{5pt}{6pt}\\selectfont'  # Set smaller font size
))
```
---

Step 3: Analysis Implementation and Table Reproduction

- Smoothing Reported Operating Earnings Using Accruals (EM1)
    $$
    \text{EM1} = \text{Median}\left(\frac{\frac{\sigma(\text{Operating Income}_{it})}{\text{Total Assets}_{i(t-1)}}}{\frac{\sigma(\text{Operating Cash Flow}_{it})}{\text{Total Assets}_{i(t-1)}}}\right)
    $$
- Smoothing and the Correlation Between Changes in Accounting Accruals and Operating Cash Flows (EM2)
    $$
    \text{EM2} = \rho\left(\frac{\Delta \text{Accruals}_{it}}{\text{Total Assets}_{i(t-1)}}, \frac{\Delta \text{CFO}_{it}}{\text{Total Assets}_{i(t-1)}}\right)
    $$
- Discretion in Reported Earnings: The Magnitude of Accruals (EM3)
    $$
    \text{EM3} = \text{Median}\left(\frac{|\text{Accruals}_{it}|}{|\text{CFO}_{it}|}\right)
    $$

 ---
- Discretion in Reported Earnings: Small Loss Avoidance (EM4)
    $$
    \text{EM4} = \frac{\text{Small Profits}}{\text{Small Losses}}
    $$
 - Aggregate Measure of Earnings Management
    $$
    \text{Aggregate EM Score}_{\text{country}} = \frac{\text{Rank}_{\text{EM1}} + \text{Rank}_{\text{EM2}} + \text{Rank}_{\text{EM3}} + \text{Rank}_{\text{EM4}}}{4}
    $$

# Results

```{python}
#| label: final_combined_table
#| echo: false
#| output: asis

# Rename the columns to ensure consistency with paper
em_results['final_combined_table'].rename(columns={
    'item6026': 'Country',
    'EM1': 'EM1 (−)',
    'EM2': 'EM2 (−)',
    'EM3': 'EM3 (+)',
    'EM4': 'EM4 (+)',
    'Aggregate_EM_Score': 'Aggregate EM score'
}, inplace=True)

print(prep_latex_table(
    em_results['final_combined_table'], 
    caption='Replicated table - Country scores for earnings management measures (Sorted by aggregate earnings management)',
    label='tab:final_combined_table',
    fontsize='\\fontsize{4pt}{5pt}\\selectfont',  # Set smaller font size

))
```

# Conclusion

## Summary of Findings
- EM1 Alignment: Strong alignment with minor differences, reflecting consistent trends in earnings management across countries.
- EM2 Findings: Similar patterns observed; Greece and Japan show more smoothing, while Canada and the U.S. show less.
- EM3 Consistency: Replicated results closely match the original, confirming trends in accrual magnitudes.
- EM4 Regional Differences: Significant regional variations in loss avoidance, consistent with original findings.
- Aggregate Scores: Strong alignment between original and replicated scores, confirming key trends across countries.
- **Overall Conclusion**: Replication confirms higher EM in European and Asian countries compared to Anglo-American counterparts.

## Key Insights

- Significant differences in earnings management practices across countries.
- Strong alignment with original findings in Leuz, Nanda, and Wysocki (2003).
- The project structure demonstrated the effectiveness of an open science and collaborative workflow.
- **Future Work**: Repository can be cloned or forked for further analysis.

# References {-}
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\noindent

### 2025-05-03 code/python/do_analysis-wscp.py

@@ -1,0 +1,232 @@
+# We start by loading the libraries that we will use in this analysis.
+import os
+import pickle
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+import statsmodels.api as sm
+from utils import read_config, setup_logging
+
+# Set up logging
+log = setup_logging()
+
+def main():
+    log.info("Starting analysis ...")
+    cfg = read_config('config/do_analysis_cfg.yaml')
+
+    # Load datasets
+    log.info("Loading computed BHR Event and BHR Annual datasets...")
+    bhr_event_results = pd.read_csv(cfg["bhr_event_output_csv"])
+    bhr_annual_results = pd.read_csv(cfg["bhr_annual_output_csv"])
+
+    # Compute summary statistics
+    df_summary = compute_summary_statistics(bhr_annual_results, bhr_event_results)
+
+    # Save summary statistics
+    summary_statistics_csv = cfg["summary_statistics_csv"]
+    df_summary.to_csv(summary_statistics_csv, index=False)
+    log.info(f"Summary statistics saved to {summary_statistics_csv}")
+
+    # Run annual regressions
+    df_regression = run_regressions(bhr_annual_results, bhr_event_results)
+
+    # Save regression results (CSV)
+    regression_results_csv = cfg["regression_results_csv"]
+    df_regression.to_csv(regression_results_csv, index=False)
+    log.info(f"Regression results saved to {regression_results_csv}")
+
+    # Generate Figure 1 replication
+    plot_figure1(df_regression, cfg["figure1_save_path"], cfg['figure1_pickle_path'])
+
+    
+    log.info("Analysis complete.")
+
+def compute_summary_statistics(bhr_annual_results, bhr_event_results):
+    """
+    Computes summary statistics (Mean, Median, Skewness, % Obs = 0, % Obs > 0) 
+    for annual returns and earnings-announcement window returns like in Table 1 by Ball(2008)
+    to compare preliminary results.
+    """
+    log.info("Computing summary statistics for comparison with Table 1...")
+
+    # **Annual Returns Summary**
+    annual_stats = {
+        "Category": "Calendar-Year Returns",
+        "No. Obs.": len(bhr_annual_results),
+        "Mean": np.mean(bhr_annual_results["BHR_Annual"]),
[…]

### 2025-05-03 code/python/prepare_data-wscp.py

@@ -1,0 +1,449 @@
+# --- Header -------------------------------------------------------------------
+# Prepare the pulled data for further analysis as per Task 3 requirements
+#
+# (C) Mel Mzv - See LICENSE file for details
+# ------------------------------------------------------------------------------
+
+import pandas as pd
+from utils import read_config, setup_logging
+
+log = setup_logging()
+
+def main():
+    log.info("Preparing data for analysis ...")
+    cfg = read_config('config/prepare_data_cfg.yaml')
+
+    # Load the pulled datasets
+    ws_stock = pd.read_csv(cfg['worldscope_sample_save_path_csv'])
+    link_ds_ws = pd.read_csv(cfg['link_ds_ws_save_path_csv'])
+    ds2dsf = pd.read_csv(cfg['datastream_sample_save_path_csv'])
+
+    # Step 1: Merge Worldscope with the Linking Table
+    log.info("Merging Worldscope with Linking Table...")
+    ws_link_merged = merge_worldscope_link(ws_stock, link_ds_ws)
+
+    # Step 2: Pivot dataset to long format
+    log.info("Pivoting merged dataset to long format...")
+    ws_long = pivot_longer_earnings(ws_link_merged)
+
+    # Step 3: Expand dataset for event windows (-1, 0, +1 days)
+    log.info("Expanding dataset for event windows...")
+    ws_expanded = expand_event_window(ws_long)
+
+    # Step 4: Merge with Datastream stock returns
+    log.info("Merging expanded dataset with Datastream stock returns...")
+    merged_dataset = merge_with_datastream(ws_expanded, ds2dsf)
+
+    # Step 5: Select firms that meet sample criteria
+    log.info("Selecting firms that meet the sample criteria (4 announcements per year)...")
+    final_dataset = select_firms_for_sample(merged_dataset)
+
+    # Step 6: Compute and Save BHR (Event Window)
+    bhr_event_results = compute_and_save_eawr_bhr(final_dataset, cfg)
+
+    # Step 7: Extract annual stock return data for firms in BHR Event dataset
+    annual_stock_data = extract_annual_stock_data(bhr_event_results, ds2dsf, cfg)
+
+    # Step 8: Compute and Save BHR (Annual Return)
+    bhr_annual_results = compute_and_save_annual_bhr(cfg)
+
+    # Step 9: Save the final dataset (full dataset with event windows)
+    final_dataset.to_csv(cfg['prepared_wrds_ds2dsf_path'], index=False)
[…]

### 2025-05-03 code/python/pull_wrds_data-wscp.py

@@ -1,0 +1,118 @@
+# --- Header -------------------------------------------------------------------
+# See LICENSE file for details
+#
+# This code pulls data from WRDS Worldscope and Datastream Databases and links them
+# ------------------------------------------------------------------------------
+import os
+from getpass import getpass
+import dotenv
+import pandas as pd
+from utils import read_config, setup_logging
+import wrds
+
+log = setup_logging()
+
+def main():
+    """
+    Main function to pull data from WRDS.
+
+    This function reads the configuration file, gets the WRDS login credentials, 
+    and pulls the data from WRDS.
+
+    The data is then saved to CSV and Parquet files.
+    """
+    cfg = read_config('config/pull_data_cfg.yaml')
+    wrds_login = get_wrds_login()
+    
+    # Pull data from WRDS
+    worldscope_df, ds_df, linkdata_df = pull_wrds_data(cfg, wrds_login)
+
+    # Save pulled data
+    worldscope_df.to_csv(cfg['worldscope_sample_save_path_csv'], index=False)
+    worldscope_df.to_parquet(cfg['worldscope_sample_save_path'], index=False)
+
+    ds_df.to_csv(cfg['datastream_sample_save_path_csv'], index=False)
+    ds_df.to_parquet(cfg['datastream_sample_save_path'], index=False)
+
+    linkdata_df.to_csv(cfg['link_ds_ws_save_path_csv'], index=False)
+    linkdata_df.to_parquet(cfg['link_ds_ws_save_path'], index=False)
+
+
+def get_wrds_login():
+    """
+    Gets the WRDS login credentials.
+    """
+    if os.path.exists('secrets.env'):
+        dotenv.load_dotenv('secrets.env')
+        wrds_username = os.getenv('WRDS_USERNAME')
+        wrds_password = os.getenv('WRDS_PASSWORD')
+    else:
+        wrds_username = input("Please provide a WRDS username: ")
+        wrds_password = getpass("Please provide a WRDS password (it will not show as you type): ")
+
+    return {'wrds_username': wrds_username, 'wrds_password': wrds_password}
+
+
+def pull_wrds_data(cfg, wrds_authentication):
+    """
+    Pulls data from WRDS (Worldscope, Datastream, and linking table).
+    """
+    db = wrds.Connection(
[…]

### 2025-05-03 code/python/pull_wrds_data.py

@@ -1,0 +1,101 @@
+# --- Header -------------------------------------------------------------------
+# See LICENSE file for details
+#
+# This code pulls data from WRDS Databases 
+# ------------------------------------------------------------------------------
+
+import os
+from getpass import getpass
+import dotenv
+
+import pandas as pd
+from utils import read_config, setup_logging
+import wrds
+
+log = setup_logging()
+
+def main():
+    '''
+    Main function to pull data from WRDS.
+
+    This function reads the configuration file, gets the WRDS login credentials, and pulls the data from WRDS.
+
+    The data is then saved to CSV and Parquet files.
+    '''
+    cfg = read_config('config/pull_data_cfg.yaml')
+    wrds_login = get_wrds_login()
+    
+    # Pull CRSP and Compustat Data
+    db = wrds.Connection(
+        wrds_username=wrds_login['wrds_username'], 
+        wrds_password=wrds_login['wrds_password']
+    )
+    
+    log.info('Logged on to WRDS ...')
+    
+    pull_crsp_data(cfg, db)
+    pull_compustat_data(cfg, db)
+    pull_link_data(db)
+    
+    db.close()
+    log.info("Disconnected from WRDS")
+
+def get_wrds_login():
+    '''
+    Gets the WRDS login credentials.
+    '''
+    if os.path.exists('secrets.env'):
+        dotenv.load_dotenv('secrets.env')
+        wrds_username = os.getenv('WRDS_USERNAME')
+        wrds_password = os.getenv('WRDS_PASSWORD')
+        return {'wrds_username': wrds_username, 'wrds_password': wrds_password}
+    else:
+        wrds_username = input('Please provide a WRDS username: ')
+        wrds_password = getpass(
+            'Please provide a WRDS password (it will not show as you type): ')
+        return {'wrds_username': wrds_username, 'wrds_password': wrds_password}
+
+def pull_link_data(db):
+    """
+    Pulls the linking table between Compustat and CRSP from WRDS.
+    """
+    linkdata_df_wrds = db.get_table(library="crsp_a_ccm", table="ccmxpf_linktable")
+    linkdata_df_wrds.to_parquet("data/pulled/linkdata_compustat_crsp.parquet")
[…]

### 2025-05-03 doc/paper.qmd
clean up

@@
 Additionally, it sketches a research design for a non-archival study to evaluate the paper’s findings.
-This code base, adapted from TREAT, demonstrates how the template applies to the project and serves as a structured guide for reproducible empirical research in accounting.
+This code base, adapted from TREAT, demonstrates how the template applies to this project and serves as a structured guide for reproducible empirical research in accounting.
   <!--
@@
 
-The following formula is used to calculate calendar-year buy-and-hold returns, that measures the stock’s total performance over the year for each firm using daily CRSP returns:
-$$
-BHR_{\text{annual}} = \prod_{t=1}^{T} (1 + R_t) - 1, \quad \text{where } R_t \text{ is the daily return and } T \text{ is the total trading days in a year.}
-$$
-
-Specifically, buy-and-hold return over the three-day event window, that measures the stock’s reaction only to earnings news, is computed as given below:
-$$
-\begin{aligned}
-BHR_{\text{event}} &= (1 + R_{t_1}) (1 + R_{t_2}) \dots (1 + R_{t_T}) - 1
-\\
-\text{where } R_t & \text{ is the daily return and } T \text{ is the total trading days in a year.}
-\end{aligned}
-$$
-
-Next, perform cross-sectional regression for each firm/year - the main objective of @ball_how_2008's research:
-$$
-R_i(\text{annual}) = a_0 + a_1 R_i(\text{window1}) + a_2 R_i(\text{window2}) + a_3 R_i(\text{window3}) + a_4 R_i(\text{window4}) + \varepsilon_i
-$$
-
-The DV (BHR_annual) is the firm’s calendar-year buy-and-hold return from CRSP, aggregated over all trading days, while the IVs (BHR_event) are the four quarterly earnings-announcement returns, computed as three-day buy-and-hold returns (-1, 0, +1) around each earnings announcement date (RDQ) from Compustat.
-Additionally, I compute abnormal R^2 by benchmarking the adjusted  R^2  from annual regressions against the 4.8% baseline.
-This lays the ground for Table 2 replication.
-
 ## Results {#sec-results1}
 Since Tables and Figure replication not finished, this section presents only the original @fig-fig1_original.
-
-```{python}
-#| label: fig-fig1_original
-#| fig-cap: "Original Figure 1 from Ball and Shivakumar (2008)"
-#| echo: false
-
-import matplotlib.pyplot as plt
-import matplotlib.image as mpimg
-
-img = mpimg.imread("fig1_original.png")
-plt.figure(figsize=(12, 20))
-plt.imshow(img)
-plt.axis("off")
-plt.show()
-
-# used in AssI Solution, I use knitr for availability of cross-referencing.
-#| label: fig1_original_png
-#| fig-cap: "Original Figure 1 from Ball and Shivakumar (2008)"
-#| echo: false
-
-#from IPython.display import Image
-
-#Image(filename="fig1_original.png")
-#
-```
 
@@
 
-```{python}
-#| label: fig-abnormal-r2
-#| fig-cap: "Replicated Figure 1: Abnormal Adjusted R² Values and Slope Coefficients from Annual Regressions"
-#| fig-align: center
-#| echo: false
-
-import pickle
-import matplotlib.pyplot as plt
-
-# Load the figure from the pickle file
-with open('../output/figure1_replication.pickle', 'rb') as f:
-    fig = pickle.load(f)
-
-# Render the figure explicitly
-plt.figure(fig.number)
-plt.show()
-```
-
 # Task 4 - Research Design for a Non-Archival Study {#sec-task4}

### 2025-05-03 code/python/do_analysis-wscp.py
Neutralize Project Elements

@@ -1,232 +1,0 @@
-# We start by loading the libraries that we will use in this analysis.
-import os
-import pickle
-import pandas as pd
-import numpy as np
-import matplotlib.pyplot as plt
-import statsmodels.api as sm
-from utils import read_config, setup_logging
-
-# Set up logging
-log = setup_logging()
-
-def main():
-    log.info("Starting analysis ...")
-    cfg = read_config('config/do_analysis_cfg.yaml')
-
-    # Load datasets
-    log.info("Loading computed BHR Event and BHR Annual datasets...")
-    bhr_event_results = pd.read_csv(cfg["bhr_event_output_csv"])
-    bhr_annual_results = pd.read_csv(cfg["bhr_annual_output_csv"])
-
-    # Compute summary statistics
-    df_summary = compute_summary_statistics(bhr_annual_results, bhr_event_results)
-
-    # Save summary statistics
-    summary_statistics_csv = cfg["summary_statistics_csv"]
-    df_summary.to_csv(summary_statistics_csv, index=False)
-    log.info(f"Summary statistics saved to {summary_statistics_csv}")
-
-    # Run annual regressions
-    df_regression = run_regressions(bhr_annual_results, bhr_event_results)
-
-    # Save regression results (CSV)
-    regression_results_csv = cfg["regression_results_csv"]
-    df_regression.to_csv(regression_results_csv, index=False)
-    log.info(f"Regression results saved to {regression_results_csv}")
-
-    # Generate Figure 1 replication
-    plot_figure1(df_regression, cfg["figure1_save_path"], cfg['figure1_pickle_path'])
-
-    
-    log.info("Analysis complete.")
-
-def compute_summary_statistics(bhr_annual_results, bhr_event_results):
-    """
-    Computes summary statistics (Mean, Median, Skewness, % Obs = 0, % Obs > 0) 
-    for annual returns and earnings-announcement window returns like in Table 1 by Ball(2008)
-    to compare preliminary results.
-    """
-    log.info("Computing summary statistics for comparison with Table 1...")
-
-    # **Annual Returns Summary**
-    annual_stats = {
-        "Category": "Calendar-Year Returns",
-        "No. Obs.": len(bhr_annual_results),
-        "Mean": np.mean(bhr_annual_results["BHR_Annual"]),
[…]

### 2025-05-03 code/python/do_analysis.py

@@ -1,0 +1,98 @@
+# do_analysis.py — Template for analysis stage (after data prep)
+import os
+import pickle
+import pandas as pd
+import matplotlib.pyplot as plt
+import statsmodels.api as sm
+from utils import read_config, setup_logging
+
+log = setup_logging()
+
+
+def main():
+    log.info("Running post-preparation analysis...")
+
+    cfg = read_config("config/do_analysis_cfg.yaml")
+
+    # === Load prepared data ===
+    df_event = pd.read_csv(cfg["bhr_event_output_csv"])
+    df_annual = pd.read_csv(cfg["bhr_annual_output_csv"])
+
+    # === Summary Statistics ===
+    summary = compute_summary(df_annual, df_event)
+    summary.to_csv(cfg["summary_statistics_csv"], index=False)
+
+    # === Regressions & Plot ===
+    results = run_regressions(df_annual, df_event)
+    results.to_csv(cfg["regression_results_csv"], index=False)
+    plot_figure(results, cfg["figure1_save_path"], cfg["figure1_pickle_path"])
+
+    log.info("Analysis complete.")
+
+
+def compute_summary(df_annual, df_event):
+    """Basic summary stats for annual and event-window returns."""
+    def stats(series):  # Helper
+        return {
+            "Mean": series.mean(),
+            "Median": series.median(),
+            "Skew": series.skew(),
+            "% = 0": (series == 0).mean() * 100,
+            "% > 0": (series > 0).mean() * 100,
+            "Obs": len(series)
+        }
+
+    summary = [{"Category": "Annual", **stats(df_annual["BHR_Annual"])}]
+    for q in ["Q1", "Q2", "Q3", "Q4"]:
+        subset = df_event[df_event["quarter"] == q]["BHR_3day"]
+        summary.append({"Category": f"3-day {q}", **stats(subset)})
+
+    return pd.DataFrame(summary)
+
+
+def run_regressions(df_annual, df_event):
+    """Run annual regressions of annual returns on quarterly 3-day returns."""
+    df_event["year_stock"] = pd.to_datetime(df_event["rdq"]).dt.year
+    merged = df_annual.merge(df_event, on=["infocode", "year_stock"], how="inner")
+
+    results = []
+    for year in sorted(merged["year_stock"].unique()):
[…]

### 2025-05-03 code/python/prepare_data-wscp.py

@@ -1,449 +1,0 @@
-# --- Header -------------------------------------------------------------------
-# Prepare the pulled data for further analysis as per Task 3 requirements
-#
-# (C) Mel Mzv - See LICENSE file for details
-# ------------------------------------------------------------------------------
-
-import pandas as pd
-from utils import read_config, setup_logging
-
-log = setup_logging()
-
-def main():
-    log.info("Preparing data for analysis ...")
-    cfg = read_config('config/prepare_data_cfg.yaml')
-
-    # Load the pulled datasets
-    ws_stock = pd.read_csv(cfg['worldscope_sample_save_path_csv'])
-    link_ds_ws = pd.read_csv(cfg['link_ds_ws_save_path_csv'])
-    ds2dsf = pd.read_csv(cfg['datastream_sample_save_path_csv'])
-
-    # Step 1: Merge Worldscope with the Linking Table
-    log.info("Merging Worldscope with Linking Table...")
-    ws_link_merged = merge_worldscope_link(ws_stock, link_ds_ws)
-
-    # Step 2: Pivot dataset to long format
-    log.info("Pivoting merged dataset to long format...")
-    ws_long = pivot_longer_earnings(ws_link_merged)
-
-    # Step 3: Expand dataset for event windows (-1, 0, +1 days)
-    log.info("Expanding dataset for event windows...")
-    ws_expanded = expand_event_window(ws_long)
-
-    # Step 4: Merge with Datastream stock returns
-    log.info("Merging expanded dataset with Datastream stock returns...")
-    merged_dataset = merge_with_datastream(ws_expanded, ds2dsf)
-
-    # Step 5: Select firms that meet sample criteria
-    log.info("Selecting firms that meet the sample criteria (4 announcements per year)...")
-    final_dataset = select_firms_for_sample(merged_dataset)
-
-    # Step 6: Compute and Save BHR (Event Window)
-    bhr_event_results = compute_and_save_eawr_bhr(final_dataset, cfg)
-
-    # Step 7: Extract annual stock return data for firms in BHR Event dataset
-    annual_stock_data = extract_annual_stock_data(bhr_event_results, ds2dsf, cfg)
-
-    # Step 8: Compute and Save BHR (Annual Return)
-    bhr_annual_results = compute_and_save_annual_bhr(cfg)
-
-    # Step 9: Save the final dataset (full dataset with event windows)
-    final_dataset.to_csv(cfg['prepared_wrds_ds2dsf_path'], index=False)
[…]

### 2025-05-03 code/python/prepare_data.py

@@ -1,0 +1,112 @@
+# --- Header -------------------------------------------------------------------
+# Prepare the pulled data for further analysis as per Task 3 requirements
+#
+# (C) Mel Mzv - See LICENSE file for details
+# ------------------------------------------------------------------------------
+
+# --- Header -------------------------------------------------------------------
+# Prepare the pulled data for further analysis as per Task 3 requirements
+# ------------------------------------------------------------------------------
+
+import pandas as pd
+from utils import read_config, setup_logging
+
+log = setup_logging()
+
+def main():
+    log.info("Preparing data for analysis ...")
+    cfg = read_config('config/prepare_data_cfg.yaml')
+
+    ws_stock = pd.read_csv(cfg['worldscope_sample_save_path_csv'])
+    link_ds_ws = pd.read_csv(cfg['link_ds_ws_save_path_csv'])
+    ds2dsf = pd.read_csv(cfg['datastream_sample_save_path_csv'])
+
+    ws_link_merged = merge_worldscope_link(ws_stock, link_ds_ws)
+    ws_long = pivot_longer_earnings(ws_link_merged)
+    ws_expanded = expand_event_window(ws_long)
+    merged_dataset = merge_with_datastream(ws_expanded, ds2dsf)
+    final_dataset = select_firms_for_sample(merged_dataset)
+    bhr_event_results = compute_and_save_eawr_bhr(final_dataset, cfg)
+    annual_stock_data = extract_annual_stock_data(bhr_event_results, ds2dsf, cfg)
+    bhr_annual_results = compute_and_save_annual_bhr(cfg)
+
+    final_dataset.to_csv(cfg['prepared_wrds_ds2dsf_path'], index=False)
+    final_dataset.to_parquet(cfg['prepared_wrds_ds2dsf_parquet'], index=False)
+    log.info("Preparation complete.")
+
+def merge_worldscope_link(ws_stock, link_ds_ws):
+    merged = ws_stock.merge(link_ds_ws, on="code", how="inner")
+    return merged[["year_", "item6105", "item5901", "item5902", "item5903", "item5904", "infocode"]]
+
+def pivot_longer_earnings(ws_link_merged):
+    df = ws_link_merged.melt(
+        id_vars=["year_", "item6105", "infocode"],
[…]

### 2025-05-03 code/python/pull_wrds_data-wscp.py

@@ -1,118 +1,0 @@
-# --- Header -------------------------------------------------------------------
-# See LICENSE file for details
-#
-# This code pulls data from WRDS Worldscope and Datastream Databases and links them
-# ------------------------------------------------------------------------------
-import os
-from getpass import getpass
-import dotenv
-import pandas as pd
-from utils import read_config, setup_logging
-import wrds
-
-log = setup_logging()
-
-def main():
-    """
-    Main function to pull data from WRDS.
-
-    This function reads the configuration file, gets the WRDS login credentials, 
-    and pulls the data from WRDS.
-
-    The data is then saved to CSV and Parquet files.
-    """
-    cfg = read_config('config/pull_data_cfg.yaml')
-    wrds_login = get_wrds_login()
-    
-    # Pull data from WRDS
-    worldscope_df, ds_df, linkdata_df = pull_wrds_data(cfg, wrds_login)
-
-    # Save pulled data
-    worldscope_df.to_csv(cfg['worldscope_sample_save_path_csv'], index=False)
-    worldscope_df.to_parquet(cfg['worldscope_sample_save_path'], index=False)
-
-    ds_df.to_csv(cfg['datastream_sample_save_path_csv'], index=False)
-    ds_df.to_parquet(cfg['datastream_sample_save_path'], index=False)
-
-    linkdata_df.to_csv(cfg['link_ds_ws_save_path_csv'], index=False)
-    linkdata_df.to_parquet(cfg['link_ds_ws_save_path'], index=False)
-
-
-def get_wrds_login():
-    """
-    Gets the WRDS login credentials.
-    """
-    if os.path.exists('secrets.env'):
-        dotenv.load_dotenv('secrets.env')
-        wrds_username = os.getenv('WRDS_USERNAME')
-        wrds_password = os.getenv('WRDS_PASSWORD')
-    else:
-        wrds_username = input("Please provide a WRDS username: ")
-        wrds_password = getpass("Please provide a WRDS password (it will not show as you type): ")
-
-    return {'wrds_username': wrds_username, 'wrds_password': wrds_password}
-
-
-def pull_wrds_data(cfg, wrds_authentication):
-    """
-    Pulls data from WRDS (Worldscope, Datastream, and linking table).
-    """
-    db = wrds.Connection(
[…]

### 2025-05-03 code/python/prepare_data.py
update

@@ -1,51 +1,42 @@
 # --- Header -------------------------------------------------------------------
-# Prepare the pulled data for further analysis as per Task 3 requirements
+# Prepare the pulled data for further analysis as per Analysis requirements
 #
 # (C) Mel Mzv - See LICENSE file for details
 # ------------------------------------------------------------------------------
 
-# --- Header -------------------------------------------------------------------
-# Prepare the pulled data for further analysis as per Task 3 requirements
-# ------------------------------------------------------------------------------
-
-import pandas as pd
-from utils import read_config, setup_logging
-
-log = setup_logging()
-
 def main():
-    log.info("Preparing data for analysis ...")
+    log.info("Preparing data ...")
     cfg = read_config('config/prepare_data_cfg.yaml')
 
-    ws_stock = pd.read_csv(cfg['worldscope_sample_save_path_csv'])
-    link_ds_ws = pd.read_csv(cfg['link_ds_ws_save_path_csv'])
-    ds2dsf = pd.read_csv(cfg['datastream_sample_save_path_csv'])
+    ws = pd.read_csv(cfg['worldscope_sample_save_path_csv'])
+    link = pd.read_csv(cfg['link_ds_ws_save_path_csv'])
+    ds = pd.read_csv(cfg['datastream_sample_save_path_csv'])
 
-    ws_link_merged = merge_worldscope_link(ws_stock, link_ds_ws)
-    ws_long = pivot_longer_earnings(ws_link_merged)
-    ws_expanded = expand_event_window(ws_long)
-    merged_dataset = merge_with_datastream(ws_expanded, ds2dsf)
-    final_dataset = select_firms_for_sample(merged_dataset)
-    bhr_event_results = compute_and_save_eawr_bhr(final_dataset, cfg)
-    annual_stock_data = extract_annual_stock_data(bhr_event_results, ds2dsf, cfg)
-    bhr_annual_results = compute_and_save_annual_bhr(cfg)
+    merged = merge_worldscope_link(ws, link)
+    long = pivot_longer_earnings(merged)
+    events = expand_event_window(long)
+    with_ret = merge_with_datastream(events, ds)
+    sample = select_firms_for_sample(with_ret)
[…]

### 2025-05-03 code/python/pull_wrds_data.py
check

@@ -7,7 +7,7 @@
 import os
 from getpass import getpass
 import dotenv
-
+cc
 import pandas as pd
 from utils import read_config, setup_logging
 import wrds

### 2025-05-03 code/python/pull_wrds_data.py
config&code - Neutralize Project Elements

@@ -7,7 +7,7 @@
 import os
 from getpass import getpass
 import dotenv
-cc
+
 import pandas as pd
 from utils import read_config, setup_logging
 import wrds

### 2025-05-03 code/python/test_wrds_connection.py
data - Neutralize Project Elements

@@ -1,6 +1,5 @@
 import os
-import dotenv
-import wrds
+
 
 def load_wrds_credentials():
     '''

### 2025-05-04 doc/presentation.qmd
Update Makefile to include presentation rendering in the build process

@@
---- ## Not rendered in this project; to be used as draft
+--- ## Draft of a presentation - include it in Makefile if you run it apart paper.qmd too
 title: |
   | Earnings Management and Investor Protection:
-  | Accounting Reading Group - Assignment III
+  | Accounting Reading Group
 author: "Mel Mzv"

### 2025-05-04 code/python/do_analysis.py

@@ -1,4 +1,3 @@
-# do_analysis.py — Template for analysis stage (after data prep)
 import os
 import pickle
 import pandas as pd

### 2025-05-04 doc/presentation.qmd
Presentation template

--- 2025-05-03 doc/presentation.qmd
@@
---- ## Not rendered in this project; to be used as draft
+--- ## Draft of a presentation - include it in Makefile if you run it apart paper.qmd too
 title: |
   | Earnings Management and Investor Protection:
-  | Accounting Reading Group - Assignment III
+  | Accounting Reading Group
 author: "Mel Mzv"
-institute: "Humboldt-Universität zu Berlin – Summer 2024"
+institute: "Humboldt-Universität zu Berlin – Summer 2025"
 date: today
@@
 # Motivation
+
+This presentation template serves as a standardized framework for thesis works at WiWi faculty, promoting consistency and professionalism in academic presentations.
 

### 2025-05-04 doc/paper.qmd
Template update - paper.qmd

@@
   | Corporate Decision-Making and Quantitative Analysis
-  | Winter 2024/25 - Individual Report\vspace{1cm}
+  | Summer 2025 - Individual Report\vspace{1cm}
 author:
@@
     lof: true  # Enables List of Figures
-    lot: false  # Disables List of Tables
+    lot: true  # Enables List of Tables
 fig_caption: yes
@@
 **CRSP**: Center for Research in Security Prices  
-**CQA**: Corporate Decision-Making and Quantitative Analysis  
 **DS**: Datastream  
-**DV**: Dependent Variable  
-**HU**: Humboldt-Universität zu Berlin  
 **IDE**: Integrated Development Environment  
-**IV**: Independent Variable  
 **TREAT**: TRR 266 Template for Reproducible Empirical Accounting Research  
-**US**: United States  
 **WRDS**: Wharton Research Data Services  
@@
 # Introduction {#sec-introduction}
-The aim of this paper is to illustrate the use of open science tools in empirical accounting research.
-This project builds on the methodological framework established in the Corporate Decision-Making and Quantitative Analysis (CQA) course, which explored a range of empirical research methods—including archival analysis, field experiments, and survey-based approaches—to develop a comprehensive understanding of corporate decision-making and quantitative analysis.
-Expanding on prior empirical research conducted in Assignments I and II, this paper extends the empirical focus of the course.
-Assignment I examined audit market concentration in the EU using Transparency Reports data, offering insights into the dominance of major audit firms and the structure of the European audit market.
-Assignment II explored the economic implications of corporate financial reporting, focusing on @graham_economic_2005 to analyze managerial decision-making, earnings management strategies, and voluntary disclosure practices.
+This project serves as a template-based example for conducting and documenting reproducible empirical accounting research.
+It leverages the TRR 266 Template for Reproducible Empirical Accounting Research (TREAT) and builds upon the methodological foundations of the Corporate Decision-Making and Quantitative Analysis course, as well as the Accounting Reading Group course, which explored archival and non-archival research methods.
+To support future empirical thesis projects, this template was adapted and further developed specifically for students at the School of Business and Economics at Humboldt-Universität zu Berlin—particularly those affiliated with the Institute of Accounting and Auditing and the Finance Group.
 
-Building on these foundations, Assignment III investigates the informativeness of quarterly earnings announcements and their contribution to annual share price movements.
-The project replicates and extends @ball_how_2008, assessing the extent to which earnings releases serve as a source of new information for financial markets.
-By integrating multiple datasets from WRDS and applying an event-study methodology, this study aims to evaluate the role of earnings disclosures in market efficiency, showcasing the valuation purpose of accounting.
-At the same time, earnings reports may play a crucial role in contracting by influencing managerial compensation, debt agreements, and other contractual mechanisms tied to financial performance [@ball_how_2008].
-The integration of multiple databases adds complexity, requiring a detailed understanding of their structures and careful scripting to extract, align, and analyze data effectively.
+Following prior assignments on audit market concentration and financial reporting effects (e.g., @graham_economic_2005), this project focuses on the informativeness of quarterly earnings announcements.
+Specifically, it replicates and extends the study by @ball_how_2008, which examines whether earnings announcements contribute new information to markets or merely confirm what is already known—contributing to the long-standing discussion on market efficiency and the valuation versus contracting roles of accounting.
 
-Beyond empirical replication, the project also documents research design choices, replication steps, and explicit assumptions made whenever the original paper was unclear on how to proceed.
-It discusses variations between original and reproduced results and provides insights into the timeliness and market impact of earnings announcements across different timeframes and jurisdictions.
-Additionally, it sketches a research design for a non-archival study to evaluate the paper’s findings through an alternative methodological lens.
-This project leverages the TRR 266 Template for Reproducible Empirical Accounting Research (TREAT) to establish an open science-oriented infrastructure, ensuring transparency and replicability for empirical accounting research.
+Using updated U.S.
+data and applying the methodology to a second country, this thesis combines multiple WRDS datasets and applies an event-study design to evaluate how much of the annual stock return variation can be explained by earnings announcement windows.
+The approach is designed to foster transparent research workflows, with a fully scripted pipeline for data pulling, preparation, analysis, and reporting.
 
-Earnings announcements provide investors with valuable information about a firm’s market value, with stock prices reacting significantly when earnings news deviate from expectations [@fink_review_2021, p.2].
-The seminal study by @ball_empirical_1968 was the first to document this relationship, showing that stock prices anticipate earnings surprises, with most of the market reaction occurring before the official announcement, suggesting that earnings reports primarily confirm rather than introduce new information.
-Over the past decades, more than a thousand studies have examined the interplay between capital markets and financial statements, a research stream that originated with Ball and Brown (1968)  [@kothari_capital_2001].
-
-Building on this foundation, the original study by @ball_how_2008 investigates the extent to which quarterly earnings announcements contribute new information to the market, assessing their role in shaping annual share price movements.
-By estimating the R² from regressions of annual stock returns on earnings announcement window returns, the paper quantifies the informativeness of earnings releases, finding that they account for only 1% to 2% of total annual volatility.
-This challenges the assumption that earnings provide substantial new information and instead suggests a confirmatory role in financial reporting.
-These findings align with the semi-strong form of market efficiency, which suggests that stock prices already reflect all publicly available information, including prior signals about earnings.
-Unlike traditional event studies that assume market efficiency, @ball_how_2008 allow for potential market mispricing by not imposing a unit slope restriction in their regressions, meaning that their approach can capture under- or overreaction to earnings announcements.
-They also emphasize that extending the event window beyond the earnings announcement period would shift the research focus from earnings “surprise” to broader financial reporting effects, such as its role in debt contracting and executive compensation.
-
-While @ball_how_2008 provide key insights into the informativeness of earnings announcements, their study was conducted within a specific timeframe and market context.
-Since then, financial markets have undergone significant changes due to regulatory reforms, economic crises, and technological advancements in financial information processing.
-Moreover, the extent to which earnings informativeness varies across international markets remains underexplored.
-By replicating their study with updated data from 2007–2023 and applying the methodology to Canada as a non-U.S.
-country, this project reassesses the robustness of their findings.
-Through this approach, it critically evaluates the generalizability of the original results.
-
-The paper is structured into sections corresponding to Tasks 1–3 (@sec-task1, @sec-task2, @sec-task3), each detailing the research design choices and assumptions, documenting the replication steps and results.
-@sec-task1 follows an exact replication approach by replicating key tables and figures from the original study using US data (1972–2006) to establish baseline results.
-@sec-task2 extends the analysis through empirical generalization, applying the same methodology to 2007–2023 US data to compare how results evolve over time.
-@sec-task3 employs a generalization and extension approach by applying the methodology to a non-US country, requiring adjustments for different databases and market structures.
-This framework aligns with the replication taxonomy outlined by @salterio_replication_2022.
-Finally, @sec-task4 sketches the survey design as a non-archival study that allows evaluating the key findings of the seminal paper, following Bloomfield’s (2016) triangulation principle.
-The concluding remarks are provided in @sec-conclusion.
+The paper is organized into four main parts.
+In @sec-task1, I replicate the core findings of the original study using U.S.
+data from 1972 to 2006.
+In @sec-task2, I apply the same methodology to U.S.
+data from 2007 to 2023 to explore temporal robustness.
+In @sec-task3, I generalize the findings to a non-U.S.
+country to assess cross-country variation and data integration requirements.
+Finally, @sec-task4 sketches a non-archival research design using a survey-based approach to evaluate the underlying assumptions of the original paper.
+The thesis concludes with reflections and key insights in @sec-conclusion.
 
 # Task 1 - Replication of Key Tables and Figures {#sec-task1}
-Task 1 is the most comprehensive, replicating the methodology of key tables and figures from @ball_how_2008 using daily US stock return data (1972–2006) to establish baseline results, ensuring alignment with the original findings before extending the analysis in @sec-task2.
-By reconstructing Table 1 (Panel A), Table 2 (all panels), and Figure 1 (all panels), the task evaluates the robustness of the original results and identifies potential discrepancies, offering insights into data consistency, methodology, and market dynamics over the examined period.
-The use of daily CRSP data is essential for capturing short-term stock price movements around earnings announcements, ensuring precise event window calculations.
-The discussion of replication findings is presented in @sec-results1.
+Task 1 replicates key tables and figures from @ball_how_2008 using daily US stock return data (1972–2006) to establish baseline results.
+By reconstructing Table 1 (Panel A), Table 2, and Figure 1, it evaluates the robustness of the original findings and highlights any discrepancies.
+Daily CRSP data ensures accurate measurement of stock reactions around earnings announcements.
+Replication results are discussed in @sec-results1, with extended analysis in @sec-task2.
 
 ## Research Design Choices and Assumptions {#sec-research_design_assumptions1}
-Following @ball_how_2008, I focus on earnings announcements and stock returns from 1972 to 2006, ensuring alignment with the original study’s methodology.
-This replication mirrors the research design, maintaining consistency in sample selection, event window definitions, and return calculations.
+In line with @ball_how_2008, this replication examines U.S.
+earnings announcements and stock returns from 1972 to 2006, following the original study’s sample design, event window (-1 to +1), and return computations.
+Unlike the original, I omit analyst expectations, focusing solely on returns and announcement dates for simplicity.
 
-Based on the original paper, I do not gather analyst expectations or estimate earnings deviations, requiring only stock return data and earnings announcement dates, simplifying data collection.
+To address gaps in methodological detail, several assumptions are applied.
+Event windows follow CRSP trading calendar conventions, shifting dates when earnings fall on non-trading days [@center_for_research_in_security_prices_crsp_crsp_2024].
+While @ball_how_2008 require a minimum of 240 trading days, I relax this constraint, as per course guidance, resulting in a broader sample.
+All firms in Compustat meeting the quarterly announcement and return data criteria are included, regardless of activity status.
+I use PERMNO to uniquely track securities, avoiding aggregation across multiple share classes [@gow_empirical_2025].
+The CRSP/Compustat Merged (CCM) database is used to link identifiers, noting that differences from the original study may arise due to updated data (WRDS version: January 2025).
+As in @ball_how_2008, the analysis is return-based (RET in CRSP), excluding bid-ask pricing.
+Earnings are grouped by announcement date (RDQ), not fiscal period, to align directly with CRSP trading dates.
 
-To ensure clarity where @ball_how_2008 lacks explicit guidance, I impose the following assumptions:
-
-1.	Earnings announcement windows are defined as the three-day period surrounding the release (-1 to +1).
-According to @center_for_research_in_security_prices_crsp_crsp_2024 guide,since CRSP excludes weekends and holidays, if Day 0 falls on a non-trading day, it shifts to the next trading day.
-Similarly, Day -1 or +1 adjusts to the nearest available trading day.
-
-2.	Firm-years are selected based on available CRSP stock return data.
-While @ball_how_2008 restricts the sample to firms with at least 240 trading days per year, I follow course guidance and do not impose this restriction, allowing for a slightly larger sample.
-
-3.	Compustat includes active and inactive firms; since the original study does not filter based on firm activity, all firms meeting sample selection criteria (four quarterly earnings announcements per year, available stock returns) are retained.
-
-4.	Permanent identifiers (`PERMNO`, `PERMCO`, `GVKEY`) ensure accurate historical tracking [@gow_empirical_2025].
-`PERMNO` is used instead of `PERMCO` as it uniquely identifies individual securities rather than companies, preventing aggregation issues from multiple share classes.
-
-5.	The replication relies on Compustat and CRSP, linked via the CRSP/Compustat Merged (CCM) database.
-As Compustat updates daily, while CRSP and CCM update annually, timing mismatches may arise.
-Unlike the original study, this replication uses the latest WRDS version as of January 2025, meaning database updates may introduce firm coverage or value differences.
-
-6. Since @ball_how_2008 analyze returns, I use `RET` variable in CRSP too to simplify the process. Bid and ask prices are ignored.
-
-7.
-Same as @ball_how_2008, I group earnings announcements based on their actual announcement dates (`RDQ` from Compustat) rather than the fiscal periods they cover.
-Since Compustat provides the precise earnings announcement date and CRSP aligns stock returns with trading days (`date` in CRSP), this approach ensures consistency.
-So, no adjustments are made to realign quarters.
-
-By following the steps outlined in @sec-replication_steps1 and adhering to the assumptions made, I aim to replicate the results from @ball_how_2008.
-A thorough step-by-step approach, with each step clearly documented, helps to understand and verify the outputs.
+These assumptions, together with the procedural details in @sec-replication_steps1, guide the replication of @ball_how_2008 and ensure transparency in design and implementation.
 
 ## Replication Steps {#sec-replication_steps1}
-The analysis involves gathering and filtering stock return and earnings announcement data, linking CRSP stock data with Compustat firm-level accounting data fundamentals, and estimating the relationship between annual returns and earnings-announcement window returns.
-The replication process follows a structured workflow of data pull, preparation, and analysis, ensuring consistency with the original study.
+This section outlines the structured replication process: data pulling, preparation, and analysis, aligned with @ball_how_2008.
 
 **Step 1: Pulling the Data and Managing the Databases**  
-Data is pulled from WRDS using Python to integrate daily stock returns from CRSP (`crsp_a_stock.dsf`) with quarterly earnings announcement data from Compustat  
-(`comp_na_daily_all.fundq`), linked via the Compustat-CRSP link table  
-(`crsp_a_ccm.ccmxpf_lnkhist`).
-`permno` unique identifier matches stock returns to firm fundamentals via `gvkey` for exact firm/year match.
-Key variables include `ret` (daily stock return) and `rdq` (earnings announcement date), enabling event study analysis from 1972 to 2023.
-While WRDS provides extensive financial datasets, documentation on precise variable selection and dataset integration is limited, requiring additional validation and cross-referencing with prior research.
-The data is pulled for the entire range 1972-2023 required for Task 1 and 2 to prevent double preparation step.
+Data from WRDS combines daily CRSP stock returns (dsf) and Compustat quarterly earnings announcements (fundq), linked via the CRSP-Compustat link table (ccmxpf_lnkhist).
+The `permno` and `gvkey` identifiers ensure accurate firm-year alignment.
+Key variables are `ret` (returns) and `rdq` (announcement dates), pulled for the full 1972–2023 period used in Tasks 1 and 2.
 
 **Step 2: Data Preparation**  
-Although not finished, the plan would be as follows: The merged dataset would integrate stock return data from CRSP (`crsp_daily_stock_returns`), firm fundamentals from Compustat (`compustat_fundq_1972_2023`), and the CRSP-Compustat linking table (`linkdata_compustat_crsp.parquet`).
-The `PERMNO` identifier from CRSP and `GVKEY` from Compustat are connected via `LPERMNO` in the linking table, ensuring correct firm-year alignment.
-Unnecessary variables from linked table, such as `linkprim`, `liid`, `linktype`, and `usedflag`, are dropped.
-The resulting dataset would retain only daily stock returns (`ret`), earnings announcement dates (`rdq`), fiscal year-end indicators (`fyr`), and firm identifiers (`permno, gvkey`), used further for Buy-and-Hold Returns (BHR) calculation and regression analysis.
-
-After defining the event windows for each earnings announcement day in Q1-Q4, retain only the firms that have four earnings announcements per year, such that each of them falls within the same calendar year.
-Additionally, for the subgroup replication of Table 2 Panels C-D classify the firms based on their fiscal year-end (`fyr`) to define whether firms have December fiscal year-end firm or in any another month.
-Control for the duplicate and missing firm/years.
+The merged dataset includes CRSP, Compustat, and the link table.
+After removing irrelevant linking variables, the dataset is trimmed to key variables: `ret`, `rdq`, `fyr`, `permno`, and `gvkey`.
+Only firms with four earnings announcements per calendar year are retained.
+Firms are also classified by fiscal year-end (`fyr`) to match Table 2 Panels C–D.
+Duplicate or incomplete firm-years are excluded.
 
 **Step 3: Analysis Implementation and Reproduction of Tables and Figure**  
-The analysis step computes the BHR returns and performs a cross-sectional regression for each distinct firm/year.
-
-Since @ball_how_2008 do not explicitly describe their return computation, I compute Buy-and-Hold Returns (BHR) following the cumulative return formula as discussed in @gundersen_returns_2022.
-The BHR formula captures compounded returns over multiple days, reflecting the stock’s actual performance, which is used to replicate Table 1.
+BHRs are computed using the cumulative return formula per @gundersen_returns_2022.
+Cross-sectional regressions are then run for each firm-year to replicate Table 1, Table 2, and Figure 1 from the original paper.
 
 ## Results {#sec-results1}
-Since Tables and Figure replication not finished, this section presents only the original @fig-fig1_original.
+This section presents the replication results.
 
 # Task 2 - Extending the Analysis to 2007–2023 {#sec-task2}
-This Section extends the analysis to 2007–2023, a period defined by major economic events such as the 2008 financial crisis and the COVID-19 pandemic, which may have influenced the informativeness of earnings announcements.
-While the task is relatively straightforward in terms of programming—requiring adjustments to the sample period—it demands careful economic interpretation of observed trends, market shifts, and potential institutional details that may influence the relationship between earnings announcements and stock returns.
+This section extends the analysis to 2007–2023, a period shaped by events like the 2008 financial crisis and the COVID-19 pandemic.
+Although the programming involves only a sample period update, the task requires thoughtful economic interpretation of how structural shifts and evolving market conditions may have affected the informativeness of earnings announcements.
 
@@
 ## Results {#sec-results2}
-The sample shift to 2007–2023 is expected to bring potential shifts in earnings announcement informativeness due to significant economic and regulatory events.
-The 2008 Global Financial Crisis likely disrupted earnings predictability, with heightened volatility and uncertainty increasing the market’s sensitivity to earnings reports.
-The COVID-19 pandemic 2020 introduced extreme market uncertainty, potentially amplifying the informativeness of earnings announcements as investors relied on them to gauge firm resilience and recovery.
-Additionally, the introduction of Regulation (EU) No 537/2014 may have influenced earnings informativeness by improving transparency, audit quality, and financial reporting compliance across EU countries.
-Differences in compliance and audit coverage could have led to variations in how earnings announcements reflect new information in stock prices, potentially contributing to cross-country disparities in earnings informativeness.
+Shifting the sample to 2007–2023 introduces potential changes in earnings announcement informativeness driven by key economic and regulatory events.
+The 2008 financial crisis and the 2020 COVID-19 pandemic likely heightened market uncertainty, increasing reliance on earnings disclosures to assess firm stability.
+Moreover, Regulation (EU) No 537/2014 may have enhanced financial reporting transparency and audit quality in EU countries, potentially affecting how earnings news is reflected in stock prices and contributing to cross-country variation in informativeness.
 
 # Task 3 - Cross-Country Replication {#sec-task3}
-Task 3 extends the analysis beyond the U.S.
-by replicating Figure 1 using 1972–2023 data from a non-U.S.
-market, examining how earnings informativeness varies across regulatory environments and market structures.
-This requires mapping CRSP/Compustat variables to their Worldscope/Datastream equivalents while adjusting for accounting standards, market liquidity, and institutional factors.
-Worldscope provides earnings announcement data, and Datastream covers stock returns.
-While the core methodology remains, careful data handling and interpretation are essential for cross-country analysis, particularly with Canada.
+Task 3 generalizes the analysis to a non-U.S.
+market by replicating Figure 1 using 1972–2023 data from Canada.
+This involves aligning CRSP/Compustat variables with Worldscope and Datastream equivalents while accounting for differences in reporting standards, market liquidity, and institutional context.
+Though the core methodology is retained, adapting it to international data demands careful mapping and interpretation.
 
 ## Research Design Choices and Assumptions {#sec-research_design_assumptions3}
-Since this task employs a generalization and extension approach by applying the same methodology to a country different from original report, it requires adjustments for different databases and market structures.
-Based on the original paper, I do not gather analyst expectations or estimate earnings deviations for my replication.
-Instead, I only require stock return data and earnings announcement dates.
+Since Task 3 applies a generalization and extension approach to a non-U.S.
+market, it requires adapting the methodology to alternative databases and market structures.
+Like in the original study, no analyst forecasts or earnings surprises are used—only stock returns and earnings announcement dates are required.
+Canada was selected for its comparable quarterly reporting frequency and institutional similarity to the U.S., making it a suitable choice for cross-country replication [@short_short_2025].
+Canadian firms reporting under U.S.
+GAAP further enhance comparability.
 
-Canada was chosen for its quarterly reporting frequency, ensuring comparability with U.S.
-data, as most firms report four times per year, with some semi-annual reports [@short_short_2025].
-Despite its smaller market size, Canada’s regulatory environment, liquidity, and institutional settings closely align with the U.S., making it an ideal candidate.
-Additionally, Canadian firms registered with the SEC can report under U.S.
-GAAP, making their financial disclosures fully compatible with U.S.
-firms.
-
-In addition, I impose the following assumptions apart from those listed in @sec-research_design_assumptions1 to ensure generalization and extension approach due to usage of other databases or where @ball_how_2008 do not provide explicit guidance:
-
-1.
-To ensure a smooth and conflict-free process, I handle the WSCP/Datastream workflow separately, preventing potential interference with the CRSP/Compustat pull process and improving debugging efficiency.
-Although a unified pull/prepare/do_analysis file for all databases would be ideal and streamline management, the large-scale nature of the extracted datasets demands modular approach.
-
-2.
-Since I do not perform subgroup analysis for December and non-December fiscal year end firms as in Table 2 Panels C-D, `item5350` for FISCAL PERIOD END DATE is not relevant for this task, as Figure 1 applies to the analysis of the entire dataset.
-So, the sample is all firm-years with available data on the quarterly Worlscope and daily Datastream files.
-
-3.
-I verified whether Worldscope and Datastream follow the same earnings announcement date convention as Compustat or if they report earnings based on fiscal periods.
-I decided to use Worldscope items 5901–5904 from historical stock-related data in `wrds_ws_stock`, which represent the actual date when a company publicly reports its earnings for each quarter, where, however, data for these fields is generally not available prior to 1992 [@thomson_financial_worldscope_2007, p.281].
-This may cause limited data sample size.
-
-4.
-This task is a generalization to a non-U.S.
-setting, meaning methodological differences will certainly arise due to variations in data structure and reporting standards.
-Worldscope - Stock Data `wrds_ws_stock` table is updated quarterly, while Datastream Daily Stock File `wrds_ds2dsf` is updated weekly, which may introduce timing mismatches when aligning financial fundamentals with stock returns.
-Given that, future applications should consider database updates and coverage differences between CRSP and Datastream.
-
-5. Similar to @sec-task1, I use returns `ret` variable in Datastream for returns, presented as percentage change. Bid and ask prices are ignored.
-
-6. I ignore the currency effect (CAD or USD) in DS table, as returns are measured in percentages, making them relative rather than absolute values.
+To address database-specific issues not covered in @sec-research_design_assumptions1, I make the following adjustments: First, Worldscope/Datastream workflows are kept separate from CRSP/Compustat to prevent interference and facilitate debugging, given the scale of data involved.
+Second, I exclude fiscal-year-end-based subgroup analysis (Table 2 Panels C–D), as Figure 1 uses the full sample without splitting by fiscal period.
+Third, I use Worldscope items 5901–5904 to define earnings dates, though limited availability before 1992 may reduce the sample size [@thomson_financial_worldscope_2007, p.281].
+Fourth, differences in update frequency between quarterly Worldscope and weekly Datastream can lead to timing mismatches when aligning fundamentals and returns.
+Fifth, I rely on Datastream’s `ret` variable for percentage returns, ignoring bid/ask spreads.
+Finally, currency effects are ignored, as percentage returns are unaffected by currency denomination.
 
@@
 **Step 1: Pulling the Data and Managing the Databases**  
-I follow the linking approach between Worldscope and Datastream based on @dai_research_2021.
-The original study uses data from 1972 onward, while Worldscope's `year_` variable only provides data starting from 1980 [@wharton_research_data_services_worldscope_2025].
-This constraint may lead to differences in sample coverage.
-Hence, filter Datastream `marketdate` to 1980 to prevent mismatched observations where stock data exists but earnings do not.
-
-Next, retrieve quarterly earnings announcement dates from Worldscope and daily stock returns from Datastream, filtering for Canadian firms from 1980 onward.
-The datasets are saved separately and later merged using firm-level identifiers.
-Only common equity (`typecode` = EQ) is included to ensure consistency with CRSP.
-
-The pull step results into three comprehensive datasets: DS table with 27,950,012, Link table with 77,944, WSCP table with 1,942,202 observations.
+Following @dai_research_2021, I link Worldscope and Datastream using code and infocode.
+Since Worldscope’s year_ starts in 1980 [@wharton_research_data_services_worldscope_2025], Datastream data is also restricted to 1980–2023 to ensure consistent coverage.
+I extract quarterly earnings announcement dates (items 5901–5904) from Worldscope and daily stock returns from Datastream, focusing on Canadian firms and filtering for common equity (typecode = EQ).
+This step yields three datasets: 27.9M (Datastream), 77.9K (link table), and 1.94M (Worldscope) observations.
 
 **Step 2: Data Preparation**  
-First, I merged Worldscope stock data with the linking table using `code` as unique identifier, resulting in 747,760 observations.
-This merged dataset was then joined with Datastream daily stock data using unique `infocode`, producing a final dataset with 10,249,737 observations.
-
-If a firm had missing values for any of the Earnings Per Share Report Date fields (item5901–item5904), it did not meet the criteria set by @ball_how_2008 and was excluded from the sample.
-This filtering step removed 3,609,492 missing firm-year observations.
-Additionally, 5,996,793 firm-year entries were dropped because their four earnings announcements did not fall within the same calendar year.
-Furthermore, I extend dynamically shift the missing values for trading days -1,0,1 within 3 day window, due to small sample size to keep more earning announcemnet windows.
-
-From this cleaned dataset, BHR event returns are computed.
-Then, annual stock returns for the firms of the filtered sample are extracted from Datastream pulled data .
-Finally, the Annual BHR Returns are calculated.
-Notably, after pulling the annual stock data for further annual return calculation, all infocodes present an average of 250 trading days, highlighting full year stock coverage.
+Worldscope is first merged with the link table via code, producing 747,760 rows.
+This is then joined with Datastream via infocode, resulting in 10.2M observations.
+I exclude firms missing any earnings date fields, removing 3.6M rows, and drop 6M rows where earnings do not span all four quarters within the same calendar year.
+To retain more valid windows, event days (-1, 0, 1) are dynamically shifted within ±3 days when needed.
+After cleaning, I compute 3-day BHRs and extract full-year stock return data from Datastream.
+All retained infocode entries have around 250 trading days per year, confirming data completeness for annual BHR computation.
 
 **Step 3: Analysis Implementation and Reproduction of Tables and Figure**  
-I conduct the earnings informativeness analysis by computing summary statistics (Table 1), running annual regressions (Table 2), and generating a replication of Figure 1.
-First, I loaded the computed event-window and annual buy-and-hold returns (BHR_3day, BHR_Annual) and produced summary statistics similar to @ball_how_2008.
-Next, I performed annual cross-sectional regressions of stock returns on earnings-announcement returns, benchmarking against an Abnormal R² (adjusted R² minus 4.8%), while handling missing data and ensuring a valid number of observations per year.
-Finally, the script plots Figure 1, visualizing Abnormal R² trends (Panel A) and slope coefficients (Panel B) over time, saving it as both a PNG and a pickle file for further use.
+The analysis replicates Table 1 (summary statistics), Table 2 (yearly regressions), and Figure 1 (Abnormal R² and slope trends).
+Using `BHR_3day` and `BHR_Annual`, I run yearly regressions of annual returns on earnings-window returns and compute Abnormal R² (adjusted R² – 4.8%), handling missing values appropriately.
+The resulting trends are plotted and saved as both PNG and pickle files for use in the paper.
 
 ## Results {#sec-results3}
-Comparing the replication results with those from @ball_how_2008 (see @fig-fig1_original) reveals both similarities and discrepancies.
+Comparing the replication results with those from @ball_how_2008 reveals both similarities and discrepancies.
 The original figure shows a general upward trend in abnormal  R² , peaking in the early 2000s, whereas the replicated figure captures greater volatility in the later years (2010–2023).
@@
 
-The research design follows the generalization goal in empirical literature, emphasizing the need to validate findings across different methods [@bloomfield_gathering_2016].
-@ball_how_2008 assess earnings informativeness through stock price reactions, but triangulation—beyond replication—requires diverse methods to examine the same question [@bloomfield_gathering_2016, p.353].
-While price reactions indicate market response, they do not capture investor perceptions.
-A survey directly evaluates whether market participants view earnings announcements as confirmatory or informative.
-If investors confirm a secondary role, it reinforces Ball and Shivakumar’s (2008) findings, ensuring informativeness reflects cognitive and behavioral mechanisms rather than price fluctuations.
-These insights could inform corporate reporting strategies by clarifying how investors prioritize disclosures.
+To extend the analysis beyond archival methods, this section proposes a survey to assess whether investors perceive earnings announcements as informative or merely confirmatory, following the triangulation principle in empirical research [@bloomfield_gathering_2016].
+While @ball_how_2008 measure market reactions via stock returns, surveys can capture investor beliefs directly, offering complementary evidence on earnings informativeness.
 
-A survey is the most suitable non-archival method, capturing investor sentiment beyond what archival data provides.
-Unlike lab experiments, which lack real-world applicability, surveys measure the dependent variable (DV) by eliciting participant perceptions [@bloomfield_gathering_2016, p.358].
-Field experiments, though capturing real-time reactions, require direct intervention, posing ethical and logistical challenges.
-A long-term research program combining surveys and field studies would enhance validity, leveraging unlimited funds to assess investor decision-making and market responses.
+A structured survey targeting 300 institutional and retail investors will be distributed through financial networks and platforms (e.g., LinkedIn, CFA societies), with a pilot study refining its clarity.
+Participants will receive a €20 honorarium, and responses will remain anonymous.
+The survey measures perceived informativeness (DV) using Likert-scale and multiple-choice questions, with IVs capturing factors such as investor preparedness and reaction timing.
 
-To ensure survey reliability, design strategies follow @bruggemann_worku_2024.
-A sample of 300 investors, analysts, and financial professionals will provide institutional and retail perspectives while balancing statistical power and feasibility.
-The survey will be distributed via email through financial networks, investor associations, and professional platforms (e.g., LinkedIn, CFA societies) and in paper format at finance conferences.
-The Humboldt-Universität zu Berlin (HU) brand may be used, subject to approval.
-Respondents will be assured confidentiality and offered an €20 honorarium.
-A pilot study with 30 participants will refine question clarity, survey length, and response distribution, helping identify unclear or missing questions and assess completion time.
-
-To minimize response bias, questions will be neutrally worded to capture pre- and post-announcement perceptions.
-The fully anonymous survey ensures candid responses without requiring personal financial history.
-
-The **DV** is the perceived informativeness of earnings announcements, measured through survey responses on whether they are confirmatory or informative.
-Independent variables (**IVs**) include investor preparation, market awareness, reaction timing, perceived novelty of earnings, key financial indicators, and external conditions.
-These are measured via Likert-scale, ranking, and multiple-choice questions, supplemented by qualitative insights from follow-up interviews.
-
-To deepen findings, selected participants will be invited for confidential follow-up interviews, ensuring diversity in investment strategies, industries, and expertise.
-These interviews will be conducted via video conferencing or phone calls.
-The following sections provide detailed sketches of the survey and interview designs.
+Follow-up interviews with selected respondents will add qualitative depth, enabling a richer understanding of investor perspectives.
+This mixed-method approach enhances validity and supports generalization of findings beyond price-based evidence.
 
@@
 # Conclusion {#sec-conclusion}
-This project effectively demonstrates the use of a systematic and collaborative workflow for empirical accounting research, leveraging the TRR 266 Template for Reproducible Empirical Accounting Research.
-By following an open science approach, I aimed to replicate key tables and figures from @ball_how_2008, providing insights into how earnings announcements contribute to price formation and market efficiency.
-By extending the analysis beyond the original study’s U.S.
-market focus, the project highlights how market structures and investor behaviors influence the informativeness of earnings disclosures in different jurisdictions.
+This thesis template demonstrates how the TRR 266 framework can be applied to structure a reproducible and transparent empirical accounting study.
+Using Quarto proved to be a particularly effective and user-friendly solution for structuring and rendering the thesis.
+It facilitates seamless integration of code, results, and narrative, making it an ideal tool for students at the Institute of Accounting and Auditing or the Finance Group at the School of Business and Economics, HU Berlin.
 
-The study contributes to the ongoing discussion on market efficiency and the economic role of corporate disclosures by providing empirical evidence on how earnings informativeness has evolved over time and across jurisdictions.
-The cross-country analysis with Canada highlights variations in market responses to earnings announcements, suggesting that institutional factors, market structures, and investor behavior play a role in shaping how financial information is processed.
-These findings have implications for investors, analysts, and policymakers, as they underscore the importance of regulatory consistency and market transparency in shaping the effectiveness of financial reporting.
-
-This assignment required a comprehensive application of programming skills and institutional knowledge gained throughout the course, integrating data analysis, replication, and visualization techniques in line with open science principles.
-In the future, this repository can be cloned or forked (if made public) to facilitate further research on earnings informativeness, enabling additional extensions or robustness tests.
-Additionally, the survey developed as part of the non-archival study proposal can be further refined and expanded, providing a structured framework for gathering primary data on market participants’ interpretations of earnings announcements.
-Through this approach, the study not only revisits a fundamental question in financial research but also provides a foundation for future empirical investigations into the evolving role of earnings announcements in global capital markets, both through archival and non-archival methodologies.
+Beyond replication, this project integrates programming, data processing, and visualization skills, offering a foundation for future extensions—whether through archival analysis or survey-based designs.
+As an open and modular template, this repository is designed to support ongoing student and academic work in empirical research.
 Thanks for reading!

### 2025-05-04 doc/declaration.tex
Declaration template - submit

\newpage
\thispagestyle{empty}  % <--- Hides page number

\section*{Declaration of Originality}

I hereby declare that the present seminar paper and the work reported herein was composed by and originated entirely from me without any help. All sources used from published or unpublished work of others are reported in the list of references. All parts of my work that are based on others’ work are cited as such.

This seminar paper has not been submitted for any degree or other purposes, neither at the Humboldt-Universität zu Berlin nor at any other university or college.

\vspace{20mm}

Berlin, \hrulefill

\vspace{30mm}

\hrulefill \\
Name and surname \hspace{20mm} Signature

### 2025-05-04 doc/paper.qmd
Declaration template - submit

@@
 author:
-  - name: Mel Mzv
+  - name: Max Mustermann
     email: m@student.hu-berlin.de
@@
     lot: true  # Enables List of Tables
+    include-before-body: 
+      text: |
+        \pagenumbering{Roman}
+    include-after-body:
+      - declaration.tex
+
 fig_caption: yes
@@
 \pagebreak
+\pagenumbering{arabic}
 

### 2025-05-04 doc/declaration.tex
Declaration - signature

@@
 
-\section*{Declaration of Originality}
+\section*{Declaration of Academic Honesty}
 
-I hereby declare that the present seminar paper and the work reported herein was composed by and originated entirely from me without any help.
-All sources used from published or unpublished work of others are reported in the list of references.
-All parts of my work that are based on others’ work are cited as such.
+I, Max Mustermann, hereby declare that I have not previously submitted the present work for other examinations.
+I wrote this work independently.
+All sources, including sources from the Internet, that I have reproduced in either an unaltered or modified form (particularly sources for texts, graphs, tables and images), have been acknowledged by me as such.
 
-This seminar paper has not been submitted for any degree or other purposes, neither at the Humboldt-Universität zu Berlin nor at any other university or college.
+I understand that violations of these principles will result in proceedings regarding deception or attempted deception.
 
-\vspace{20mm}
+\vspace{2cm}
 
-Berlin, \hrulefill
+\rule{60mm}{0.4pt}
 
-\vspace{30mm}
+\vspace{0.5cm}
 
-\hrulefill \\
-Name and surname \hspace{20mm} Signature
+\text{Max Mustermann}
+
+Berlin, \today

### 2025-05-04 doc/abstract.tex
Added abstract, acknowledgments, list of abbreviations

--- 2025-05-04 doc/paper.qmd
@@
----
-title: |
-  | How Much New Information Is There in Earnings?
-  | Corporate Decision-Making and Quantitative Analysis
-  | Summer 2025 - Individual Report\vspace{1cm}
-author:
-  - name: Mel Mzv
-    email: m@student.hu-berlin.de
-    affiliations:
-      - Humboldt-Universität zu Berlin  
-date: today
-date-format: MMM D, YYYY [\vspace{0cm}] # Adjusted the vertical space after the date from 1cm
-abstract: |
-  | This project uses the TRR 266 Template for Reproducible Empirical Accounting Research (TREAT) to provide infrastructure for open science-oriented empirical projects.
+\clearpage
+\pagenumbering{Roman}
+\setcounter{page}{3}  % Optional: set to next page after acknowledgements
+\section*{Abstract}
+\addcontentsline{toc}{section}{Abstract}
+
+This project uses the TRR 266 Template for Reproducible Empirical Accounting Research (TREAT) to provide infrastructure for open science-oriented empirical projects.
 Leveraging data from the CRSP and Compustat databases, alongside Worldscope and Datastream via WRDS, this repository showcases a reproducible workflow integrating Python scripts for data preparation, analysis, and visualization.
 Integrating multiple databases adds complexity, requiring a detailed understanding of their structures and careful scripting to extract, align, and analyze data effectively.
+
 The project replicates and extends Ball and Shivakumar (2008) to analyze the informativeness of quarterly earnings announcements and their contribution to share price movements, highlighting their critical role in investment decisions and impact on investors and analysts.
@@
 This code base, adapted from TREAT, demonstrates how the template applies to this project and serves as a structured guide for reproducible empirical research in accounting.
-  <!--
-  | \vspace{6cm} commenting away to decrease distance
-  -->
 
-bibliography: references.bib
-biblio-style: apsr
-format:
-  pdf:
-    documentclass: article
-    number-sections: true
-    toc: true
-    lof: true  # Enables List of Figures
-    lot: true  # Enables List of Tables
-fig_caption: yes
-fontsize: 11pt
-ident: yes
-always_allow_html: yes
-number-sections: true 
-header-includes:
-  - \usepackage[nolists]{endfloat}    
-  - \usepackage{setspace}\doublespacing
-  - \setlength{\parindent}{4em}
-  - \setlength{\parskip}{0em}
-  - \usepackage[hang,flushmargin]{footmisc}
-  - \usepackage{caption} 
-  - \captionsetup[table]{skip=24pt,font=bf}
-  - \usepackage{array}
-  - \usepackage{threeparttable}
-  - \usepackage{adjustbox}
-  - \usepackage{graphicx}
-  - \usepackage{csquotes}
-  - \usepackage{indentfirst}  # Added this line to ensure the first paragraph is indented for better readability
-  - \usepackage[margin=1in]{geometry}
----
-
-\pagebreak
-
-# List of Abbreviations
-\setlength{\parindent}{0em}
-**BHR**: Buy-and-Hold Returns   
-**CRSP**: Center for Research in Security Prices  
-**DS**: Datastream  
-**IDE**: Integrated Development Environment  
-**TREAT**: TRR 266 Template for Reproducible Empirical Accounting Research  
-**WRDS**: Wharton Research Data Services  
-**WSCP**: Worldscope
-
-\setlength{\parindent}{4em}
-
-\pagebreak
-
-# Introduction {#sec-introduction}
-This project serves as a template-based example for conducting and documenting reproducible empirical accounting research.
-It leverages the TRR 266 Template for Reproducible Empirical Accounting Research (TREAT) and builds upon the methodological foundations of the Corporate Decision-Making and Quantitative Analysis course, as well as the Accounting Reading Group course, which explored archival and non-archival research methods.
-To support future empirical thesis projects, this template was adapted and further developed specifically for students at the School of Business and Economics at Humboldt-Universität zu Berlin—particularly those affiliated with the Institute of Accounting and Auditing and the Finance Group.
-
-Following prior assignments on audit market concentration and financial reporting effects (e.g., @graham_economic_2005), this project focuses on the informativeness of quarterly earnings announcements.
-Specifically, it replicates and extends the study by @ball_how_2008, which examines whether earnings announcements contribute new information to markets or merely confirm what is already known—contributing to the long-standing discussion on market efficiency and the valuation versus contracting roles of accounting.
-
-Using updated U.S.
-data and applying the methodology to a second country, this thesis combines multiple WRDS datasets and applies an event-study design to evaluate how much of the annual stock return variation can be explained by earnings announcement windows.
-The approach is designed to foster transparent research workflows, with a fully scripted pipeline for data pulling, preparation, analysis, and reporting.
-
-The paper is organized into four main parts.
-In @sec-task1, I replicate the core findings of the original study using U.S.
-data from 1972 to 2006.
-In @sec-task2, I apply the same methodology to U.S.
-data from 2007 to 2023 to explore temporal robustness.
-In @sec-task3, I generalize the findings to a non-U.S.
-country to assess cross-country variation and data integration requirements.
-Finally, @sec-task4 sketches a non-archival research design using a survey-based approach to evaluate the underlying assumptions of the original paper.
-The thesis concludes with reflections and key insights in @sec-conclusion.
-
-# Task 1 - Replication of Key Tables and Figures {#sec-task1}
-Task 1 replicates key tables and figures from @ball_how_2008 using daily US stock return data (1972–2006) to establish baseline results.
-By reconstructing Table 1 (Panel A), Table 2, and Figure 1, it evaluates the robustness of the original findings and highlights any discrepancies.
-Daily CRSP data ensures accurate measurement of stock reactions around earnings announcements.
-Replication results are discussed in @sec-results1, with extended analysis in @sec-task2.
-
-## Research Design Choices and Assumptions {#sec-research_design_assumptions1}
-In line with @ball_how_2008, this replication examines U.S.
-earnings announcements and stock returns from 1972 to 2006, following the original study’s sample design, event window (-1 to +1), and return computations.
-Unlike the original, I omit analyst expectations, focusing solely on returns and announcement dates for simplicity.
-
-To address gaps in methodological detail, several assumptions are applied.
-Event windows follow CRSP trading calendar conventions, shifting dates when earnings fall on non-trading days [@center_for_research_in_security_prices_crsp_crsp_2024].
-While @ball_how_2008 require a minimum of 240 trading days, I relax this constraint, as per course guidance, resulting in a broader sample.
-All firms in Compustat meeting the quarterly announcement and return data criteria are included, regardless of activity status.
-I use PERMNO to uniquely track securities, avoiding aggregation across multiple share classes [@gow_empirical_2025].
-The CRSP/Compustat Merged (CCM) database is used to link identifiers, noting that differences from the original study may arise due to updated data (WRDS version: January 2025).
-As in @ball_how_2008, the analysis is return-based (RET in CRSP), excluding bid-ask pricing.
-Earnings are grouped by announcement date (RDQ), not fiscal period, to align directly with CRSP trading dates.
-
-These assumptions, together with the procedural details in @sec-replication_steps1, guide the replication of @ball_how_2008 and ensure transparency in design and implementation.
-
-## Replication Steps {#sec-replication_steps1}
-This section outlines the structured replication process: data pulling, preparation, and analysis, aligned with @ball_how_2008.
-
-**Step 1: Pulling the Data and Managing the Databases**  
-Data from WRDS combines daily CRSP stock returns (dsf) and Compustat quarterly earnings announcements (fundq), linked via the CRSP-Compustat link table (ccmxpf_lnkhist).
-The `permno` and `gvkey` identifiers ensure accurate firm-year alignment.
-Key variables are `ret` (returns) and `rdq` (announcement dates), pulled for the full 1972–2023 period used in Tasks 1 and 2.
-
-**Step 2: Data Preparation**  
-The merged dataset includes CRSP, Compustat, and the link table.
-After removing irrelevant linking variables, the dataset is trimmed to key variables: `ret`, `rdq`, `fyr`, `permno`, and `gvkey`.
-Only firms with four earnings announcements per calendar year are retained.
-Firms are also classified by fiscal year-end (`fyr`) to match Table 2 Panels C–D.
-Duplicate or incomplete firm-years are excluded.
-
-**Step 3: Analysis Implementation and Reproduction of Tables and Figure**  
-BHRs are computed using the cumulative return formula per @gundersen_returns_2022.
-Cross-sectional regressions are then run for each firm-year to replicate Table 1, Table 2, and Figure 1 from the original paper.
-
-## Results {#sec-results1}
-This section presents the replication results.
-
-# Task 2 - Extending the Analysis to 2007–2023 {#sec-task2}
-This section extends the analysis to 2007–2023, a period shaped by events like the 2008 financial crisis and the COVID-19 pandemic.
-Although the programming involves only a sample period update, the task requires thoughtful economic interpretation of how structural shifts and evolving market conditions may have affected the informativeness of earnings announcements.
-
-## Research Design Choices and Assumptions {#sec-research_design_assumptions2}
-
-Assumptions from @sec-research_design_assumptions1 still hold.
-
-## Replication Steps {#sec-replication_steps2}
-The pull, preparation, and analysis step would remain the same as designed in @sec-replication_steps1 with the only change in data range.
-
-## Results {#sec-results2}
-Shifting the sample to 2007–2023 introduces potential changes in earnings announcement informativeness driven by key economic and regulatory events.
-The 2008 financial crisis and the 2020 COVID-19 pandemic likely heightened market uncertainty, increasing reliance on earnings disclosures to assess firm stability.
-Moreover, Regulation (EU) No 537/2014 may have enhanced financial reporting transparency and audit quality in EU countries, potentially affecting how earnings news is reflected in stock prices and contributing to cross-country variation in informativeness.
-
-# Task 3 - Cross-Country Replication {#sec-task3}
-Task 3 generalizes the analysis to a non-U.S.
-market by replicating Figure 1 using 1972–2023 data from Canada.
-This involves aligning CRSP/Compustat variables with Worldscope and Datastream equivalents while accounting for differences in reporting standards, market liquidity, and institutional context.
-Though the core methodology is retained, adapting it to international data demands careful mapping and interpretation.
-
-## Research Design Choices and Assumptions {#sec-research_design_assumptions3}
-Since Task 3 applies a generalization and extension approach to a non-U.S.
-market, it requires adapting the methodology to alternative databases and market structures.
-Like in the original study, no analyst forecasts or earnings surprises are used—only stock returns and earnings announcement dates are required.
-Canada was selected for its comparable quarterly reporting frequency and institutional similarity to the U.S., making it a suitable choice for cross-country replication [@short_short_2025].
-Canadian firms reporting under U.S.
-GAAP further enhance comparability.
-
-To address database-specific issues not covered in @sec-research_design_assumptions1, I make the following adjustments: First, Worldscope/Datastream workflows are kept separate from CRSP/Compustat to prevent interference and facilitate debugging, given the scale of data involved.
-Second, I exclude fiscal-year-end-based subgroup analysis (Table 2 Panels C–D), as Figure 1 uses the full sample without splitting by fiscal period.
-Third, I use Worldscope items 5901–5904 to define earnings dates, though limited availability before 1992 may reduce the sample size [@thomson_financial_worldscope_2007, p.281].
-Fourth, differences in update frequency between quarterly Worldscope and weekly Datastream can lead to timing mismatches when aligning fundamentals and returns.
-Fifth, I rely on Datastream’s `ret` variable for percentage returns, ignoring bid/ask spreads.
-Finally, currency effects are ignored, as percentage returns are unaffected by currency denomination.
-
-## Replication Steps {#sec-replication_steps3}
-The process is same as in @sec-task1.
-
-**Step 1: Pulling the Data and Managing the Databases**  
-Following @dai_research_2021, I link Worldscope and Datastream using code and infocode.
-Since Worldscope’s year_ starts in 1980 [@wharton_research_data_services_worldscope_2025], Datastream data is also restricted to 1980–2023 to ensure consistent coverage.
-I extract quarterly earnings announcement dates (items 5901–5904) from Worldscope and daily stock returns from Datastream, focusing on Canadian firms and filtering for common equity (typecode = EQ).
-This step yields three datasets: 27.9M (Datastream), 77.9K (link table), and 1.94M (Worldscope) observations.
-
-**Step 2: Data Preparation**  
-Worldscope is first merged with the link table via code, producing 747,760 rows.
-This is then joined with Datastream via infocode, resulting in 10.2M observations.
-I exclude firms missing any earnings date fields, removing 3.6M rows, and drop 6M rows where earnings do not span all four quarters within the same calendar year.
-To retain more valid windows, event days (-1, 0, 1) are dynamically shifted within ±3 days when needed.
-After cleaning, I compute 3-day BHRs and extract full-year stock return data from Datastream.
-All retained infocode entries have around 250 trading days per year, confirming data completeness for annual BHR computation.
-
-**Step 3: Analysis Implementation and Reproduction of Tables and Figure**  
-The analysis replicates Table 1 (summary statistics), Table 2 (yearly regressions), and Figure 1 (Abnormal R² and slope trends).
-Using `BHR_3day` and `BHR_Annual`, I run yearly regressions of annual returns on earnings-window returns and compute Abnormal R² (adjusted R² – 4.8%), handling missing values appropriately.
-The resulting trends are plotted and saved as both PNG and pickle files for use in the paper.
-
-## Results {#sec-results3}
-Comparing the replication results with those from @ball_how_2008 reveals both similarities and discrepancies.
-The original figure shows a general upward trend in abnormal  R² , peaking in the early 2000s, whereas the replicated figure captures greater volatility in the later years (2010–2023).
-Additionally, Ball’s figure suggests a smoother long-term dynamic, while the replication exhibits more pronounced fluctuations, particularly post-2010.
-Panel B in replicated figure exhibits higher volatility and larger coefficient magnitudes, suggesting increased earnings informativeness post-2006, while Ball’s original figure has more stable and lower-magnitude coefficients.
-
-The post-2000 period in the replication shows more inconsistent abnormal  R^2  values, possibly reflecting evolving market structures, shifts in disclosure practices, and increased earnings informativeness following regulatory changes (e.g., IFRS adoption, post-SOX adjustments, different financial market structure).
-The pronounced fluctuations could also be driven by global financial crisis, Canada's higher market concentration, fewer publicly traded firms, and potential regulatory differences, which may affect earnings informativeness and stock return patterns.
-The variability in slope coefficients further suggests that earnings announcements’ impact on stock returns is less stable than in prior decades, potentially due to differences in investor response.
-
-# Task 4 - Research Design for a Non-Archival Study {#sec-task4}
-
-To extend the analysis beyond archival methods, this section proposes a survey to assess whether investors perceive earnings announcements as informative or merely confirmatory, following the triangulation principle in empirical research [@bloomfield_gathering_2016].
-While @ball_how_2008 measure market reactions via stock returns, surveys can capture investor beliefs directly, offering complementary evidence on earnings informativeness.
-
-A structured survey targeting 300 institutional and retail investors will be distributed through financial networks and platforms (e.g., LinkedIn, CFA societies), with a pilot study refining its clarity.
-Participants will receive a €20 honorarium, and responses will remain anonymous.
-The survey measures perceived informativeness (DV) using Likert-scale and multiple-choice questions, with IVs capturing factors such as investor preparedness and reaction timing.
-
-Follow-up interviews with selected respondents will add qualitative depth, enabling a richer understanding of investor perspectives.
-This mixed-method approach enhances validity and supports generalization of findings beyond price-based evidence.
-
-## Survey Questions
-
-1. **How do you typically prepare for earnings announcements in your investment decisions?**
-    * I conduct in-depth research and adjust my positions in advance 
-    * I monitor but rarely adjust positions pre-announcement
-    * I rely on market consensus, analyst forecasts, and AI-driven insights
-    * I do not make investment decisions based on earnings announcements
-
-2. **How do you typically react to earnings announcements?**
-    * I adjust my investment strategy immediately based on the announcement.
-    * I wait for further analysis before making changes.
-    * I rarely make investment decisions based on earnings announcements.
-
-3. **To what extent do you believe that earnings announcements provide new information beyond what is already reflected in market prices?**
-    * Always
-    * Most of the time
-    * Sometimes
-    * Rarely
-    * Never
-
-4. **Rank the following factors in order of importance when evaluating earnings announcements (1 = most important, 5 = least important):**
-    * Earnings per share (EPS) compared to analyst forecasts
-    * Revenue growth
-    * Management guidance and commentary
-    * Market reaction on the day of the announcement
-    * Industry trends and macroeconomic conditions
-
-5. **Do you consider earnings announcements to be more confirmatory or informative?**
-    * Primarily confirmatory (reinforce existing expectations)
-    * Primarily informative (provide new insights)
-    * A mix of both
-
-## Follow-Up Interview Questions
-
-1. Can you describe a recent instance where an earnings announcement significantly influenced your investment decision?
-2. In your experience, are there specific industries where earnings announcements are more informative than confirmatory?
-3. Do you use earnings announcements differently depending on market conditions (e.g., economic downturn vs. growth periods)?
-
-# Conclusion {#sec-conclusion}
-This thesis template demonstrates how the TRR 266 framework can be applied to structure a reproducible and transparent empirical accounting study.
-Using Quarto proved to be a particularly effective and user-friendly solution for structuring and rendering the thesis.
-It facilitates seamless integration of code, results, and narrative, making it an ideal tool for students at the Institute of Accounting and Auditing or the Finance Group at the School of Business and Economics, HU Berlin.
-
-Beyond replication, this project integrates programming, data processing, and visualization skills, offering a foundation for future extensions—whether through archival analysis or survey-based designs.
-As an open and modular template, this repository is designed to support ongoing student and academic work in empirical research.
-Thanks for reading!
-
-\pagebreak
-
-\setcounter{table}{0}
-\renewcommand{\thetable}{\arabic{table}}
-
-# References {-}
-\setlength{\parindent}{-0.2in}
-\setlength{\leftskip}{0.2in}
-\setlength{\parskip}{8pt}
-\noindent
+\clearpage

### 2025-05-04 doc/acknowledgements.tex
Added abstract, acknowledgments, list of abbreviations

--- 2025-05-04 doc/paper.qmd
@@
----
-title: |
-  | How Much New Information Is There in Earnings?
-  | Corporate Decision-Making and Quantitative Analysis
-  | Summer 2025 - Individual Report\vspace{1cm}
-author:
-  - name: Mel Mzv
-    email: m@student.hu-berlin.de
-    affiliations:
-      - Humboldt-Universität zu Berlin  
-date: today
-date-format: MMM D, YYYY [\vspace{0cm}] # Adjusted the vertical space after the date from 1cm
-abstract: |
-  | This project uses the TRR 266 Template for Reproducible Empirical Accounting Research (TREAT) to provide infrastructure for open science-oriented empirical projects.
-Leveraging data from the CRSP and Compustat databases, alongside Worldscope and Datastream via WRDS, this repository showcases a reproducible workflow integrating Python scripts for data preparation, analysis, and visualization.
-Integrating multiple databases adds complexity, requiring a detailed understanding of their structures and careful scripting to extract, align, and analyze data effectively.
-The project replicates and extends Ball and Shivakumar (2008) to analyze the informativeness of quarterly earnings announcements and their contribution to share price movements, highlighting their critical role in investment decisions and impact on investors and analysts.
-Key tasks include replicating and comparing original results, extending the analysis to 2007–2023, and applying the methodology to a non-U.S.
-country.
-The project also documents research design choices, discusses variations between original and reproduced results, and provides insights into earnings informativeness across different timeframes and jurisdictions.
-Additionally, it sketches a research design for a non-archival study to evaluate the paper’s findings.
-This code base, adapted from TREAT, demonstrates how the template applies to this project and serves as a structured guide for reproducible empirical research in accounting.
-  <!--
-  | \vspace{6cm} commenting away to decrease distance
-  -->
+\clearpage
+\pagenumbering{Roman}
+\setcounter{page}{2}  % Or adjust based on your layout
+\section*{Acknowledgements}
+\addcontentsline{toc}{section}{Acknowledgements}
 
-bibliography: references.bib
-biblio-style: apsr
-format:
-  pdf:
-    documentclass: article
-    number-sections: true
-    toc: true
-    lof: true  # Enables List of Figures
-    lot: true  # Enables List of Tables
-fig_caption: yes
-fontsize: 11pt
-ident: yes
-always_allow_html: yes
-number-sections: true 
-header-includes:
-  - \usepackage[nolists]{endfloat}    
-  - \usepackage{setspace}\doublespacing
-  - \setlength{\parindent}{4em}
-  - \setlength{\parskip}{0em}
-  - \usepackage[hang,flushmargin]{footmisc}
-  - \usepackage{caption} 
-  - \captionsetup[table]{skip=24pt,font=bf}
-  - \usepackage{array}
-  - \usepackage{threeparttable}
-  - \usepackage{adjustbox}
-  - \usepackage{graphicx}
-  - \usepackage{csquotes}
-  - \usepackage{indentfirst}  # Added this line to ensure the first paragraph is indented for better readability
-  - \usepackage[margin=1in]{geometry}
----
+I would like to thank my professors, colleagues, and friends for their support and insights throughout the preparation of this thesis.
+Special thanks go to [Name], whose guidance was invaluable during the project.
 
-\pagebreak
-
-# List of Abbreviations
-\setlength{\parindent}{0em}
-**BHR**: Buy-and-Hold Returns   
-**CRSP**: Center for Research in Security Prices  
-**DS**: Datastream  
-**IDE**: Integrated Development Environment  
-**TREAT**: TRR 266 Template for Reproducible Empirical Accounting Research  
-**WRDS**: Wharton Research Data Services  
-**WSCP**: Worldscope
-
-\setlength{\parindent}{4em}
-
-\pagebreak
-
-# Introduction {#sec-introduction}
-This project serves as a template-based example for conducting and documenting reproducible empirical accounting research.
-It leverages the TRR 266 Template for Reproducible Empirical Accounting Research (TREAT) and builds upon the methodological foundations of the Corporate Decision-Making and Quantitative Analysis course, as well as the Accounting Reading Group course, which explored archival and non-archival research methods.
-To support future empirical thesis projects, this template was adapted and further developed specifically for students at the School of Business and Economics at Humboldt-Universität zu Berlin—particularly those affiliated with the Institute of Accounting and Auditing and the Finance Group.
-
-Following prior assignments on audit market concentration and financial reporting effects (e.g., @graham_economic_2005), this project focuses on the informativeness of quarterly earnings announcements.
-Specifically, it replicates and extends the study by @ball_how_2008, which examines whether earnings announcements contribute new information to markets or merely confirm what is already known—contributing to the long-standing discussion on market efficiency and the valuation versus contracting roles of accounting.
-
-Using updated U.S.
-data and applying the methodology to a second country, this thesis combines multiple WRDS datasets and applies an event-study design to evaluate how much of the annual stock return variation can be explained by earnings announcement windows.
-The approach is designed to foster transparent research workflows, with a fully scripted pipeline for data pulling, preparation, analysis, and reporting.
-
-The paper is organized into four main parts.
-In @sec-task1, I replicate the core findings of the original study using U.S.
-data from 1972 to 2006.
-In @sec-task2, I apply the same methodology to U.S.
-data from 2007 to 2023 to explore temporal robustness.
-In @sec-task3, I generalize the findings to a non-U.S.
-country to assess cross-country variation and data integration requirements.
-Finally, @sec-task4 sketches a non-archival research design using a survey-based approach to evaluate the underlying assumptions of the original paper.
-The thesis concludes with reflections and key insights in @sec-conclusion.
-
-# Task 1 - Replication of Key Tables and Figures {#sec-task1}
-Task 1 replicates key tables and figures from @ball_how_2008 using daily US stock return data (1972–2006) to establish baseline results.
-By reconstructing Table 1 (Panel A), Table 2, and Figure 1, it evaluates the robustness of the original findings and highlights any discrepancies.
-Daily CRSP data ensures accurate measurement of stock reactions around earnings announcements.
-Replication results are discussed in @sec-results1, with extended analysis in @sec-task2.
-
-## Research Design Choices and Assumptions {#sec-research_design_assumptions1}
-In line with @ball_how_2008, this replication examines U.S.
-earnings announcements and stock returns from 1972 to 2006, following the original study’s sample design, event window (-1 to +1), and return computations.
-Unlike the original, I omit analyst expectations, focusing solely on returns and announcement dates for simplicity.
-
-To address gaps in methodological detail, several assumptions are applied.
-Event windows follow CRSP trading calendar conventions, shifting dates when earnings fall on non-trading days [@center_for_research_in_security_prices_crsp_crsp_2024].
-While @ball_how_2008 require a minimum of 240 trading days, I relax this constraint, as per course guidance, resulting in a broader sample.
-All firms in Compustat meeting the quarterly announcement and return data criteria are included, regardless of activity status.
-I use PERMNO to uniquely track securities, avoiding aggregation across multiple share classes [@gow_empirical_2025].
-The CRSP/Compustat Merged (CCM) database is used to link identifiers, noting that differences from the original study may arise due to updated data (WRDS version: January 2025).
-As in @ball_how_2008, the analysis is return-based (RET in CRSP), excluding bid-ask pricing.
-Earnings are grouped by announcement date (RDQ), not fiscal period, to align directly with CRSP trading dates.
-
-These assumptions, together with the procedural details in @sec-replication_steps1, guide the replication of @ball_how_2008 and ensure transparency in design and implementation.
-
-## Replication Steps {#sec-replication_steps1}
-This section outlines the structured replication process: data pulling, preparation, and analysis, aligned with @ball_how_2008.
-
-**Step 1: Pulling the Data and Managing the Databases**  
-Data from WRDS combines daily CRSP stock returns (dsf) and Compustat quarterly earnings announcements (fundq), linked via the CRSP-Compustat link table (ccmxpf_lnkhist).
-The `permno` and `gvkey` identifiers ensure accurate firm-year alignment.
-Key variables are `ret` (returns) and `rdq` (announcement dates), pulled for the full 1972–2023 period used in Tasks 1 and 2.
-
-**Step 2: Data Preparation**  
-The merged dataset includes CRSP, Compustat, and the link table.
-After removing irrelevant linking variables, the dataset is trimmed to key variables: `ret`, `rdq`, `fyr`, `permno`, and `gvkey`.
-Only firms with four earnings announcements per calendar year are retained.
-Firms are also classified by fiscal year-end (`fyr`) to match Table 2 Panels C–D.
-Duplicate or incomplete firm-years are excluded.
-
-**Step 3: Analysis Implementation and Reproduction of Tables and Figure**  
-BHRs are computed using the cumulative return formula per @gundersen_returns_2022.
-Cross-sectional regressions are then run for each firm-year to replicate Table 1, Table 2, and Figure 1 from the original paper.
-
-## Results {#sec-results1}
-This section presents the replication results.
-
-# Task 2 - Extending the Analysis to 2007–2023 {#sec-task2}
-This section extends the analysis to 2007–2023, a period shaped by events like the 2008 financial crisis and the COVID-19 pandemic.
-Although the programming involves only a sample period update, the task requires thoughtful economic interpretation of how structural shifts and evolving market conditions may have affected the informativeness of earnings announcements.
-
-## Research Design Choices and Assumptions {#sec-research_design_assumptions2}
-
-Assumptions from @sec-research_design_assumptions1 still hold.
-
-## Replication Steps {#sec-replication_steps2}
-The pull, preparation, and analysis step would remain the same as designed in @sec-replication_steps1 with the only change in data range.
-
-## Results {#sec-results2}
-Shifting the sample to 2007–2023 introduces potential changes in earnings announcement informativeness driven by key economic and regulatory events.
-The 2008 financial crisis and the 2020 COVID-19 pandemic likely heightened market uncertainty, increasing reliance on earnings disclosures to assess firm stability.
-Moreover, Regulation (EU) No 537/2014 may have enhanced financial reporting transparency and audit quality in EU countries, potentially affecting how earnings news is reflected in stock prices and contributing to cross-country variation in informativeness.
-
-# Task 3 - Cross-Country Replication {#sec-task3}
-Task 3 generalizes the analysis to a non-U.S.
-market by replicating Figure 1 using 1972–2023 data from Canada.
-This involves aligning CRSP/Compustat variables with Worldscope and Datastream equivalents while accounting for differences in reporting standards, market liquidity, and institutional context.
-Though the core methodology is retained, adapting it to international data demands careful mapping and interpretation.
-
-## Research Design Choices and Assumptions {#sec-research_design_assumptions3}
-Since Task 3 applies a generalization and extension approach to a non-U.S.
-market, it requires adapting the methodology to alternative databases and market structures.
-Like in the original study, no analyst forecasts or earnings surprises are used—only stock returns and earnings announcement dates are required.
-Canada was selected for its comparable quarterly reporting frequency and institutional similarity to the U.S., making it a suitable choice for cross-country replication [@short_short_2025].
-Canadian firms reporting under U.S.
-GAAP further enhance comparability.
-
-To address database-specific issues not covered in @sec-research_design_assumptions1, I make the following adjustments: First, Worldscope/Datastream workflows are kept separate from CRSP/Compustat to prevent interference and facilitate debugging, given the scale of data involved.
-Second, I exclude fiscal-year-end-based subgroup analysis (Table 2 Panels C–D), as Figure 1 uses the full sample without splitting by fiscal period.
-Third, I use Worldscope items 5901–5904 to define earnings dates, though limited availability before 1992 may reduce the sample size [@thomson_financial_worldscope_2007, p.281].
-Fourth, differences in update frequency between quarterly Worldscope and weekly Datastream can lead to timing mismatches when aligning fundamentals and returns.
-Fifth, I rely on Datastream’s `ret` variable for percentage returns, ignoring bid/ask spreads.
-Finally, currency effects are ignored, as percentage returns are unaffected by currency denomination.
-
-## Replication Steps {#sec-replication_steps3}
-The process is same as in @sec-task1.
-
-**Step 1: Pulling the Data and Managing the Databases**  
-Following @dai_research_2021, I link Worldscope and Datastream using code and infocode.
-Since Worldscope’s year_ starts in 1980 [@wharton_research_data_services_worldscope_2025], Datastream data is also restricted to 1980–2023 to ensure consistent coverage.
-I extract quarterly earnings announcement dates (items 5901–5904) from Worldscope and daily stock returns from Datastream, focusing on Canadian firms and filtering for common equity (typecode = EQ).
-This step yields three datasets: 27.9M (Datastream), 77.9K (link table), and 1.94M (Worldscope) observations.
-
-**Step 2: Data Preparation**  
-Worldscope is first merged with the link table via code, producing 747,760 rows.
-This is then joined with Datastream via infocode, resulting in 10.2M observations.
-I exclude firms missing any earnings date fields, removing 3.6M rows, and drop 6M rows where earnings do not span all four quarters within the same calendar year.
-To retain more valid windows, event days (-1, 0, 1) are dynamically shifted within ±3 days when needed.
-After cleaning, I compute 3-day BHRs and extract full-year stock return data from Datastream.
-All retained infocode entries have around 250 trading days per year, confirming data completeness for annual BHR computation.
-
-**Step 3: Analysis Implementation and Reproduction of Tables and Figure**  
-The analysis replicates Table 1 (summary statistics), Table 2 (yearly regressions), and Figure 1 (Abnormal R² and slope trends).
-Using `BHR_3day` and `BHR_Annual`, I run yearly regressions of annual returns on earnings-window returns and compute Abnormal R² (adjusted R² – 4.8%), handling missing values appropriately.
-The resulting trends are plotted and saved as both PNG and pickle files for use in the paper.
-
-## Results {#sec-results3}
-Comparing the replication results with those from @ball_how_2008 reveals both similarities and discrepancies.
-The original figure shows a general upward trend in abnormal  R² , peaking in the early 2000s, whereas the replicated figure captures greater volatility in the later years (2010–2023).
-Additionally, Ball’s figure suggests a smoother long-term dynamic, while the replication exhibits more pronounced fluctuations, particularly post-2010.
-Panel B in replicated figure exhibits higher volatility and larger coefficient magnitudes, suggesting increased earnings informativeness post-2006, while Ball’s original figure has more stable and lower-magnitude coefficients.
-
-The post-2000 period in the replication shows more inconsistent abnormal  R^2  values, possibly reflecting evolving market structures, shifts in disclosure practices, and increased earnings informativeness following regulatory changes (e.g., IFRS adoption, post-SOX adjustments, different financial market structure).
-The pronounced fluctuations could also be driven by global financial crisis, Canada's higher market concentration, fewer publicly traded firms, and potential regulatory differences, which may affect earnings informativeness and stock return patterns.
-The variability in slope coefficients further suggests that earnings announcements’ impact on stock returns is less stable than in prior decades, potentially due to differences in investor response.
-
-# Task 4 - Research Design for a Non-Archival Study {#sec-task4}
-
-To extend the analysis beyond archival methods, this section proposes a survey to assess whether investors perceive earnings announcements as informative or merely confirmatory, following the triangulation principle in empirical research [@bloomfield_gathering_2016].
-While @ball_how_2008 measure market reactions via stock returns, surveys can capture investor beliefs directly, offering complementary evidence on earnings informativeness.
-
-A structured survey targeting 300 institutional and retail investors will be distributed through financial networks and platforms (e.g., LinkedIn, CFA societies), with a pilot study refining its clarity.
-Participants will receive a €20 honorarium, and responses will remain anonymous.
-The survey measures perceived informativeness (DV) using Likert-scale and multiple-choice questions, with IVs capturing factors such as investor preparedness and reaction timing.
-
-Follow-up interviews with selected respondents will add qualitative depth, enabling a richer understanding of investor perspectives.
-This mixed-method approach enhances validity and supports generalization of findings beyond price-based evidence.
-
-## Survey Questions
-
-1. **How do you typically prepare for earnings announcements in your investment decisions?**
-    * I conduct in-depth research and adjust my positions in advance 
-    * I monitor but rarely adjust positions pre-announcement
-    * I rely on market consensus, analyst forecasts, and AI-driven insights
-    * I do not make investment decisions based on earnings announcements
-
-2. **How do you typically react to earnings announcements?**
-    * I adjust my investment strategy immediately based on the announcement.
-    * I wait for further analysis before making changes.
-    * I rarely make investment decisions based on earnings announcements.
-
-3. **To what extent do you believe that earnings announcements provide new information beyond what is already reflected in market prices?**
-    * Always
-    * Most of the time
-    * Sometimes
-    * Rarely
-    * Never
-
-4. **Rank the following factors in order of importance when evaluating earnings announcements (1 = most important, 5 = least important):**
-    * Earnings per share (EPS) compared to analyst forecasts
-    * Revenue growth
-    * Management guidance and commentary
-    * Market reaction on the day of the announcement
-    * Industry trends and macroeconomic conditions
-
-5. **Do you consider earnings announcements to be more confirmatory or informative?**
-    * Primarily confirmatory (reinforce existing expectations)
-    * Primarily informative (provide new insights)
-    * A mix of both
-
-## Follow-Up Interview Questions
-
-1. Can you describe a recent instance where an earnings announcement significantly influenced your investment decision?
-2. In your experience, are there specific industries where earnings announcements are more informative than confirmatory?
-3. Do you use earnings announcements differently depending on market conditions (e.g., economic downturn vs. growth periods)?
-
-# Conclusion {#sec-conclusion}
-This thesis template demonstrates how the TRR 266 framework can be applied to structure a reproducible and transparent empirical accounting study.
-Using Quarto proved to be a particularly effective and user-friendly solution for structuring and rendering the thesis.
-It facilitates seamless integration of code, results, and narrative, making it an ideal tool for students at the Institute of Accounting and Auditing or the Finance Group at the School of Business and Economics, HU Berlin.
-
-Beyond replication, this project integrates programming, data processing, and visualization skills, offering a foundation for future extensions—whether through archival analysis or survey-based designs.
-As an open and modular template, this repository is designed to support ongoing student and academic work in empirical research.
-Thanks for reading!
-
-\pagebreak
-
-\setcounter{table}{0}
-\renewcommand{\thetable}{\arabic{table}}
-
-# References {-}
-\setlength{\parindent}{-0.2in}
-\setlength{\leftskip}{0.2in}
-\setlength{\parskip}{8pt}
-\noindent
+\clearpage

### 2025-05-04 doc/paper.qmd
signature upd
Added abstract, acknowledgments, list of abbreviations

@@
     include-before-body: 
-      text: |
-        \pagenumbering{Roman}
+    - acknowledgements.tex
+    - abstract.tex
+    - abbreviations.tex
     include-after-body:
@@
 
-# List of Abbreviations
-\setlength{\parindent}{0em}
-**BHR**: Buy-and-Hold Returns   
-**CRSP**: Center for Research in Security Prices  
-**DS**: Datastream  
-**IDE**: Integrated Development Environment  
-**TREAT**: TRR 266 Template for Reproducible Empirical Accounting Research  
-**WRDS**: Wharton Research Data Services  
-**WSCP**: Worldscope
-
-\setlength{\parindent}{4em}
-
-\pagebreak
 \pagenumbering{arabic}

### 2025-05-04 doc/paper.qmd
Title page done

@@
 ---
-title: |
-  | How Much New Information Is There in Earnings?
-  | Corporate Decision-Making and Quantitative Analysis
-  | Summer 2025 - Individual Report\vspace{1cm}
-author:
-  - name: Max Mustermann
-    email: m@student.hu-berlin.de
-    affiliations:
-      - Humboldt-Universität zu Berlin  
-date: today
-date-format: MMM D, YYYY [\vspace{0cm}] # Adjusted the vertical space after the date from 1cm
-abstract: |
-  | This project uses the TRR 266 Template for Reproducible Empirical Accounting Research (TREAT) to provide infrastructure for open science-oriented empirical projects.
-Leveraging data from the CRSP and Compustat databases, alongside Worldscope and Datastream via WRDS, this repository showcases a reproducible workflow integrating Python scripts for data preparation, analysis, and visualization.
-Integrating multiple databases adds complexity, requiring a detailed understanding of their structures and careful scripting to extract, align, and analyze data effectively.
-The project replicates and extends Ball and Shivakumar (2008) to analyze the informativeness of quarterly earnings announcements and their contribution to share price movements, highlighting their critical role in investment decisions and impact on investors and analysts.
-Key tasks include replicating and comparing original results, extending the analysis to 2007–2023, and applying the methodology to a non-U.S.
-country.
-The project also documents research design choices, discusses variations between original and reproduced results, and provides insights into earnings informativeness across different timeframes and jurisdictions.
-Additionally, it sketches a research design for a non-archival study to evaluate the paper’s findings.
-This code base, adapted from TREAT, demonstrates how the template applies to this project and serves as a structured guide for reproducible empirical research in accounting.
-  <!--
-  | \vspace{6cm} commenting away to decrease distance
-  -->
-
-bibliography: references.bib
-biblio-style: apsr
 format:
@@
     toc: true
-    lof: true  # Enables List of Figures
-    lot: true  # Enables List of Tables
+    lof: true
+    lot: true
     include-before-body:
+      - titlepage.tex
       - acknowledgements.tex
@@
       - declaration.tex
+
+bibliography: references.bib
+biblio-style: apsr
 
@@
   - \usepackage{csquotes}
-  - \usepackage{indentfirst}  # Added this line to ensure the first paragraph is indented for better readability
+  - \usepackage{indentfirst}
   - \usepackage[margin=1in]{geometry}
@@
 Specifically, it replicates and extends the study by @ball_how_2008, which examines whether earnings announcements contribute new information to markets or merely confirm what is already known—contributing to the long-standing discussion on market efficiency and the valuation versus contracting roles of accounting.
-
-Using updated U.S.
-data and applying the methodology to a second country, this thesis combines multiple WRDS datasets and applies an event-study design to evaluate how much of the annual stock return variation can be explained by earnings announcement windows.
-The approach is designed to foster transparent research workflows, with a fully scripted pipeline for data pulling, preparation, analysis, and reporting.
 
@@
 All firms in Compustat meeting the quarterly announcement and return data criteria are included, regardless of activity status.
-I use PERMNO to uniquely track securities, avoiding aggregation across multiple share classes [@gow_empirical_2025].
 The CRSP/Compustat Merged (CCM) database is used to link identifiers, noting that differences from the original study may arise due to updated data (WRDS version: January 2025).

### 2025-05-04 doc/titlepage.tex
Title page done

\begin{titlepage}
    \begin{center}
    \vspace*{0.5cm}
    \LARGE{How Much New Information Is There in Earnings?}\\[3cm]
    \Large{Corporate Decision-Making and Quantitative Analysis}\\[1cm]
    \normalsize{
    Marketing Seminar\\
    Summer Term 2025 \\
    School of Business and Economics\\
    Humboldt University Berlin \\[3cm]
    }

    Submitted by:\\
    Max Mustermann (Matriculation no.: 123456)\\[3cm]

    Submitted to:\\
    Prof. Dr. Anastasia Danilov\\[2cm]

    Berlin, \today
    \end{center}
    \end{titlepage}

### 2025-05-05 doc/titlepage.tex
Finalized title page tex

@@
 \begin{titlepage}
-    \begin{center}
-    \vspace*{0.5cm}
-    \LARGE{How Much New Information Is There in Earnings?}\\[3cm]
-    \Large{Corporate Decision-Making and Quantitative Analysis}\\[1cm]
-    \normalsize{
-    Marketing Seminar\\
-    Summer Term 2025 \\
-    School of Business and Economics\\
-    Humboldt University Berlin \\[3cm]
+    \thispagestyle{empty}
+
+    \vspace*{1.5cm}
+
+    % Thesis title (2-line format)
+    {\Large
+    \noindent Analysis of the short-term economic effects\\
+    of the German subsidy for scrap materials}
+
+    \vspace{2cm}
+
+    % Type of thesis - largest font (not bold)
+    {\LARGE
+    \noindent
+    Master’s thesis}
+
+    \vspace{2cm}
+
+    % Degree info
+    {\normalsize
+    \noindent for acquiring the degree of\\
+    Master of Science (M.Sc.)
+
+    \vspace{0.5cm}
+
+    \noindent in Business Administration
+
+    \vspace{0.5cm}
+
+    \noindent at the School of Business and Economics\\
+    of Humboldt-Universität zu Berlin
+
+    \vspace{2cm}
+
+    % Submission info
+    \noindent submitted by\\
+    Max Mustermann\\
+    Student no. 1234567
+
+    \vspace{2cm}
+
+    \noindent First Examiner: << name of first examiner >>\\
+    Second Examiner: << name of second examiner >>
+
+    \vspace{1cm}
+
+    \noindent Berlin, \today
     }
 
-    Submitted by:\\
-    Max Mustermann (Matriculation no.: 123456)\\[3cm]
-
-    Submitted to:\\
-    Prof. Dr. Anastasia Danilov\\[2cm]
-
-    Berlin, \today
-    \end{center}
 \end{titlepage}

### 2025-05-05 doc/abstract.tex
Adjust abstract&titlepage to uni guidelines

@@
 
-This project uses the TRR 266 Template for Reproducible Empirical Accounting Research (TREAT) to provide infrastructure for open science-oriented empirical projects.
-Leveraging data from the CRSP and Compustat databases, alongside Worldscope and Datastream via WRDS, this repository showcases a reproducible workflow integrating Python scripts for data preparation, analysis, and visualization.
-Integrating multiple databases adds complexity, requiring a detailed understanding of their structures and careful scripting to extract, align, and analyze data effectively.
-
-The project replicates and extends Ball and Shivakumar (2008) to analyze the informativeness of quarterly earnings announcements and their contribution to share price movements, highlighting their critical role in investment decisions and impact on investors and analysts.
-Key tasks include replicating and comparing original results, extending the analysis to 2007–2023, and applying the methodology to a non-U.S.
-country.
-The project also documents research design choices, discusses variations between original and reproduced results, and provides insights into earnings informativeness across different timeframes and jurisdictions.
-Additionally, it sketches a research design for a non-archival study to evaluate the paper’s findings.
-This code base, adapted from TREAT, demonstrates how the template applies to this project and serves as a structured guide for reproducible empirical research in accounting.
+This thesis examines how much new information is conveyed in quarterly earnings announcements, replicating and extending the study by Ball and Shivakumar (2008).
+Using a reproducible workflow based on the TRR 266 Template, I analyze U.S.
+and Canadian data from 1972 to 2023, applying an event-study design to assess the explanatory power of earnings windows on annual stock returns.
+The findings show that earnings informativeness varies across time and jurisdictions, with evidence suggesting institutional factors influence market reactions.
+A non-archival survey design is also proposed to explore investor perceptions, enriching the archival analysis with complementary behavioral insights.
 

### 2025-05-05 doc/presentation.qmd
Neutralize Presentation Elements

@@
 title: |
-  | Earnings Management and Investor Protection:
-  | Accounting Reading Group
-author: "Mel Mzv"
+  | How Much New Information Is There in Earnings?
+  | Evidence from Earnings Releases in the U.S. and Canada
+author: "Max Mustermann"
 institute: "Humboldt-Universität zu Berlin – Summer 2025"

### 2025-05-05 doc/paper.qmd
Structure Aligned with Accounting Chair Requieremnts

@@
       - titlepage.tex
-      - acknowledgements.tex
-      - abstract.tex
-      - abbreviations.tex
+      - preamble.tex  # This sets Roman page numbering before the TOC
     include-after-body:
@@
 fig_caption: yes
-fontsize: 11pt
+fontsize: 12pt
 ident: yes
@@
   - \usepackage[nolists]{endfloat}
-  - \usepackage{setspace}\doublespacing
+  - \usepackage{setspace}\onehalfspacing
   - \setlength{\parindent}{4em}
@@
   - \usepackage[margin=1in]{geometry}
+  - \usepackage{ragged2e} #for declaration intend
 ---
@@
 
+\clearpage
+# Acknowledgements {-}
+I would like to thank my professors, colleagues, and friends for their support and insights throughout the preparation of this thesis.
+Special thanks go to [Name], whose guidance was invaluable during the project.
+
+\clearpage
+
+# Abstract {-}
+This thesis examines how much new information is conveyed in quarterly earnings announcements, replicating and extending the study by Ball and Shivakumar (2008).
+Using a reproducible workflow based on the TRR 266 Template, I analyze U.S.
+and Canadian data from 1972 to 2023, applying an event-study design to assess the explanatory power of earnings windows on annual stock returns.
+The findings show that earnings informativeness varies across time and jurisdictions, with evidence suggesting institutional factors influence market reactions.
+A non-archival survey design is also proposed to explore investor perceptions, enriching the archival analysis with complementary behavioral insights.
+
+\newpage
+
+# List of Abbreviations {-}
+\setlength{\parindent}{0em}
+**BHR**: Buy-and-Hold Returns   
+**CRSP**: Center for Research in Security Prices  
+**DS**: Datastream  
+**IDE**: Integrated Development Environment  
+**TREAT**: TRR 266 Template for Reproducible Empirical Accounting Research  
+**WRDS**: Wharton Research Data Services  
+**WSCP**: Worldscope
+
+\newpage
 \pagenumbering{arabic}
+\setlength{\parindent}{4em}
+\setlength{\parskip}{0em}
+\newgeometry{left=6cm, right=1.5cm, top=2cm, bottom=2cm}
 
@@
 
+\restoregeometry
 \setcounter{table}{0}

### 2025-05-05 doc/paper.qmd
Upload figure; Update qmd

@@
 
-# Task 1 - Replication of Key Tables and Figures {#sec-task1}
-Task 1 replicates key tables and figures from @ball_how_2008 using daily US stock return data (1972–2006) to establish baseline results.
+# Replication of Key Tables and Figures {#sec-task1}
+This section replicates key tables and figures from @ball_how_2008 using daily US stock return data (1972–2006) to establish baseline results.
 By reconstructing Table 1 (Panel A), Table 2, and Figure 1, it evaluates the robustness of the original findings and highlights any discrepancies.
@@
 The CRSP/Compustat Merged (CCM) database is used to link identifiers, noting that differences from the original study may arise due to updated data (WRDS version: January 2025).
-As in @ball_how_2008, the analysis is return-based (RET in CRSP), excluding bid-ask pricing.
-Earnings are grouped by announcement date (RDQ), not fiscal period, to align directly with CRSP trading dates.
+As in @ball_how_2008, the analysis is return-based (`ret` in CRSP), excluding bid-ask pricing.
+Earnings are grouped by announcement date (`rdq`), not fiscal period, to align directly with CRSP trading dates.
 
@@
 ## Results {#sec-results1}
-This section presents the replication results.
+This section presents the replication results, comparing them with the original @ by @ball_how_2008.
 
-# Task 2 - Extending the Analysis to 2007–2023 {#sec-task2}
+# Extending the Analysis to 2007–2023 {#sec-task2}
 This section extends the analysis to 2007–2023, a period shaped by events like the 2008 financial crisis and the COVID-19 pandemic.
-Although the programming involves only a sample period update, the task requires thoughtful economic interpretation of how structural shifts and evolving market conditions may have affected the informativeness of earnings announcements.
+Although the programming involves only a sample period update, the process requires thoughtful economic interpretation of how structural shifts and evolving market conditions may have affected the informativeness of earnings announcements.
 
@@
 
-# Task 3 - Cross-Country Replication {#sec-task3}
-Task 3 generalizes the analysis to a non-U.S.
+# Cross-Country Replication {#sec-task3}
+This section generalizes the analysis to a non-U.S.
 market by replicating Figure 1 using 1972–2023 data from Canada.
@@
 ## Research Design Choices and Assumptions {#sec-research_design_assumptions3}
-Since Task 3 applies a generalization and extension approach to a non-U.S.
+Since this secition applies a generalization and extension approach to a non-U.S.
 market, it requires adapting the methodology to alternative databases and market structures.
@@
 Following @dai_research_2021, I link Worldscope and Datastream using code and infocode.
-Since Worldscope’s year_ starts in 1980 [@wharton_research_data_services_worldscope_2025], Datastream data is also restricted to 1980–2023 to ensure consistent coverage.
+Since Worldscope’s `year_` starts in 1980 [@wharton_research_data_services_worldscope_2025], Datastream data is also restricted to 1980–2023 to ensure consistent coverage.
 I extract quarterly earnings announcement dates (items 5901–5904) from Worldscope and daily stock returns from Datastream, focusing on Canadian firms and filtering for common equity (typecode = EQ).
@@
 Using `BHR_3day` and `BHR_Annual`, I run yearly regressions of annual returns on earnings-window returns and compute Abnormal R² (adjusted R² – 4.8%), handling missing values appropriately.
-The resulting trends are plotted and saved as both PNG and pickle files for use in the paper.
 
@@
 
-The post-2000 period in the replication shows more inconsistent abnormal  R^2  values, possibly reflecting evolving market structures, shifts in disclosure practices, and increased earnings informativeness following regulatory changes (e.g., IFRS adoption, post-SOX adjustments, different financial market structure).
+The post-2000 period in the replication shows more inconsistent abnormal  R²  values, possibly reflecting evolving market structures, shifts in disclosure practices, and increased earnings informativeness following regulatory changes (e.g., IFRS adoption, post-SOX adjustments, different financial market structure).
 The pronounced fluctuations could also be driven by global financial crisis, Canada's higher market concentration, fewer publicly traded firms, and potential regulatory differences, which may affect earnings informativeness and stock return patterns.
@@
 
-# Task 4 - Research Design for a Non-Archival Study {#sec-task4}
+# Research Design for a Non-Archival Study {#sec-task4}
 
@@
 It facilitates seamless integration of code, results, and narrative, making it an ideal tool for students at the Institute of Accounting and Auditing or the Finance Group at the School of Business and Economics, HU Berlin.
-
-Beyond replication, this project integrates programming, data processing, and visualization skills, offering a foundation for future extensions—whether through archival analysis or survey-based designs.
-As an open and modular template, this repository is designed to support ongoing student and academic work in empirical research.
 Thanks for reading!

### 2025-05-05 doc/paper.qmd
Added a figure template

@@
 ## Results {#sec-results1}
-This section presents the replication results, comparing them with the original @ by @ball_how_2008.
+This section presents the replication results, comparing them with the original \ref{fig:fig1_original} by @ball_how_2008.
+
+```{=latex}
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=0.95\textwidth]{fig1_original.pdf}
+  \caption{Original Figure 1 from Ball and Shivakumar (2008)}
+  \label{fig:fig1_original}
+\end{figure}
+```
 

### 2025-05-05 doc/paper.qmd
Inserted formula

--- 2025-05-05 doc/paper.qmd
@@
 ## Results {#sec-results1}
-This section presents the replication results, comparing them with the original @ by @ball_how_2008.
+This section presents the replication results, comparing them with the original @fig-fig1_original by @ball_how_2008.
+
+```{r}
+#| label: fig-fig1_original
+#| fig-cap: "Original Figure 1 from Ball and Shivakumar (2008)"
+#| echo: false
+#| fig-align: center
+knitr::include_graphics("fig1_original.pdf")
+```
 
@@
 ## Results {#sec-results2}
-Shifting the sample to 2007–2023 introduces potential changes in earnings announcement informativeness driven by key economic and regulatory events.
+Shifting the sample to 2007–2023 introduces potential changes in earnings announcement informativeness driven by key economic events.
 The 2008 financial crisis and the 2020 COVID-19 pandemic likely heightened market uncertainty, increasing reliance on earnings disclosures to assess firm stability.
-Moreover, Regulation (EU) No 537/2014 may have enhanced financial reporting transparency and audit quality in EU countries, potentially affecting how earnings news is reflected in stock prices and contributing to cross-country variation in informativeness.
+These developments may have influenced how earnings news is reflected in stock prices and contributed to variation in informativeness across countries.
 
@@
 Using `BHR_3day` and `BHR_Annual`, I run yearly regressions of annual returns on earnings-window returns and compute Abnormal R² (adjusted R² – 4.8%), handling missing values appropriately.
+
+Specifically, buy-and-hold return over the three-day event window, that measures the stock’s reaction only to earnings news, is computed as given below:
+$$
+\begin{aligned}
+BHR_{\text{event}} &= (1 + R_{t_1}) (1 + R_{t_2}) \dots (1 + R_{t_T}) - 1,
+\\
+\text{where } R_t & \text{ is the daily return and } T \text{ is the total trading days in a year.}
+\end{aligned}
+$$
 

### 2025-05-05 doc/paper.qmd
Working w/ page numbering

@@
   - \usepackage{ragged2e} #for declaration intend
+  - \usepackage{fancyhdr}
+  - \pagestyle{fancy}
+  - \fancyhf{}  % clear headers/footers
+  - \renewcommand{\headrulewidth}{0pt}
+  - \renewcommand{\footrulewidth}{0pt}
+  - \fancyfoot[R]{\thepage}  % right-aligned footer
 ---

### 2025-05-05 doc/paper.qmd
right-aligned footer

--- 2025-05-05 doc/paper.qmd
@@
   - \usepackage{ragged2e} #for declaration intend
+  - \usepackage{fancyhdr}
+  - \pagestyle{fancy}
+  - \fancyhf{}  % clear headers/footers
+  - \renewcommand{\headrulewidth}{0pt}
+  - \renewcommand{\footrulewidth}{0pt}
+  - \fancyfoot[R]{\makebox[0pt][r]{\thepage\hspace{1in}}} % right-aligned footer
 ---

### 2025-05-06 doc/paper.qmd
Final paper refine-tuning

--- 2025-05-05 doc/paper.qmd
@@
 fig_caption: yes
-fontsize: 12pt
+fontsize: 12pt # 12-point font acc. to Accounting guidelines
 ident: yes
@@
   - \usepackage[nolists]{endfloat}
-  - \usepackage{setspace}\onehalfspacing
+  - \usepackage{setspace}\onehalfspacing # 1.5 line spacing acc. to Accounting guidelines
   - \setlength{\parindent}{4em}
@@
   - \usepackage{ragged2e} #for declaration intend
+  - \usepackage{fancyhdr}
+  - \pagestyle{fancy}
+  - \fancyhf{}  % clear headers/footers
+  - \renewcommand{\headrulewidth}{0pt}
+  - \renewcommand{\footrulewidth}{0pt}
+  - \fancyfoot[R]{\makebox[0pt][r]{\thepage\hspace{1in}}} % right-aligned footer # page numbering on lower right margin acc. to WiWi guidelines
+  - \setlength{\footskip}{0cm} # acc. to Wiwi guidelines - control a 1 to 1.5 cm gap below page numbers
 ---

### 2025-05-06 doc/paper.qmd
Updated list of abbreviations

@@
 \setlength{\parindent}{0em}
-**BHR**: Buy-and-Hold Returns   
-**CRSP**: Center for Research in Security Prices  
-**DS**: Datastream  
-**IDE**: Integrated Development Environment  
-**TREAT**: TRR 266 Template for Reproducible Empirical Accounting Research  
-**WRDS**: Wharton Research Data Services  
-**WSCP**: Worldscope
+\begin{tabular}{@{}ll}
+\textbf{BHR}   & Buy-and-Hold Returns \\
+\textbf{CRSP}  & Center for Research in Security Prices \\
+\textbf{DS}    & Datastream \\
+\textbf{IDE}   & Integrated Development Environment \\
+\textbf{TREAT} & TRR 266 Template for Reproducible Empirical Accounting Research \\
+\textbf{WRDS}  & Wharton Research Data Services \\
+\textbf{WSCP}  & Worldscope \\
+\end{tabular}
 

### 2025-05-06 doc/paper.qmd
Formatted equation with Quarto label for referencing

@@
 To support future empirical thesis projects, this template was adapted and further developed specifically for students at the School of Business and Economics at Humboldt-Universität zu Berlin—particularly those affiliated with the Institute of Accounting and Auditing and the Finance Group.
+The template was developed in adherence to the formal formatting and content guidelines of the School of Business and Economics, and in particular, those of the Institute of Accounting and Auditing.
 
-Following prior assignments on audit market concentration and financial reporting effects (e.g., @graham_economic_2005), this project focuses on the informativeness of quarterly earnings announcements.
+The template project showcases a research on audit market concentration and financial reporting effects (e.g., @graham_economic_2005), and focuses on the informativeness of quarterly earnings announcements.
 Specifically, it replicates and extends the study by @ball_how_2008, which examines whether earnings announcements contribute new information to markets or merely confirm what is already known—contributing to the long-standing discussion on market efficiency and the valuation versus contracting roles of accounting.
@@
 
-Specifically, buy-and-hold return over the three-day event window, that measures the stock’s reaction only to earnings news, is computed as given below:
+Specifically, buy-and-hold return over the three-day event window, that measures the stock’s reaction only to earnings news, is computed as shown in @eq-bhr_event:
 $$
@@
 \end{aligned}
-$$
+$$ {#eq-bhr_event}
 

### 2025-05-12 doc/paper_ResearchTopicsInFinance.qmd
Initial commit of academicWorkbench

---
title: |
  | AI-Enhanced Satellite Imagery in Institutional Lending Decisions: A Difference-in-Differences Study of Retail Firm Performance
  | Research Proposal 
  | Winter 2024/25 - Research Topics in Finance I\vspace{1cm}
author:
  - name: Melisa Mazaeva
    email: melisa.mazaeva@student.hu-berlin.de
    affiliations:
      - Humboldt-Universität zu Berlin  
date: today
date-format: MMM D, YYYY [\vspace{1cm}] # Adjusted the vertical space after the date from 1cm
abstract: |
  | This proposal explores whether advanced artificial intelligence techniques applied to satellite imagery can enhance retail firm performance measurement and influence institutional lending behavior. Building on Kang’s (2024) findings that satellite-derived car counts reduce lenders’ reliance on private information, this study introduces a richer performance metric using deep learning methods—including Convolutional Neural Networks for feature extraction, Faster Region-based Convolutional Neural Network for object detection, and Temporal Convolutional Networks for temporal analysis. The AI-enhanced metric is incorporated into a difference-in-differences framework to test whether superior predictive signals alter lenders’ behavior compared to traditional car count data. Syndicated loan records from DealScan (2020–2024) will be matched with Compustat/CRSP firm data and high-resolution satellite-derived parking lot metrics. Improved measures—such as vehicle turnover, density, and contextual indicators—are expected to reduce reliance on private information and lower institutional participation in loan syndicates. The findings are expected to inform institutional decision-making and regulatory standards, contributing to finance literature by reshaping information asymmetry in credit markets while addressing prior data limitations.
  | \vspace{6cm}

bibliography: references-ResearchTopicsInFinance.bib
biblio-style: apsr
format:
  pdf:
    documentclass: article
    number-sections: true
    toc: true
    lof: false  ## Enables List of Figures
    lot: false  ## Disables List of Tables
fig_caption: yes
fontsize: 11pt ##
ident: yes
always_allow_html: yes
number-sections: true 
header-includes:
  - \usepackage[nolists]{endfloat}    
  - \usepackage{setspace}\onehalfspacing ## adjust to \onehalfspacing if needed
  - \setlength{\parindent}{4em}
  - \setlength{\parskip}{0em}
  - \usepackage[hang,flushmargin]{footmisc}
  - \usepackage{caption} 
  - \captionsetup[table]{skip=24pt,font=bf}
  - \usepackage{array}
  - \usepackage{threeparttable}
  - \usepackage{adjustbox}
  - \usepackage{graphicx}
  - \usepackage{csquotes}
  - \usepackage{indentfirst}  # Added this line to ensure the first paragraph is indented for better readability
  - \usepackage[margin=1in]{geometry}
---

\pagebreak

# List of Abbreviations
\setlength{\parindent}{0em}
**AI**: Artificial Intelligence  
**CNN**: Convolutional Neural Network  
**CRSP**: Center for Research in Security Prices    
**DiD**: Difference-in-Differences  
**DV**: Dependent Variable  
**Faster R-CNN**: Faster Region-based Convolutional Neural Network  
**IV**: Independent Variable  
**TCN**: Temporal Convolutional Network  
**US**: United States  
**WRDS**: Wharton Research Data Services

\setlength{\parindent}{4em}

\pagebreak

# Introduction {#sec-introduction}
Institutional lenders are known to gain an information advantage by accessing borrowers’ private data through lending relationships and using it for informed trading [@ivashina_institutional_2011]. However, the rise of alternative data, particularly satellite imagery, can offer real-time insights into firm performance that can erode this advantage [@kang_gone_2024]. @kang_gone_2024 provided evidence that once daily car-count data from satellite images became available for retail firms, institutional investors were less inclined to participate in those firms’ syndicated loans. The intuition is that accurate, timely public signals of performance undermine the exclusive value of private information. The main research question addressed in this study is: If institutional lenders have access to AI-enhanced performance signals from satellite imagery, do they further reduce their reliance on private information (and thus participation in loans) compared to the era of simpler satellite metrics? This research question is important because it investigates how advances in Artificial Intelligence (AI) can fundamentally transform the competitive dynamics of credit markets, making public alternative data nearly as informative as proprietary private data, which has significant implications for regulatory policy and lending practices.

Advancements in computer vision now allow a reexamination of satellite imagery. Whereas traditional methods like Orbital Insight’s car-count algorithm used by @kang_gone_2024 provide useful but limited measures (e.g. excluding covered parking), modern AI models can overcome these gaps. Convolutional Neural Networks (CNNs) can extract high-level features from raw images [@krizhevsky_imagenet_2017]; Faster Region-based Convolutional Neural Network (Faster R-CNN) improves vehicle detection and context [@ren_faster_2017]; and Temporal Convolutional Networks (TCNs) capture temporal patterns to yield a more stable performance indicator [@lea_temporal_2017]. This research examines whether refined public signals from AI-enhanced imagery further reduce institutional lenders’ reliance on private information. The structure of this proposal is as follows: the next section reviews related literature, followed by details on data and methodology, and concludes with anticipated results and contributions.

# Related Literature {#sec-relatedlit}
This study builds on and extends existing research on alternative data in capital markets and institutional lending by exploring the evolving role of AI-enhanced satellite imagery in reducing information asymmetry. Prior work indicates that the introduction of satellite data coverage for major retailers enables sophisticated investors to design profitable trading strategies [@katona_capital_2024]. @feng_empirical_2022 further find that parking lot traffic predicts forward-looking performance (Tobin’s Q) for retailers, underlining the value of such data even for future expectations. @kang_gone_2024 showed that the availability of daily satellite data reduced institutional participation in loans, suggesting that robust public signals can diminish the need for private data. However, these basic metrics miss nuances such as covered parking and the distinction between short- and long-term trends. It is unclear whether advanced AI techniques can yield more informative public signals. This study employs sophisticated AI to extract richer features, positing that these enhanced signals will provide deeper insights into firm performance and more strongly influence investor behavior.

Based on these insights, the primary hypothesis is that the integration of AI-enhanced performance signals from satellite imagery will lead to a further reduction in institutional lenders’ participation in syndicated loans relative to traditional car count measures, as more precise signals reduce the marginal benefit of acquiring private information for informed trading.

# Data {#sec-data}
Data will focus on U.S. retail firms during 2020–2024—a period when deep learning for image analysis has matured, following the commercialization of car-count data since 2015 [@kang_gone_2024]. High-resolution raw satellite images from providers such as Planet Labs or Maxar are expected to yield tens of thousands of daily observations across roughly 500 firms. These images will be processed using a fine-tuned Faster R-CNN for object detection and a TCN to capture temporal variations, yielding a multidimensional performance index that overcomes limitations outlined by @kang_gone_2024 (e.g., exclusion of covered parking lots).

The AI-derived metrics will be merged with syndicated loan data from DealScan and firm financials from Compustat and Center for Research in Security Prices (CRSP) using common identifiers; all three datasets are available through Wharton Research Data Services (WRDS) [@wharton_research_data_services_get_2025]. Key variables include vehicle count, parking lot utilization rate, turnover (entries/exits), and spatial distribution of vehicles, which will be aggregated into a performance index. The study will use the same syndicated loan data from DealScan as @kang_gone_2024 and match loan issuance details with firm financial data from Compustat and CRSP. The treatment group consists of retail firms with AI-enhanced metrics from 2020 onward, while control firms are the ones that did not experience significant improvements in data measurement, with 2015–2019 as the baseline and 2020–2024 as the post-treatment period. A difference-in-differences (DiD) approach will be employed to isolate the effect of AI-enhanced signals on lending behavior. Data limitations—such as missing imagery due to poor weather or resolution variations—will be addressed via cleaning, imputation, and robustness checks. With unlimited funds, the cost of these datasets is not a constraint, and the data are readily available.

# Methodology {#sec-methodology}
The empirical strategy employs a DiD design to compare changes in institutional lending behavior for treated versus control firms before and after the introduction of AI-enhanced performance metrics. Conceptually, Kang (2024) implemented a DiD where the “treatment” was the introduction of basic satellite car-count coverage. This study maintains a similar framework but redefines the treatment in terms of AI-enhanced performance signals. An indicator variable (`AI_Treat_i`) is defined for each firm such that it equals 1 if the firm is measured using AI-enhanced satellite metrics, and 0 otherwise. A post-period indicator (`Post_t`) is set to 1 for observations from 2020 to 2024 and 0 for those from 2015 to 2019. The interaction term (`AI_Treat_i × Post_t`) thus captures the effect of the enhanced performance signal on institutional lending outcomes.

The baseline DiD regression model is specified as follows:
$$
\text{InstPart}_{i,t} = \alpha + \beta_1 \, (\text{AI\_Treat}_i \times \text{Post}_t) + \beta_2 \, \text{AI\_Treat}_i + \beta_3 \, \text{Post}_t + \gamma X_{i,t} + \varepsilon_{i,t},
$$
where $\text{InstPart}_{i,t}$ represents the measure of institutional lender participation (e.g., a dummy for institutional presence or the percentage share of the loan) and serves as the dependent variable (DV). The key independent variable (IV) is the interaction term $\text{AI\_Treat}_{i} \times \text{Post}_{t}$, which measures the impact of AI-enhanced performance signals on lending behavior. Additional IVs include the binary treatment indicator $\text{AI\_Treat}_{i}$, the post-period indicator $\text{Post}_{t}$, and control variables $X_{i,t}$ such as firm size, credit rating, and loan maturity. Fixed effects at the industry×time level will be incorporated to account for broader trends. A significantly negative $\beta_1$ is anticipated, indicating that improved public signals reduce the need for private information.

The AI pipeline involves several deep learning models to construct the enhanced performance index. First, Convolutional Neural Networks (e.g., ResNet) will extract features from raw satellite images. Next, Faster R-CNN will be applied for precise vehicle detection and to capture spatial arrangements, hence, overcoming limitations such as the exclusion of covered parking data as outlined by @kang_gone_2024. Finally, TCNs will aggregate daily observations into a smooth, temporally robust index. This AI-derived performance index will then be merged with syndicated loan data from DealScan and firm financials from Compustat and CRSP using common identifiers.

The AI component of the methodology is implemented in Python using libraries such as PyTorch and torchvision. Initially, satellite images are pre-processed and converted into numerical arrays that the deep learning model can work with efficiently. A pre-trained Faster R-CNN model (specifically, fasterrcnn_resnet50_fpn) is then used to detect vehicles within these images, thereby estimating the number of cars present. The detected counts are aggregated into daily metrics, which are subsequently fed into a TCN to capture temporal patterns such as seasonal trends and fluctuations. The resulting time series signals—representing refined measures of retail firm performance, such as weekly average parking lot occupancy or turnover rates—serve as the key independent variables in the difference-in-differences regression analysis.

Robustness checks will test the parallel trends assumption by examining pre-2020 lending patterns for both groups, and additional falsification tests (e.g., using a false treatment date) will be conducted to ensure that the observed effects are not driven by other unrelated factors.

# Proposed Contribution {#sec-contribution}
This study offers several contributions. First, it advances the literature on alternative data by demonstrating how AI-driven methods can extract a richer, multidimensional signal from satellite imagery. While @kang_gone_2024 relied on basic car counts—which had limitations such as exclusion of images from covered parking lots and limited contextual detail—the proposed AI approach captures additional dimensions like spatial distribution, vehicle turnover, and temporal patterns. Second, by integrating this refined metric with established syndicated loan data within a DiD framework, the study contributes to the understanding of how improved public information affects institutional lending behavior. The results are anticipated to provide policy-relevant insights on the shifting balance between public and private information in credit markets. Finally, from a methodological perspective, the study presents a blueprint for incorporating innovative AI techniques into empirical finance research, offering a model that can be adapted for other applications.

# Anticipated Results {#sec-results}
It is expected that institutional lenders will reduce their participation in syndicated loans for firms with AI-enhanced performance metrics relative to those measured by traditional car counts [@kang_gone_2024]. In a DiD framework, a negative and significant coefficient on the interaction term (`AI_Treat × Post`) would indicate that superior public signals lessen the need for private information, with a larger effect than observed with simple car counts. Treated firms are also anticipated to face less favorable loan terms (e.g., higher spreads, stricter covenants). Robustness checks will assess whether the effect is especially strong for opaque firms, and the study will verify that the AI-driven index more accurately predicts firm performance (e.g., sales growth) than traditional metrics.

Following approval and completion, the findings will be shared in academic journals and policy briefings to influence regulatory frameworks and lending practices. The results may inform policymakers that advanced public data can substitute for private information, reshaping credit market transparency. Study limitations—such as the reliance on high-quality satellite imagery and difficulty isolating AI effects—will be critically examined. Future research may expand to other geographic regions to examine whether similar AI-enhanced metrics can improve public signals of firm performance and influence lending behavior in diverse market environments, thereby shedding light on the broader applicability of these methods in different regulatory and economic contexts.

\pagebreak

\setcounter{table}{0}
\renewcommand{\thetable}{\arabic{table}}

# References {-}
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\noindent

### 2025-05-20 code/python/do_analysis.py
pulled & prepared data

@@ -1,96 +1,73 @@
 import os
-import pickle
 import pandas as pd
 import matplotlib.pyplot as plt
-import statsmodels.api as sm
+import seaborn as sns
+import pickle
 from utils import read_config, setup_logging
 
 log = setup_logging()
 
 
 def main():
-    log.info("Running post-preparation analysis...")
-
     cfg = read_config("config/do_analysis_cfg.yaml")
 
     # === Load prepared data ===
-    df_event = pd.read_csv(cfg["bhr_event_output_csv"])
-    df_annual = pd.read_csv(cfg["bhr_annual_output_csv"])
+    log.info("Loading prepared Gapminder 2007 data ...")
+    df = pd.read_parquet(cfg["gapminder_2007_input"])
 
-    # === Summary Statistics ===
-    summary = compute_summary(df_annual, df_event)
+    # === Summary statistics ===
+    log.info("Computing summary statistics by continent ...")
+    summary = compute_summary(df)
     summary.to_csv(cfg["summary_statistics_csv"], index=False)
+    log.info("Summary statistics saved.")
 
-    # === Regressions & Plot ===
-    results = run_regressions(df_annual, df_event)
-    results.to_csv(cfg["regression_results_csv"], index=False)
-    plot_figure(results, cfg["figure1_save_path"], cfg["figure1_pickle_path"])
-
-    log.info("Analysis complete.")
-
-
-def compute_summary(df_annual, df_event):
-    """Basic summary stats for annual and event-window returns."""
-    def stats(series):  # Helper
-        return {
-            "Mean": series.mean(),
-            "Median": series.median(),
-            "Skew": series.skew(),
-            "% = 0": (series == 0).mean() * 100,
-            "% > 0": (series > 0).mean() * 100,
-            "Obs": len(series)
-        }
-
-    summary = [{"Category": "Annual", **stats(df_annual["BHR_Annual"])}]
-    for q in ["Q1", "Q2", "Q3", "Q4"]:
-        subset = df_event[df_event["quarter"] == q]["BHR_3day"]
-        summary.append({"Category": f"3-day {q}", **stats(subset)})
-
-    return pd.DataFrame(summary)
-
-
-def run_regressions(df_annual, df_event):
-    """Run annual regressions of annual returns on quarterly 3-day returns."""
[…]

### 2025-05-20 code/python/prepare_data.py

@@ -1,99 +1,26 @@
-# --- Header -------------------------------------------------------------------
-# Prepare the pulled data for further analysis as per Analysis requirements
-#
-# (C) Mel Mzv - See LICENSE file for details
-# ------------------------------------------------------------------------------
+import pandas as pd
+import os
+from utils import read_config, setup_logging
+
+log = setup_logging()
 
 def main():
-    log.info("Preparing data ...")
     cfg = read_config('config/prepare_data_cfg.yaml')
 
-    ws = pd.read_csv(cfg['worldscope_sample_save_path_csv'])
-    link = pd.read_csv(cfg['link_ds_ws_save_path_csv'])
-    ds = pd.read_csv(cfg['datastream_sample_save_path_csv'])
+    log.info("Loading raw Gapminder 2007 data ...")
+    df = pd.read_parquet(cfg['input_path'])
 
-    merged = merge_worldscope_link(ws, link)
-    long = pivot_longer_earnings(merged)
-    events = expand_event_window(long)
-    with_ret = merge_with_datastream(events, ds)
-    sample = select_firms_for_sample(with_ret)
-    bhr_event = compute_and_save_eawr_bhr(sample, cfg)
-    annual_data = extract_annual_stock_data(bhr_event, ds, cfg)
-    bhr_annual = compute_and_save_annual_bhr(cfg)
+    log.info("Checking for missing values ...")
+    missing = df.isnull().sum()
+    log.info(f"Missing values per column:\n{missing}")
 
-    sample.to_csv(cfg['prepared_wrds_ds2dsf_path'], index=False)
-    sample.to_parquet(cfg['prepared_wrds_ds2dsf_parquet'], index=False)
-    log.info("Done.")
+    log.info("Rounding numeric columns to 0 decimal places ...")
+    numeric_cols = df.select_dtypes(include='number').columns
+    df[numeric_cols] = df[numeric_cols].round(0)
 
-def merge_worldscope_link(ws, link):
-    df = ws.merge(link, on="code", how="inner")
-    return df[["year_", "item6105", "item5901", "item5902", "item5903", "item5904", "infocode"]]
+    os.makedirs(os.path.dirname(cfg['output_path']), exist_ok=True)
+    df.to_parquet(cfg['output_path'])
[…]

### 2025-05-20 code/python/pull_data.py

@@ -1,0 +1,27 @@
+# --- Header -------------------------------------------------------------------
+# See LICENSE file for details
+#
+# This code pulls 2007 data from the Gapminder Database
+# ------------------------------------------------------------------------------
+
+import pandas as pd
+from utils import read_config, setup_logging
+from gapminder import gapminder
+
+log = setup_logging()
+
+def main():
+    cfg = read_config('config/pull_data_cfg.yaml')
+
+    # 1. Pull
+    log.info("Loading Gapminder dataset from package ...")
+    df = gapminder.copy()
+    log.info("Done loading.")
+
+    # 2. Subset for 2007 only
+    df_2007 = df[df['year'] == 2007].copy()
+    df_2007.to_parquet(cfg['save_path_2007_parquet'])
+    log.info("Saved 2007 subset as Parquet data.")
+
+if __name__ == '__main__':
+    main()

### 2025-05-20 code/python/pull_wrds_data.py

@@ -1,101 +1,0 @@
-# --- Header -------------------------------------------------------------------
-# See LICENSE file for details
-#
-# This code pulls data from WRDS Databases 
-# ------------------------------------------------------------------------------
-
-import os
-from getpass import getpass
-import dotenv
-
-import pandas as pd
-from utils import read_config, setup_logging
-import wrds
-
-log = setup_logging()
-
-def main():
-    '''
-    Main function to pull data from WRDS.
-
-    This function reads the configuration file, gets the WRDS login credentials, and pulls the data from WRDS.
-
-    The data is then saved to CSV and Parquet files.
-    '''
-    cfg = read_config('config/pull_data_cfg.yaml')
-    wrds_login = get_wrds_login()
-    
-    # Pull CRSP and Compustat Data
-    db = wrds.Connection(
-        wrds_username=wrds_login['wrds_username'], 
-        wrds_password=wrds_login['wrds_password']
-    )
-    
-    log.info('Logged on to WRDS ...')
-    
-    pull_crsp_data(cfg, db)
-    pull_compustat_data(cfg, db)
-    pull_link_data(db)
-    
-    db.close()
-    log.info("Disconnected from WRDS")
-
-def get_wrds_login():
-    '''
-    Gets the WRDS login credentials.
-    '''
-    if os.path.exists('secrets.env'):
-        dotenv.load_dotenv('secrets.env')
-        wrds_username = os.getenv('WRDS_USERNAME')
-        wrds_password = os.getenv('WRDS_PASSWORD')
-        return {'wrds_username': wrds_username, 'wrds_password': wrds_password}
-    else:
-        wrds_username = input('Please provide a WRDS username: ')
-        wrds_password = getpass(
-            'Please provide a WRDS password (it will not show as you type): ')
-        return {'wrds_username': wrds_username, 'wrds_password': wrds_password}
-
-def pull_link_data(db):
-    """
-    Pulls the linking table between Compustat and CRSP from WRDS.
-    """
-    linkdata_df_wrds = db.get_table(library="crsp_a_ccm", table="ccmxpf_linktable")
-    linkdata_df_wrds.to_parquet("data/pulled/linkdata_compustat_crsp.parquet")
[…]

### 2025-05-20 code/python/test_wrds_connection.py

@@ -1,32 +1,0 @@
-import os
-
-
-def load_wrds_credentials():
-    '''
-    Load WRDS credentials from secrets.env file.
-    '''
-    if os.path.exists('secrets.env'):
-        dotenv.load_dotenv('secrets.env')
-        wrds_username = os.getenv('WRDS_USERNAME')
-        wrds_password = os.getenv('WRDS_PASSWORD')
-        return {'wrds_username': wrds_username, 'wrds_password': wrds_password}
-    else:
-        raise FileNotFoundError("secrets.env file not found")
-
-def test_wrds_connection():
-    '''
-    Test WRDS connection with credentials from secrets.env.
-    '''
-    credentials = load_wrds_credentials()
-    db = wrds.Connection(
-        wrds_username=credentials['wrds_username'],
-        wrds_password=credentials['wrds_password']
-    )
-    print("Connection to WRDS successful!")
-    db.close()
-
-if __name__ == '__main__':
-    try:
-        test_wrds_connection()
-    except Exception as e:
-        print(f"Error connecting to WRDS: {e}")

### 2025-05-20 doc/paper.qmd
pulled & prepared data
analysis done; qmd building

@@
 # Abstract {-}
-This thesis examines how much new information is conveyed in quarterly earnings announcements, replicating and extending the study by Ball and Shivakumar (2008).
-Using a reproducible workflow based on the TRR 266 Template, I analyze U.S.
-and Canadian data from 1972 to 2023, applying an event-study design to assess the explanatory power of earnings windows on annual stock returns.
-The findings show that earnings informativeness varies across time and jurisdictions, with evidence suggesting institutional factors influence market reactions.
-A non-archival survey design is also proposed to explore investor perceptions, enriching the archival analysis with complementary behavioral insights.
+This thesis template explores global development patterns using the publicly available Gapminder dataset, with a specific focus on the year 2007.
+Leveraging a reproducible workflow adapted from the TRR 266 Template, the project presents descriptive statistics by continent and a visual analysis of the relationship between life expectancy, income per capita, and population.
+The findings reveal clear disparities in health and wealth outcomes across world regions.
+A comparison with Gapminder’s 2021 global development chart highlights persistent global inequalities, while illustrating notable progress in certain regions.
+The analysis serves as a modern, transparent example of public data storytelling in empirical research.
 
@@
 \begin{tabular}{@{}ll}
-\textbf{BHR}   & Buy-and-Hold Returns \\
-\textbf{CRSP}  & Center for Research in Security Prices \\
-\textbf{DS}    & Datastream \\
+\textbf{GDP}   & Gross Domestic Product \\
 \textbf{IDE}   & Integrated Development Environment \\
 \textbf{TREAT} & TRR 266 Template for Reproducible Empirical Accounting Research \\
-\textbf{WRDS}  & Wharton Research Data Services \\
-\textbf{WSCP}  & Worldscope \\
 \end{tabular}
@@
 # Introduction {#sec-introduction}
-This project serves as a template-based example for conducting and documenting reproducible empirical accounting research.
-It leverages the TRR 266 Template for Reproducible Empirical Accounting Research (TREAT) and builds upon the methodological foundations of the Corporate Decision-Making and Quantitative Analysis course, as well as the Accounting Reading Group course, which explored archival and non-archival research methods.
-To support future empirical thesis projects, this template was adapted and further developed specifically for students at the School of Business and Economics at Humboldt-Universität zu Berlin—particularly those affiliated with the Institute of Accounting and Auditing and the Finance Group.
+This project serves as a template-based example for conducting and documenting reproducible empirical data analysis using publicly available datasets.
+It leverages the TRR 266 Template for Reproducible Empirical Accounting Research (TREAT) and builds upon the methodological foundations of the Corporate Decision-Making and Quantitative Analysis, as well as the Accounting Reading Group course.
+To support future empirical thesis projects, this template was adapted and developed specifically for students at the School of Business and Economics at Humboldt-Universität zu Berlin—particularly those affiliated with the Institute of Accounting and Auditing and the Finance Group.
 The template was developed in adherence to the formal formatting and content guidelines of the School of Business and Economics, and in particular, those of the Institute of Accounting and Auditing.
+It is not a completed thesis but a demonstrative project hosted in the [template repository](https://github.com/melmzv/huwiwiQuartoThesis).
 
-The template project showcases a research on audit market concentration and financial reporting effects (e.g., @graham_economic_2005), and focuses on the informativeness of quarterly earnings announcements.
-Specifically, it replicates and extends the study by @ball_how_2008, which examines whether earnings announcements contribute new information to markets or merely confirm what is already known—contributing to the long-standing discussion on market efficiency and the valuation versus contracting roles of accounting.
+This template project explores global development patterns using the Gapminder dataset, focusing on the year 2007.
+The project illustrates how reproducible research workflows can be applied beyond traditional financial datasets by using public and open-source data.
+Specifically, it presents a summary statistics table by continent and visualizes the relationship between GDP per capita and life expectancy, offering a static snapshot of world development during 2007.
+These descriptive results are presented in @sec-task1.
 
-The paper is organized into four main parts.
-In @sec-task1, I replicate the core findings of the original study using U.S.
-data from 1972 to 2006.
-In @sec-task2, I apply the same methodology to U.S.
-data from 2007 to 2023 to explore temporal robustness.
-In @sec-task3, I generalize the findings to a non-U.S.
-country to assess cross-country variation and data integration requirements.
-Finally, @sec-task4 sketches a non-archival research design using a survey-based approach to evaluate the underlying assumptions of the original paper.
-The thesis concludes with reflections and key insights in @sec-conclusion.
+The results are further contrasted with Gapminder’s interactive global development visualizations as of 2021 [@gapminder_health_chart], highlighting the potential and limits of static data analysis.
+This comparative discussion is provided in @sec-task2.
+The project concludes with reflections on transparency, reproducibility, and the importance of making research accessible to a broader audience, as discussed in @sec-conclusion.
 
@@
 ## Results {#sec-results1}
-This section presents the replication results, comparing them with the original @fig-fig1_original by @ball_how_2008.
+This section compares the static 2007 snapshot with Gapminder’s dynamic global development chart @fig-gapminder_original [@gapminder_health_chart].
 
 ```{r}
-#| label: fig-fig1_original
-#| fig-cap: "Original Figure 1 from Ball and Shivakumar (2008)"
+#| label: fig-gapminder_original
+#| fig-cap: "Original World Health Chart 2021 from Gapminder Foundation"
 #| echo: false
 #| fig-align: center
-knitr::include_graphics("fig1_original.pdf")
+knitr::include_graphics("countries_health_wealth_2021.pdf")
 ```
 
-# Extending the Analysis to 2007–2023 {#sec-task2}
-This section extends the analysis to 2007–2023, a period shaped by events like the 2008 financial crisis and the COVID-19 pandemic.
-Although the programming involves only a sample period update, the process requires thoughtful economic interpretation of how structural shifts and evolving market conditions may have affected the informativeness of earnings announcements.
-
-## Research Design Choices and Assumptions {#sec-research_design_assumptions2}
-
-Assumptions from @sec-research_design_assumptions1 still hold.
-
-## Replication Steps {#sec-replication_steps2}
-The pull, preparation, and analysis step would remain the same as designed in @sec-replication_steps1 with the only change in data range.
-
-## Results {#sec-results2}
-Shifting the sample to 2007–2023 introduces potential changes in earnings announcement informativeness driven by key economic events.
-The 2008 financial crisis and the 2020 COVID-19 pandemic likely heightened market uncertainty, increasing reliance on earnings disclosures to assess firm stability.
-These developments may have influenced how earnings news is reflected in stock prices and contributed to variation in informativeness across countries.
-
-# Cross-Country Replication {#sec-task3}
+# Cross-Country Replication {#sec-task2}
 This section generalizes the analysis to a non-U.S.
@@
 
-# Research Design for a Non-Archival Study {#sec-task4}
-
-To extend the analysis beyond archival methods, this section proposes a survey to assess whether investors perceive earnings announcements as informative or merely confirmatory, following the triangulation principle in empirical research [@bloomfield_gathering_2016].
-While @ball_how_2008 measure market reactions via stock returns, surveys can capture investor beliefs directly, offering complementary evidence on earnings informativeness.
-
-A structured survey targeting 300 institutional and retail investors will be distributed through financial networks and platforms (e.g., LinkedIn, CFA societies), with a pilot study refining its clarity.
-Participants will receive a €20 honorarium, and responses will remain anonymous.
-The survey measures perceived informativeness (DV) using Likert-scale and multiple-choice questions, with IVs capturing factors such as investor preparedness and reaction timing.
-
-Follow-up interviews with selected respondents will add qualitative depth, enabling a richer understanding of investor perspectives.
-This mixed-method approach enhances validity and supports generalization of findings beyond price-based evidence.
-
-## Survey Questions
-
-1. **How do you typically prepare for earnings announcements in your investment decisions?**
-    * I conduct in-depth research and adjust my positions in advance 
-    * I monitor but rarely adjust positions pre-announcement
-    * I rely on market consensus, analyst forecasts, and AI-driven insights
-    * I do not make investment decisions based on earnings announcements
-
-2. **How do you typically react to earnings announcements?**
-    * I adjust my investment strategy immediately based on the announcement.
-    * I wait for further analysis before making changes.
-    * I rarely make investment decisions based on earnings announcements.
-
-3. **To what extent do you believe that earnings announcements provide new information beyond what is already reflected in market prices?**
-    * Always
-    * Most of the time
-    * Sometimes
-    * Rarely
-    * Never
-
-4. **Rank the following factors in order of importance when evaluating earnings announcements (1 = most important, 5 = least important):**
-    * Earnings per share (EPS) compared to analyst forecasts
-    * Revenue growth
-    * Management guidance and commentary
-    * Market reaction on the day of the announcement
-    * Industry trends and macroeconomic conditions
-
-5. **Do you consider earnings announcements to be more confirmatory or informative?**
-    * Primarily confirmatory (reinforce existing expectations)
-    * Primarily informative (provide new insights)
-    * A mix of both
-
-## Follow-Up Interview Questions
-
-1. Can you describe a recent instance where an earnings announcement significantly influenced your investment decision?
-2. In your experience, are there specific industries where earnings announcements are more informative than confirmatory?
-3. Do you use earnings announcements differently depending on market conditions (e.g., economic downturn vs. growth periods)?
-
 # Conclusion {#sec-conclusion}

### 2025-05-20 code/python/do_analysis.py

@@ -19,13 +19,15 @@ def main():
     log.info("Computing summary statistics by continent ...")
     summary = compute_summary(df)
     summary.to_csv(cfg["summary_statistics_csv"], index=False)
+    with open(cfg["summary_statistics_pickle"], "wb") as f:
+        pickle.dump(summary, f)
     log.info("Summary statistics saved.")
 
     # === Plot ===
     log.info("Creating and saving figure ...")
     fig = plot_figure(df)
-    fig.savefig(cfg["figure1_save_path"])
-    with open(cfg["figure1_pickle_path"], "wb") as f:
+    fig.savefig(cfg["gapminder_figure_png"])
+    with open(cfg["gapminder_figure_pickle"], "wb") as f:
         pickle.dump(fig, f)
     log.info("Figure saved as PNG and pickle.")
 
@@ -46,29 +48,54 @@ def compute_summary(df):
         )
         .reset_index()
     )
+    
+    # Round numeric columns to 0 decimal places
+    numeric_cols = summary.select_dtypes(include="number").columns
+    summary[numeric_cols] = summary[numeric_cols].round(2)
+
     return summary
 
 
 def plot_figure(df):
     """Scatter plot: GDP per capita vs. Life Expectancy (bubble size = pop)."""
     fig, ax = plt.subplots(figsize=(10, 6))
+    
+    continent_palette = {
+        "Asia": "#1E90FF",        
+        "Europe": "#C71585",      
+        "Africa": "#32CD32",      
+        "Americas": "#FF4500",    
+        "Oceania": "#48D1CC"      
+    }
+
     sns.scatterplot(
         data=df,
         x="gdpPercap",
         y="lifeExp",
         size="pop",
         hue="continent",
+        palette=continent_palette,
         sizes=(40, 800),
         alpha=0.7,
         ax=ax,
-        legend="full"
+        legend="brief"
     )
-    ax.set_xscale("log")
-    ax.set_title("Gapminder 2007: Life Expectancy vs. GDP per Capita")
-    ax.set_xlabel("GDP per Capita (log scale)")
-    ax.set_ylabel("Life Expectancy")
-    return fig
 
+    ax.set_xscale("log")
+    ax.set_title("World Development in 2007")
+    ax.set_xlabel("GDP per Capita [in USD] (log scale)")
+    ax.set_ylabel("Life Expectancy [in years]")
+
+    # Move legend outside the plot
+    ax.legend(
[…]

### 2025-05-21 doc/paper.qmd
build paper step

@@
 
-# Replication of Key Tables and Figures {#sec-task1}
-This section replicates key tables and figures from @ball_how_2008 using daily US stock return data (1972–2006) to establish baseline results.
-By reconstructing Table 1 (Panel A), Table 2, and Figure 1, it evaluates the robustness of the original findings and highlights any discrepancies.
-Daily CRSP data ensures accurate measurement of stock reactions around earnings announcements.
-Replication results are discussed in @sec-results1, with extended analysis in @sec-task2.
+# Descriptive Analysis of Global Development in 2007 {#sec-task1}
+This section explores global development patterns using Gapminder data for the year 2007.
+It presents a continent-level summary statistics table and a scatter plot showing the relationship between GDP per capita and life expectancy.
+These outputs serve as a static snapshot of global well-being during that year and demonstrate how transparent and reproducible workflows can be applied to public datasets.
+Results are later contrasted with Gapminder’s interactive chart from 2021 in @sec-task2.
 
 ## Research Design Choices and Assumptions {#sec-research_design_assumptions1}
-In line with @ball_how_2008, this replication examines U.S.
-earnings announcements and stock returns from 1972 to 2006, following the original study’s sample design, event window (-1 to +1), and return computations.
-Unlike the original, I omit analyst expectations, focusing solely on returns and announcement dates for simplicity.
+In line with @gapminder_health_chart, the analysis uses the Gapminder dataset available via the `gapminder` Python package.
+This dataset contains country-level information on population, GDP per capita (inflation-adjusted), and life expectancy for various years in five-year intervals.
+For this project, I focus exclusively on the year 2007—the most recent year available in the dataset at the time.
 
-To address gaps in methodological detail, several assumptions are applied.
-Event windows follow CRSP trading calendar conventions, shifting dates when earnings fall on non-trading days [@center_for_research_in_security_prices_crsp_crsp_2024].
-While @ball_how_2008 require a minimum of 240 trading days, I relax this constraint, as per course guidance, resulting in a broader sample.
-All firms in Compustat meeting the quarterly announcement and return data criteria are included, regardless of activity status.
-The CRSP/Compustat Merged (CCM) database is used to link identifiers, noting that differences from the original study may arise due to updated data (WRDS version: January 2025).
-As in @ball_how_2008, the analysis is return-based (`ret` in CRSP), excluding bid-ask pricing.
-Earnings are grouped by announcement date (`rdq`), not fiscal period, to align directly with CRSP trading dates.
+To ensure clarity and consistency, several assumptions are applied.
+The analysis is restricted to the year 2007, assuming that cross-sectional variation in life expectancy and GDP per capita during that year sufficiently captures key development patterns.
+No imputation is performed; only complete observations are included.
+Population is treated as a size indicator in the visualization, assuming that within-year population estimates are comparable across countries.
+GDP per capita is used in inflation-adjusted international dollars, as provided by Gapminder, without applying additional currency normalization.
+These choices prioritize interpretability and reproducibility, while acknowledging that dynamic trends and structural differences across regions are not captured in this static snapshot.
+All data originates from the Gapminder Foundation, which curates and harmonizes publicly available indicators to promote global development awareness [@gapminder_foundation_2021].
 
-These assumptions, together with the procedural details in @sec-replication_steps1, guide the replication of @ball_how_2008 and ensure transparency in design and implementation.
+These assumptions, together with the procedural details in @sec-replication_steps1, guide the replication and ensure transparency in design and implementation.
 
 ## Replication Steps {#sec-replication_steps1}
-This section outlines the structured replication process: data pulling, preparation, and analysis, aligned with @ball_how_2008.
+This section outlines the modular workflow used to generate the summary table and figure, aligned with @gapminder_health_chart.
 
-**Step 1: Pulling the Data and Managing the Databases**  
-Data from WRDS combines daily CRSP stock returns (dsf) and Compustat quarterly earnings announcements (fundq), linked via the CRSP-Compustat link table (ccmxpf_lnkhist).
-The `permno` and `gvkey` identifiers ensure accurate firm-year alignment.
-Key variables are `ret` (returns) and `rdq` (announcement dates), pulled for the full 1972–2023 period used in Tasks 1 and 2.
+**Step 1: Pulling the Data**  
+The data is loaded from the `gapminder` Python package, ensuring version consistency and avoiding external download links.
+The 2007 subset is saved in `.parquet` format for modern data handling and versioning.
 
 **Step 2: Data Preparation**  
-The merged dataset includes CRSP, Compustat, and the link table.
-After removing irrelevant linking variables, the dataset is trimmed to key variables: `ret`, `rdq`, `fyr`, `permno`, and `gvkey`.
-Only firms with four earnings announcements per calendar year are retained.
-Firms are also classified by fiscal year-end (`fyr`) to match Table 2 Panels C–D.
-Duplicate or incomplete firm-years are excluded.
+The dataset is checked for missing values, which are reported and excluded.
+All numerical columns are rounded to two decimal places.
+No transformations, sorting, or new variables are added to preserve the original data structure.
 
 **Step 3: Analysis Implementation and Reproduction of Tables and Figure**  
-BHRs are computed using the cumulative return formula per @gundersen_returns_2022.
-Cross-sectional regressions are then run for each firm-year to replicate Table 1, Table 2, and Figure 1 from the original paper.
+In the final step, two key outputs are generated to illustrate global development patterns in 2007.
+First, a summary table is produced, aggregating life expectancy, GDP per capita, and total population figures by continent.
+This table provides a concise overview of regional development differences.
+Second, a scatter plot is created to visualize the relationship between GDP per capita and life expectancy, where each country’s bubble size corresponds to its population and color represents its continent.
 
 ## Results {#sec-results1}
-This section compares the static 2007 snapshot with Gapminder’s dynamic global development chart @fig-gapminder_original [@gapminder_health_chart].
+This section presents a static snapshot of global development in 2007 using the Gapminder dataset.
+The analysis includes both a summary table by continent and a scatter plot visualizing the relationship between GDP per capita and life expectancy.
+The figure and table illustrate clear regional disparities in wealth and health outcomes—highlighting, for example, the relatively high life expectancy in Europe and Oceania compared to lower-income regions like Africa.
 
-```{r}
-#| label: fig-gapminder_original
-#| fig-cap: "Original World Health Chart 2021 from Gapminder Foundation"
-#| echo: false
-#| fig-align: center
-knitr::include_graphics("countries_health_wealth_2021.pdf")
-```
+The summary table (see @tbl-gapminder2007_summary) provides aggregated statistics by continent and complements the visual by quantifying central tendencies and demographic scale.
+
+The scatter plot (see @fig-gapminder2007_plot) shows a positive relationship between economic prosperity and life expectancy, with bubble sizes reflecting country population.
+The logarithmic x-axis reveals substantial income inequality while preserving interpretability across income levels.
+The legend is placed outside the plot area for clarity.
 
@@
 
+This section compares the static 2007 snapshot with Gapminder’s dynamic global development chart @fig-gapminder_original [@gapminder_health_chart].
+```{r}
+#| label: fig-gapminder_original
+#| fig-cap: "Original World Health Chart 2021 from Gapminder Foundation"
+#| echo: false
+#| fig-align: center
+knitr::include_graphics("countries_health_wealth_2021.pdf")
+```
+
 # Conclusion {#sec-conclusion}

### 2025-05-22 doc/paper.qmd
rendered figure

@@
 \textbf{GDP}   & Gross Domestic Product \\
-\textbf{IDE}   & Integrated Development Environment \\
 \textbf{TREAT} & TRR 266 Template for Reproducible Empirical Accounting Research \\
@@
 These choices prioritize interpretability and reproducibility, while acknowledging that dynamic trends and structural differences across regions are not captured in this static snapshot.
-All data originates from the Gapminder Foundation, which curates and harmonizes publicly available indicators to promote global development awareness [@gapminder_foundation_2021].
+All data originates from the Gapminder Foundation, which curates and harmonizes publicly available indicators to promote global development awareness [@gapminder_health_chart].
 
@@
 This section presents a static snapshot of global development in 2007 using the Gapminder dataset.
-The analysis includes both a summary table by continent and a scatter plot visualizing the relationship between GDP per capita and life expectancy.
+The analysis includes both a summary table by continent and a scatter plot visualizing the relationship between gross domestic product (GDP) per capita and life expectancy.
 The figure and table illustrate clear regional disparities in wealth and health outcomes—highlighting, for example, the relatively high life expectancy in Europe and Oceania compared to lower-income regions like Africa.
@@
 The summary table (see @tbl-gapminder2007_summary) provides aggregated statistics by continent and complements the visual by quantifying central tendencies and demographic scale.
+It presents average and median values for life expectancy and GDP per capita, along with total population and number of countries represented per continent in 2007.
 
-The scatter plot (see @fig-gapminder2007_plot) shows a positive relationship between economic prosperity and life expectancy, with bubble sizes reflecting country population.
-The logarithmic x-axis reveals substantial income inequality while preserving interpretability across income levels.
-The legend is placed outside the plot area for clarity.
+@fig-gapminder2007_plot illustrates a scatter plot based on Gapminder’s 2007 data, mapping GDP per capita against life expectancy.
+Each country is represented as a bubble whose size corresponds to its population and whose color indicates its continent.
+The plot reveals a positive relationship between economic prosperity and life expectancy.
+The use of a logarithmic x-axis emphasizes differences in income levels across regions while maintaining interpretability across a wide range of values.
+
+```{r}
+#| label: fig-gapminder2007_plot
+#| fig-cap: "Gapminder 2007: Life Expectancy vs. GDP per Capita"
+#| echo: false
+#| fig-align: center
+knitr::include_graphics("../output/gapminder_figure.png")
+```
 

### 2025-05-22 code/python/do_analysis.py

@@ -27,9 +27,7 @@ def main():
     log.info("Creating and saving figure ...")
     fig = plot_figure(df)
     fig.savefig(cfg["gapminder_figure_png"])
-    with open(cfg["gapminder_figure_pickle"], "wb") as f:
-        pickle.dump(fig, f)
-    log.info("Figure saved as PNG and pickle.")
+    log.info("Figure saved as PNG.")
 
 
 def compute_summary(df):
@@ -48,7 +46,7 @@ def compute_summary(df):
         )
         .reset_index()
     )
-    
+
     # Round numeric columns to 0 decimal places
     numeric_cols = summary.select_dtypes(include="number").columns
     summary[numeric_cols] = summary[numeric_cols].round(2)
@@ -59,13 +57,13 @@ def compute_summary(df):
 def plot_figure(df):
     """Scatter plot: GDP per capita vs. Life Expectancy (bubble size = pop)."""
     fig, ax = plt.subplots(figsize=(10, 6))
-    
+
     continent_palette = {
-        "Asia": "#1E90FF",        
-        "Europe": "#C71585",      
-        "Africa": "#32CD32",      
-        "Americas": "#FF4500",    
-        "Oceania": "#48D1CC"      
+        "Asia": "#1E90FF",
+        "Europe": "#C71585",
+        "Africa": "#32CD32",
+        "Americas": "#FF4500",
+        "Oceania": "#48D1CC"
     }
 
     sns.scatterplot(
@@ -96,6 +94,7 @@ def plot_figure(df):
     plt.tight_layout(pad=2.5)
 
     return fig
+
 
 if __name__ == "__main__":
     main()

### 2025-05-23 doc/paper.qmd
table rendered

@@
 \pagebreak
+
+```{python}
+#| echo: false
+#| output: false
+
+import pickle
+import pandas as pd
+
+# Load the prepared summary table
+with open('../output/summary_statistics.pickle', 'rb') as f:
+    summary = pickle.load(f)
+
+# Clean column names for LaTeX rendering
+def escape_special_characters(df):
+    df.columns = [col.replace('_', '\\_').replace('#', '\\#') for col in df.columns]
+    return df
+
+# Formatting for float values
+def custom_float_format(x):
+    return f"{x:.2f}" if pd.notnull(x) else ""
+
+# Function to prep LaTeX table
+def prep_latex_table(df, caption=None, label=None):
+    df = escape_special_characters(df.reset_index(drop=True))
+    num_columns = len(df.columns)
+    column_format = 'l' + 'r' * (num_columns - 1)
+    latex_table = df.to_latex(
+        column_format=column_format,
+        index=False,
+        float_format=custom_float_format,
+    )
+    latex_table_lines = [
+        "\\begin{table}",
+        "\\centering",
+        "\\resizebox{\\textwidth}{!}{%",
+        latex_table,
+        "}",
+        f"\\caption{{{caption}}}" if caption else "",
+        f"\\label{{{label}}}" if label else "",
+        "\\end{table}"
+    ]
+    return "\n".join(line for line in latex_table_lines if line)
+```
 
@@
 
+```{python}
+#| label: tbl-gapminder2007_summary
+#| tbl-cap: "Summary Statistics by Continent (Gapminder 2007)"
+#| echo: false
+#| output: asis
+
+print(prep_latex_table(summary))
+print("\n\\caption*{\\textnormal{This table reports the mean and median life expectancy and GDP per capita, along with the total population and number of countries per continent for the year 2007.}}")
+```
+
 @fig-gapminder2007_plot illustrates a scatter plot based on Gapminder’s 2007 data, mapping GDP per capita against life expectancy.
@@
 
-```{r}
+```{python}
 #| label: fig-gapminder2007_plot
@@
 #| echo: false
-#| fig-align: center
-knitr::include_graphics("../output/gapminder_figure.png")
+
+from IPython.display import Image
+Image(filename="../output/gapminder_figure.png")
 ```
+
+Together, these outputs provide a clear and interpretable overview of global development disparities, offering insights into how income and health outcomes align across continents at a given point in time.
 
@@
 This section compares the static 2007 snapshot with Gapminder’s dynamic global development chart @fig-gapminder_original [@gapminder_health_chart].
-```{r}
+```{python}
 #| label: fig-gapminder_original
@@
 #| echo: false
-#| fig-align: center
-knitr::include_graphics("countries_health_wealth_2021.pdf")
+
+from IPython.display import Image
+Image(filename="countries_health_wealth_2021-1.png")
 ```

### 2025-05-23 doc/paper.qmd
table caption added

@@
 # Function to prep LaTeX table
-def prep_latex_table(df, caption=None, label=None):
+# Function to prep LaTeX table with optional description
+def prep_latex_table(df, caption=None, label=None, description=None):
     df = escape_special_characters(df.reset_index(drop=True))
@@
     column_format = 'l' + 'r' * (num_columns - 1)
+
     latex_table = df.to_latex(
@@
         float_format=custom_float_format,
+        escape=False
     )
+
+    latex_caption = f"\\caption{{{caption}}}" if caption else ""
+    latex_description = f"\\caption*{{\\textnormal{{{description}}}}}" if description else ""
+    latex_label = f"\\label{{{label}}}" if label else ""
+
     latex_table_lines = [
@@
         latex_table,
-        "}",
-        f"\\caption{{{caption}}}" if caption else "",
-        f"\\label{{{label}}}" if label else "",
+        "}",  # End resizebox
+        latex_caption,
+        latex_description,
+        latex_label,
         "\\end{table}"
@@
 #| output: asis
-
-print(prep_latex_table(summary))
-print("\n\\caption*{\\textnormal{This table reports the mean and median life expectancy and GDP per capita, along with the total population and number of countries per continent for the year 2007.}}")
+#| 
+print(prep_latex_table(
+    summary,
+    description="\\fontsize{10pt}{12pt}\\selectfont This table reports the mean and median life expectancy and GDP per capita, along with the total population and number of countries per continent for the year 2007."
+))
 ```

### 2025-05-23 doc/paper.qmd
add packages

@@
 header-includes:
-  - \usepackage[nolists]{endfloat}
+  - \usepackage[nolists]{endfloat} # if commenting out, you place figures&tables inside text
+  - \usepackage{float} # in order to place tables correctly
+  - \floatplacement{table}{H} # in order to place tables correctly
   - \usepackage{setspace}\onehalfspacing # 1.5 line spacing acc. to Accounting guidelines
@@
   - \usepackage{caption}
-  - \captionsetup[table]{skip=24pt,font=bf}
+  - \captionsetup[table]{skip=15pt,font=bf} # was 24pt initially
+  - \captionsetup[figure]{skip=15pt, font=bf} # consistency with tables
   - \usepackage{array}
@@
 #| output: asis
-#| 
+
 print(prep_latex_table(

### 2025-05-23 doc/paper.qmd
paper done; analysis adjusted

@@
         "\\centering",
+        "\\fontsize{12pt}{16pt}\\selectfont", #  12pt to adjust font size acc. to guidelines, 16pt for 1.5 line spacing in table
         "\\resizebox{\\textwidth}{!}{%",
@@
 This project serves as a template-based example for conducting and documenting reproducible empirical data analysis using publicly available datasets.
-It leverages the TRR 266 Template for Reproducible Empirical Accounting Research (TREAT) and builds upon the methodological foundations of the Corporate Decision-Making and Quantitative Analysis, as well as the Accounting Reading Group course.
-To support future empirical thesis projects, this template was adapted and developed specifically for students at the School of Business and Economics at Humboldt-Universität zu Berlin—particularly those affiliated with the Institute of Accounting and Auditing and the Finance Group.
-The template was developed in adherence to the formal formatting and content guidelines of the School of Business and Economics, and in particular, those of the Institute of Accounting and Auditing.
+It leverages the TRR 266 Template for Reproducible Empirical Accounting Research (TREAT) and builds upon the methodological foundations of the Corporate Decision-Making and Quantitative Analysis course, as well as the Accounting Reading Group.
+To support future empirical thesis projects, this template was developed in adherence to the formal formatting and content guidelines of the Institute of Accounting and Auditing at the School of Business and Economics, Humboldt-Universität zu Berlin.
 It is not a completed thesis but a demonstrative project hosted in the [template repository](https://github.com/melmzv/huwiwiQuartoThesis).
@@
 This template project explores global development patterns using the Gapminder dataset, focusing on the year 2007.
-The project illustrates how reproducible research workflows can be applied beyond traditional financial datasets by using public and open-source data.
-Specifically, it presents a summary statistics table by continent and visualizes the relationship between GDP per capita and life expectancy, offering a static snapshot of world development during 2007.
-These descriptive results are presented in @sec-task1.
+The project illustrates how reproducible research workflows can be applied by using public and open-source data.
+Specifically, it presents a summary statistics table by continent and visualizes the relationship between GDP per capita and life expectancy, offering a static snapshot of world development in 2007.
+These descriptive insights are presented in @sec-task1.
 
-The results are further contrasted with Gapminder’s interactive global development visualizations as of 2021 [@gapminder_health_chart], highlighting the potential and limits of static data analysis.
+The results are further contrasted with Gapminder’s interactive global development visualizations as of 2021 [@gapminder_health_chart].
 This comparative discussion is provided in @sec-task2.
@@
 These outputs serve as a static snapshot of global well-being during that year and demonstrate how transparent and reproducible workflows can be applied to public datasets.
-Results are later contrasted with Gapminder’s interactive chart from 2021 in @sec-task2.
 
@@
 This dataset contains country-level information on population, GDP per capita (inflation-adjusted), and life expectancy for various years in five-year intervals.
-For this project, I focus exclusively on the year 2007—the most recent year available in the dataset at the time.
+For this project, I focus on the year 2007 - the most recent year available in the dataset at the time.
 
@@
 No imputation is performed; only complete observations are included.
-Population is treated as a size indicator in the visualization, assuming that within-year population estimates are comparable across countries.
-GDP per capita is used in inflation-adjusted international dollars, as provided by Gapminder, without applying additional currency normalization.
+GDP per capita is used in inflation-adjusted international dollars, as provided by @gapminder_health_chart, without applying additional currency normalization.
 These choices prioritize interpretability and reproducibility, while acknowledging that dynamic trends and structural differences across regions are not captured in this static snapshot.
@@
 **Step 1: Pulling the Data**  
-The data is loaded from the `gapminder` Python package, ensuring version consistency and avoiding external download links.
-The 2007 subset is saved in `.parquet` format for modern data handling and versioning.
+The data is loaded from the `gapminder` Python package. The 2007 subset is saved in `.parquet` format for modern data handling and versioning.
 
@@
 This section presents a static snapshot of global development in 2007 using the Gapminder dataset.
-The analysis includes both a summary table by continent and a scatter plot visualizing the relationship between gross domestic product (GDP) per capita and life expectancy.
-The figure and table illustrate clear regional disparities in wealth and health outcomes—highlighting, for example, the relatively high life expectancy in Europe and Oceania compared to lower-income regions like Africa.
+The figure and table illustrate clear regional disparities in wealth and health outcomes - highlighting, for example, the relatively high life expectancy in Europe and Oceania compared to lower-income regions like Africa.
 
-The summary table (see @tbl-gapminder2007_summary) provides aggregated statistics by continent and complements the visual by quantifying central tendencies and demographic scale.
+@tbl-gapminder2007_summary provides aggregated statistics by continent and complements the visual by quantifying central tendencies and demographic scale.
 It presents average and median values for life expectancy and GDP per capita, along with total population and number of countries represented per continent in 2007.
+A total of `{python} f"{summary['Num_Countries'].sum():,}"` countries are included in the 2007 snapshot.
 
@@
 The plot reveals a positive relationship between economic prosperity and life expectancy.
-The use of a logarithmic x-axis emphasizes differences in income levels across regions while maintaining interpretability across a wide range of values.
 
@@
 
-# Cross-Country Replication {#sec-task2}
-This section generalizes the analysis to a non-U.S.
-market by replicating Figure 1 using 1972–2023 data from Canada.
-This involves aligning CRSP/Compustat variables with Worldscope and Datastream equivalents while accounting for differences in reporting standards, market liquidity, and institutional context.
-Though the core methodology is retained, adapting it to international data demands careful mapping and interpretation.
+# Global Development Patterns in 2021 {#sec-task2}
+This section compares the 2007 snapshot with Gapminder’s 2021 visualization to assess whether key patterns - like the link between GDP per capita and life expectancy - persist and to highlight shifts in global inequality and regional development.
 
-## Research Design Choices and Assumptions {#sec-research_design_assumptions3}
-Since this secition applies a generalization and extension approach to a non-U.S.
-market, it requires adapting the methodology to alternative databases and market structures.
-Like in the original study, no analyst forecasts or earnings surprises are used—only stock returns and earnings announcement dates are required.
-Canada was selected for its comparable quarterly reporting frequency and institutional similarity to the U.S., making it a suitable choice for cross-country replication [@short_short_2025].
-Canadian firms reporting under U.S.
-GAAP further enhance comparability.
+## Comparative Assumptions and Design Choices {#sec-research_design_assumptions3}
+This section outlines the assumptions for comparing the 2007 Gapminder dataset to a visualization of global development patterns in 2021.
 
-To address database-specific issues not covered in @sec-research_design_assumptions1, I make the following adjustments: First, Worldscope/Datastream workflows are kept separate from CRSP/Compustat to prevent interference and facilitate debugging, given the scale of data involved.
-Second, I exclude fiscal-year-end-based subgroup analysis (Table 2 Panels C–D), as Figure 1 uses the full sample without splitting by fiscal period.
-Third, I use Worldscope items 5901–5904 to define earnings dates, though limited availability before 1992 may reduce the sample size [@thomson_financial_worldscope_2007, p.281].
-Fourth, differences in update frequency between quarterly Worldscope and weekly Datastream can lead to timing mismatches when aligning fundamentals and returns.
-Fifth, I rely on Datastream’s `ret` variable for percentage returns, ignoring bid/ask spreads.
-Finally, currency effects are ignored, as percentage returns are unaffected by currency denomination.
+The assumptions guiding the comparison are as follows: visual inspection is used to assess trends in life expectancy and income per capita; bubble size reflects each country’s population, as defined by Gapminder; and no new calculations are applied - the 2021 snapshot displays curated values from the Gapminder Foundation [-@gapminder_health_chart].
 
-## Replication Steps {#sec-replication_steps3}
-The process is same as in @sec-task1.
-
-**Step 1: Pulling the Data and Managing the Databases**  
-Following @dai_research_2021, I link Worldscope and Datastream using code and infocode.
-Since Worldscope’s `year_` starts in 1980 [@wharton_research_data_services_worldscope_2025], Datastream data is also restricted to 1980–2023 to ensure consistent coverage.
-I extract quarterly earnings announcement dates (items 5901–5904) from Worldscope and daily stock returns from Datastream, focusing on Canadian firms and filtering for common equity (typecode = EQ).
-This step yields three datasets: 27.9M (Datastream), 77.9K (link table), and 1.94M (Worldscope) observations.
-
-**Step 2: Data Preparation**  
-Worldscope is first merged with the link table via code, producing 747,760 rows.
-This is then joined with Datastream via infocode, resulting in 10.2M observations.
-I exclude firms missing any earnings date fields, removing 3.6M rows, and drop 6M rows where earnings do not span all four quarters within the same calendar year.
-To retain more valid windows, event days (-1, 0, 1) are dynamically shifted within ±3 days when needed.
-After cleaning, I compute 3-day BHRs and extract full-year stock return data from Datastream.
-All retained infocode entries have around 250 trading days per year, confirming data completeness for annual BHR computation.
-
-**Step 3: Analysis Implementation and Reproduction of Tables and Figure**  
-The analysis replicates Table 1 (summary statistics), Table 2 (yearly regressions), and Figure 1 (Abnormal R² and slope trends).
-Using `BHR_3day` and `BHR_Annual`, I run yearly regressions of annual returns on earnings-window returns and compute Abnormal R² (adjusted R² – 4.8%), handling missing values appropriately.
-
-Specifically, buy-and-hold return over the three-day event window, that measures the stock’s reaction only to earnings news, is computed as shown in @eq-bhr_event:
+The relationship between country-level variables and their graphical representation in the visualizations can be summarized as shown in @eq-bubble_mapping:
 $$
-\begin{aligned}
-BHR_{\text{event}} &= (1 + R_{t_1}) (1 + R_{t_2}) \dots (1 + R_{t_T}) - 1,
-\\
-\text{where } R_t & \text{ is the daily return and } T \text{ is the total trading days in a year.}
-\end{aligned}
-$$ {#eq-bhr_event}
+\text{Bubble}_{i} = f\left(\text{GDP}_i,\ \text{LifeExp}_i,\ \text{Pop}_i,\ \text{Continent}_i\right)
+$$ {#eq-bubble_mapping}
 
 ## Results {#sec-results3}
-Comparing the replication results with those from @ball_how_2008 reveals both similarities and discrepancies.
-The original figure shows a general upward trend in abnormal  R² , peaking in the early 2000s, whereas the replicated figure captures greater volatility in the later years (2010–2023).
-Additionally, Ball’s figure suggests a smoother long-term dynamic, while the replication exhibits more pronounced fluctuations, particularly post-2010.
-Panel B in replicated figure exhibits higher volatility and larger coefficient magnitudes, suggesting increased earnings informativeness post-2006, while Ball’s original figure has more stable and lower-magnitude coefficients.
-
-The post-2000 period in the replication shows more inconsistent abnormal  R²  values, possibly reflecting evolving market structures, shifts in disclosure practices, and increased earnings informativeness following regulatory changes (e.g., IFRS adoption, post-SOX adjustments, different financial market structure).
-The pronounced fluctuations could also be driven by global financial crisis, Canada's higher market concentration, fewer publicly traded firms, and potential regulatory differences, which may affect earnings informativeness and stock return patterns.
-The variability in slope coefficients further suggests that earnings announcements’ impact on stock returns is less stable than in prior decades, potentially due to differences in investor response.
-
-This section compares the static 2007 snapshot with Gapminder’s dynamic global development chart @fig-gapminder_original [@gapminder_health_chart].
+This section compares the 2007 snapshot, rendered from Gapminder data in Python in @sec-task1, with Gapminder’s official visualization of global development in 2021 presented below.
 ```{python}
@@
 Using Quarto proved to be a particularly effective and user-friendly solution for structuring and rendering the thesis.
-It facilitates seamless integration of code, results, and narrative, making it an ideal tool for students at the Institute of Accounting and Auditing or the Finance Group at the School of Business and Economics, HU Berlin.
+It facilitates seamless integration of code, results, and narrative, making it an ideal tool for students at the Institute of Accounting and Auditing at the School of Business and Economics, HU Berlin.
 Thanks for reading!

### 2025-05-23 code/python/do_analysis.py

@@ -18,7 +18,6 @@ def main():
     # === Summary statistics ===
     log.info("Computing summary statistics by continent ...")
     summary = compute_summary(df)
-    summary.to_csv(cfg["summary_statistics_csv"], index=False)
     with open(cfg["summary_statistics_pickle"], "wb") as f:
         pickle.dump(summary, f)
     log.info("Summary statistics saved.")

### 2025-05-24 doc/paper.qmd
Makefile configuration

@@
 
-The assumptions guiding the comparison are as follows: visual inspection is used to assess trends in life expectancy and income per capita; bubble size reflects each country’s population, as defined by Gapminder; and no new calculations are applied - the 2021 snapshot displays curated values from the Gapminder Foundation [-@gapminder_health_chart].
+The assumptions guiding the comparison are as follows: visual inspection is used to assess trends in life expectancy and income per capita; bubble size reflects each country’s population, as defined by Gapminder; no new calculations are applied - the 2021 snapshot displays curated values from the Gapminder Foundation [-@gapminder_health_chart].
 

### 2025-05-24 code/python/prepare_data.py

@@ -20,7 +20,7 @@ def main():
 
     os.makedirs(os.path.dirname(cfg['output_path']), exist_ok=True)
     df.to_parquet(cfg['output_path'])
-    log.info("Prepared data saved.")
+    log.info("Prepared data is now saved.")
 
 if __name__ == '__main__':
     main()

### 2025-05-25 doc/paper.qmd
png adapted

@@
 This comparative discussion is provided in @sec-task2.
-The project concludes with reflections on transparency, reproducibility, and the importance of making research accessible to a broader audience, as discussed in @sec-conclusion.
+The project concludes with reflections on transparency, reproducibility, and importance of making research accessible to a broader audience, as discussed in @sec-conclusion.
 
